### 2015-11-30 documents/paper/paper.tex
Started paper

\documentclass[journal]{IEEEtran}
\usepackage{blindtext}
\usepackage{graphicx}
\usepackage{algorithm2e}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{tabularx}
\usepackage{float}
\usepackage{bm}

\begin{document}

\title{GPU-Accelerated Parallelization of Edge-Based Motion Detection Algorithms with the NVIDIA CUDA Framework}

\author{
	Emilio Del Vecchio, Kevin Lin, Senthil Natarajan\\
	Department of Electrical and Computer Engineering, Rice University, Houston, TX\\
	\{edd5, kevinlin, ssn3\}@rice.edu
}

\maketitle

\begin{abstract}
Graphics processing units (GPUs) are rapidly gaining popularity as a platform for parallelized computations on massive sets of data. Since much of the computations in image processing and computer vision are easily parallelized, graphics operations on GPUs achieve significant speedups compared to those done on their serial, CPU counterparts. Further, SDKs like the NVIDIA CUDA framework provide developers easy APIs to take advantage of the parallel computing power of GPUs. In this paper, we construct a GPU-based method of real-time motion detection based on edges computed from frames of a continuous video stream, and demonstrate the speedup we can achieve compared to reference CPU implementation.
\end{abstract}

\section{Introduction}
Placeholder text

\subsection{GPU Computation}
Placeholder text

\subsection{The NVIDIA CUDA Framework}
Placeholder text

\section{Edge Detection Algorithm}
Edge detection is the process of determining the locations of boundaries that separate objects, or sections of objects, in an image. This edge data simplifies the features of an image to only its basic outlines, which makes it a convenient input for many other problems in computer vision, including motion detection and tracking, object classification, three-dimensional reconstruction, and others. The edge detection algorithm we present is composed of three stages:
\begin{enumerate}
	\item Reduce the amount of high-frequency noise from the image using a two-dimensional low pass filter.
	\item Determine the two-dimensional gradient of the filtered image by applying partial derivatives in both the horizontal and vertical directions.
	\item Classify edges by suppressing surrounding pixels in the gradient direction whose magnitude is lesser, and applying selective thresholding to a limited apron around each selected pixel.
\end{enumerate}
Edges correspond to the points on the image that exhibit a high gradient magnitude. The edge data from our algorithm is then used as input to our motion detection algorithm.

\subsection{Separable Convolution}
The filtering necessary in steps (1) and (2) from our algorithm above involve the convolution of a two-dimensional image and a two-dimensional kernel filter of constant width and height. A na\"{\i}ve implementation of two-dimensional convolution would require a double sum across the kernel and image for each individual pixel:
\begin{equation}
	\label{2d-naive-convolution}
	y[m, n] = x[m, n] * h[m, n] = \sum_{i = -\infty}^{\infty} \sum_{j = -\infty}^{\infty} h[i, j] x[m - i, n - j]
\end{equation}
For an image of size $M$ x $N$ and kernel of size $k$, a direct implementation would require $O(MNk^2)$ time, which is infeasible for a real-time implementation. Instead, it is possible to exploit the separable property of certain convolution matrices $h$, e.g. that $h$ can be represented as the product of two one-dimensional matrices $h_1$ and $h_2$. This effectively reduces our two-dimensional convolution of \eqref{2d-naive-convolution} to two separate instances of one-dimensional convolution:
$$y[m, n] = h_1[n] * (h_2[m] * x[m, n])$$
which is implemented as
$$y[m, n] = \sum_{i = -\infty}^{\infty}h_1[i] \sum_{j = -\infty}^{\infty}h_2[j] x[m - i, n - j]$$
For the same $M$ x $N$ input and square kernel of size $k$, a separable implementation reduces the computational complexity to $O(MNk)$. By comparison, an implementation using the FFT would cost $O(MN\log{MN})$. In practice, consider the following results for a convolution of a $3$ x $3$ kernel with an image of size $2500$ x $2500$:
\begin{table}[H]
	\small
	\centering
	\caption{Comparison of Computational Complexity of Different Convolution Implementation Algorithms}
	\label{convolution-complexity-comparison}
	\begin{tabular}{llll}
	 & Complexity & Execution time (s) \\
	 Direct convolution & $O(MNk^2)$ & $3$ \\
	 Separated convolution & $O(MNk)$ & $3$ \\
	 FFT & $O(MN\log{MN})$ & $3$
	\end{tabular}
\end{table}
As detailed in later sections, we take advantage of the fast computational complexity of separable convolution in our implementation, as all of the filters we apply are separable into one-dimensional matrices.

\subsection{Noise Elimination: Two-Dimensional Gaussian Filtering}
The first task in our edge detection algorithm is denoising the input in order to reduce the amount of high-frequency content in the image by as much as possible without destroying critical information points in the image (e.g. the real edges). We filter out high-frequency noise so that random noise is not mistakenly interpreted as an edge, as edges correspond to points in the image where the gradient has an above-threshold magnitude. \par
For example, consider the following $5312$ x $2988$ pixel image (taken at a high ISO to deliberately introduce noise) and a plot of its grayscale intensity versus the image's spatial dimensions:
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{noisy_image.jpg}
	\caption{Unfiltered noisy image}
    \label{noisy-image}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{noisy_image_mesh.jpg}
	\caption{Grayscale intensity of unfiltered noisy image at each pixel}
    \label{noisy-image-mesh}
\end{figure}
To low-pass filter our image, we apply a discrete Gaussian filter. Generally, a Gaussian blur kernel of size $2n + 1$ x $2n + 1$ ($\forall n \in \mathbb{Z}^+$, and with parameter $\sigma$) is given by
$$h[i, j] = \frac{1}{2\pi \sigma^2}e^{-\frac{(i - k - 1)^2 + (j - k - 1)^2}{2\sigma^2}}$$
Our implementation of Gaussian filtering uses the constant-sized ($k = 3$), constant-$\sigma$ kernel
\[
h[i, j] = \frac{1}{16}
\begin{bmatrix}
	1 & 2 & 1 \\
	2 & 4 & 2 \\
	1 & 2 & 1
\end{bmatrix}
\]
This kernel is implemented separably as
\[
h[i, j] = \frac{1}{16}
\begin{bmatrix}
	1 & 2 & 1
\end{bmatrix}
*
\begin{bmatrix}
	1 \\ 2 \\ 1
\end{bmatrix}
\] \par
Applying a Gaussian filter with parameters $k = 5$ and $\sigma = 5$ to the above image significantly denoises the image without sacrificing edge precision, which can be seen from the spatial intensity plot in Figure \ref{filtered-noisy-image-mesh}.
\begin{figure}[H]
	\centering
	\includegraphics[width=0.45\textwidth]{filtered_noisy_image.jpg}
	\caption{Gaussian-filtered noisy image with parameters $k = 3$ and $\sigma = 5$}
    \label{filtered-noisy-image}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{filtered_noisy_image_mesh.jpg}
	\caption{Grayscale intensity of Gaussian-filtered noisy image at each pixel}
    \label{filtered-noisy-image-mesh}
\end{figure}

\subsection{Gradient Computation: The Sobel Operator}
Edges in the image correspond to pixel locations at which there is a rapid change in intensity with respect to the image's spatial dimensions. Thus, edge pixels are defined as those whose gradient magnitude $||\boldsymbol{G}||$ is maximized along the gradient direction $\theta_G$.
\begin{equation}
	\label{gradient-magnitude}
	||\boldsymbol{G}_{i,j}|| = \sqrt{G_x^2 + G_y^2}
\end{equation}
\begin{equation}
	\label{gradient-angle}
	\theta_G = \arctan{\frac{G_y}{G_x}}
\end{equation}
where $G_x$ and $G_y$ are the values of the partial derivatives along the $x$ and $y$ directions, respectively, at the pixel located at $(i, j)$ of the image $A$.
$$G_x = \frac{\partial \boldsymbol{A}}{\partial x}\Bigr|_{\substack{(x, y) = (i, j)}}$$
$$G_y = \frac{\partial \boldsymbol{A}}{\partial y}\Bigr|_{\substack{(x, y) = (i, j)}}$$
\par A variety of different discrete differentiation operators exist to approximate the gradient of the image. The operator we choose in our implementation is the Sobel operator.
\[
\boldsymbol{G_x} =
\begin{bmatrix}
	-1 & 0 & 1 \\
	-2 & 0 & 2 \\
	-1 & 0 & 1
\end{bmatrix} * \boldsymbol{A}
\]
\[
\boldsymbol{G_y} =
\begin{bmatrix}
	-1 & -2 & -1 \\
	0 & 0 & 0 \\
	1 & 2 & 1
\end{bmatrix} * \boldsymbol{A}
\]
Like our Gaussian kernel, the Sobel operator (in both the $x$ and $y$ directions) is separable:
\[
\boldsymbol{G_x} =
\begin{bmatrix}
	1 \\
	2 \\
	1
\end{bmatrix} * 
(\begin{bmatrix}
	1 & 0 & -1
\end{bmatrix} * \boldsymbol{A})
\]
\[
\boldsymbol{G_y} =
\begin{bmatrix}
	1 \\
	0 \\
	-1
\end{bmatrix} * 
(\begin{bmatrix}
	1 & 2 & 1
\end{bmatrix} * \boldsymbol{A})
\]
\par Thus, we perform two separable convolutions with the separated Sobel filter on the Gaussian-filtered image to obtain gradients in the horizontal and vertical directions, and determine the magnitude and angle matrices $\boldsymbol{G}$ and $\boldsymbol{\theta}$ from equations \eqref{gradient-magnitude} and \eqref{gradient-angle}.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{pcb.jpg}
	\caption{Original image input to Sobel filter}
    \label{pcb-original}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{sobel_pcb.jpg}
	\caption{Magnitude of the gradient $||\boldsymbol{G}||$, as determined by applying a Sobel filter in both the $x$ and $y$ directions, at each pixel of the Gaussian-filtered original image with parameters $k = 5$ and $\sigma = 5$}
    \label{sobel-pcb}
\end{figure}
\begin{figure}[h]
	\centering
	\includegraphics[width=0.45\textwidth]{sobel_pcb_mesh.jpg}
	\caption{Gradient of \ref{sobel-pcb} visualized as a three-dimensional mesh}
    \label{sobel-pcb}
\end{figure}

\subsection{Non-Maximum Suppression and Selective Thresholding}
Given the gradient magnitude and direction of the Gaussian-filtered image, the final step of the edge detection algorithm is to determine which pixels should be selected as edge pixels, and which should be rejected. We define two procedures, \textit{non-maximum suppression} and \textit{selective thresholding}, to gain a reasonably accurate edge map from the gradient.
\par Non-maximum suppression selects and rejects edge pixels according to the following criteria:
\begin{enumerate}
	\item Pixels whose gradient magnitude is below a user-defined low threshold, $t_l$, is immediately rejected
	\item If the gradient magnitude of a pixel is greater than that of the pixels in either direction of the gradient angle of that pixel, then the pixel is accepted if its gradient magnitude is greater than a user-defined high threshold, $t_h$
\end{enumerate}
More precisely, a pixel at $(i, j)$ with gradient magnitude $G_{i, j} > t_l$ and angle $\theta$ is accepted if any of the following hold true:
\begin{itemize}
	\item $(-\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi) \wedge G_{i, j} > G_{i, j \pm 1}$
	\item $(\frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}) \wedge G_{i, j} > G_{i \pm 1, \pm j}$
	\item $(\frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}) \wedge G_{i, j} > G_{i \pm 1, j}$
	\item $(\frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}) \wedge G_{i, j} > G_{i \pm 1, \mp j}$
\end{itemize}
The selective thresholding technique extends on the procedure of non-maximum suppression. For any pixel that passes crtiteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel. If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted. This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.

\section{Motion Detection Algorithm}
Placeholder text

\subsection{Thresholded Frame-by-Frame Difference Map}
Placeholder text

\subsection{Motion Area Estimation via a Spatial Difference Density Map}
Placeholder text

\section{GPU Speedup Results}
Placeholder text

\section{Conclusion}
Placeholder text

\appendices
\section{Proof of the First Zonklar Equation}
Some text for the appendix.

\section*{Acknowledgment}
\begin{thebibliography}{1}

\bibitem{IEEEhowto:kopka}
H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.

\end{thebibliography}
\end{document}

### 2015-11-30 documents/paper/paper.tex
Add introduction and non-maximum suppression and selective thresholding algorithm

@@
 \begin{abstract}
+GPUs provide a powerful platform for parallel computations on large data inputs, namely images.
+In this paper, we explore a GPU-based implementation of a simplified adaptation of existing edge detection algorithms fast enough to operate on frames of a continuous video stream in real-time.
+We also demonstrate practical applications of edge detection, including an edge-based method for motion detection estimation.
+Finally, we demonstrate the speedups of over 2x we achieve with GPU parallelism, as compared to a reference serial CPU-based implementation.
+\end{abstract}
+
+\section{Introduction}
 Graphics processing units (GPUs) are rapidly gaining popularity as a platform for parallelized computations on massive sets of data.
@@
 Further, SDKs like the NVIDIA CUDA framework provide developers easy APIs to take advantage of the parallel computing power of GPUs.
-In this paper, we construct a GPU-based method of real-time motion detection based on edges computed from frames of a continuous video stream, and demonstrate the speedup we can achieve compared to reference CPU implementation.
-\end{abstract}
+We take full advantage of the computational benefits of GPUs by implementing edge detection and motion detection algorithms in CUDA C.
 
-\section{Introduction}
-Placeholder text
-
-\subsection{GPU Computation}
-Placeholder text
-
-\subsection{The NVIDIA CUDA Framework}
-Placeholder text
+\subsection{GPU Computation and the NVIDIA CUDA Framework}
+Formally, NVIDIA CUDA is a parallel computing platform and programming model that exposes familiar C-based APIs for parallelized computations.
+CUDA is NVIDIA's platform for general-purpose computing on graphics processing units (commonly, GPGPU or GP$^2$U), or the use of a GPU for computations traditionally handled by the CPU.
+Generally, GPGPU is used to exploit the improved multithreaded performance and raw floating-point computational ability of GPUs over CPUs.
+For example, on modern hardware, an NVIDIA GeForce GTX 970 (1664 CUDA cores) exhibits peak single-precision floating point performance of nearly 3500 GFLOPS (floating-point operations per second), while an Intel Core i7 4790K (8C, 16T) achieves 100 GFLOPS.
+Our goal is to demonstrate the performance of GPU computing by solving a handful of existing problems in computer vision with CUDA: namely, edge detection and motion detection.
 
@@
 By comparison, an implementation using the FFT would cost $O(MN\log{MN})$.
-In practice, consider the following results for a convolution of a $3$ x $3$ kernel with an image of size $2500$ x $2500$:
+In practice, consider the following results\footnote{Results were obtained by measuring the execution time of a MATLAB script on an Intel Core i7 4790K @ 4.7 GHz.} for a convolution of a $3$ x $3$ kernel with an image of size $2500$ x $2500$:
 \begin{table}[H]
@@
 \par
-For example, consider the following $5312$ x $2988$ pixel image (taken at a high ISO to deliberately introduce noise) and a plot of its grayscale intensity versus the image's spatial dimensions:
+For example, consider the $5312$ x $2988$ pixel image (taken at a high ISO to deliberately introduce noise) of Figure \ref{noisy-image} and a plot of its grayscale intensity versus the image's spatial dimensions in Figure \ref{noisy-image-mesh}.
+It is clear from the mesh that there is much high-frequency noise, which can be removed with a low-pass filter.
 \begin{figure}[H]
@@
 	\includegraphics[width=0.45\textwidth]{sobel_pcb_mesh.jpg}
-	\caption{Gradient of \ref{sobel-pcb} visualized as a three-dimensional mesh}
-    \label{sobel-pcb}
+	\caption{Gradient of Figure \ref{sobel-pcb} visualized as a three-dimensional mesh}
+    \label{sobel-pcb-mesh}
 \end{figure}
@@
 This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.
+\begin{algorithm}[h]
+	\smaller
+	\label{non-maximum-suppression-selective-thresholding-algorithm}
+	\caption{Non-maximum suppression and selective thresholding: pseudocode for a serial, CPU implementation}
+
+	\KwIn{Gradient magnitude matrix $\boldsymbol{G}_{i, j}$ and gradient angle matrix $\boldsymbol{\theta}_{i, j}$ for an image $\boldsymbol{A}_{i, j}$, $\forall i \in [0, M - 1], \forall j \in [0, N - 1]$, $\alpha$ (representing the size of the box around which to apply selective thresholding), $t_h$ (the high threshold), and $t_l$ (the low threshold)}
+	\KwOut{Edge matrix $\boldsymbol{E}$ of dimensions $M$ x $N$ that represent the edges of $\boldsymbol{A}$; any entry $\boldsymbol{E}_{i, j}$ is $255$ if an edge is present, and $0$ if an edge is not present}
+
+	\nl Initialize $\boldsymbol{E}$ to 0 $\forall i \in [0, M - 1], \forall j \in [0, N - 1]$\;
+	\nl \ForEach {$i \in [\alpha + 1, M - \alpha - 2]$} {
+		\nl \ForEach {$j \in [\alpha + 1, N - \alpha - 2]$} {
+			\nl HorizontalGradient $\gets -\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi$\;
+			\nl HorizontalMax $\gets$ HorizontalGradient $\wedge G_{i, j} > G_{i, j \pm 1}$\;
+			\nl PosDiagonalGradient $\gets \frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}$\;
+			\nl PosDiagonalMax $\gets$ PosDiagonalGradient $\wedge G_{i, j} > G_{i \pm 1, \pm j}$\;
+			\nl VerticalGradient $\gets \frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}$\;
+			\nl VerticalMax $\gets$ VerticalGradient $\wedge G_{i, j} > G_{i \pm 1, j}$\;
+			\nl NegDiagonalGradient $\gets \frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}$\;
+			\nl NegDiagonalMax $\gets$ NegDiagonalGradient $\wedge G_{i, j} > G_{i \pm 1, \mp j}$\;
+
+			\nl \If {$\boldsymbol{G}_{i, j} > t_h \wedge ($\textnormal{HorizontalMax} $\vee$ \textnormal{PosDiagonalMax} $\vee$ \textnormal{VerticalMax} $\vee$ \textnormal{NegDiagonalMax}$)$} {
+				\For {$i' \in [-\alpha, \alpha]$} {
+					\For {$j' \in [-\alpha, \alpha]$} {
+						\nl \If{$\boldsymbol{E}_{i + i', j + j'} > t_l$} {
+							\nl $\boldsymbol{E}_{i + i', j + j'} \gets 255$\;
+						}
+					}
+				}
+			}
+		}
+	}
+	\
+\end{algorithm}
 
 \section{Motion Detection Algorithm}
-Placeholder text
+Our motion detection algorithm compares the edges from two frames temporally separated by a small interval of time and attempts to estimate regions of the frame that exhibit movement or motion.
 
@@
 
-\appendices
-\section{Proof of the First Zonklar Equation}
-Some text for the appendix.
-
 \section*{Acknowledgment}

### 2015-11-30 documents/paper/paper.tex
Introduction for motion detection algorithm

@@
 		\nl \ForEach {$j \in [\alpha + 1, N - \alpha - 2]$} {
-			\nl HorizontalGradient $\gets -\frac{\pi}{6} < \theta < \frac{\pi}{6} \vee -\pi < \theta < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \theta < \pi$\;
-			\nl HorizontalMax $\gets$ HorizontalGradient $\wedge G_{i, j} > G_{i, j \pm 1}$\;
-			\nl PosDiagonalGradient $\gets \frac{\pi}{6} < \theta < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \theta < -\frac{2 \pi}{3}$\;
-			\nl PosDiagonalMax $\gets$ PosDiagonalGradient $\wedge G_{i, j} > G_{i \pm 1, \pm j}$\;
-			\nl VerticalGradient $\gets \frac{\pi}{3} < \theta < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \theta < -\frac{\pi}{3}$\;
-			\nl VerticalMax $\gets$ VerticalGradient $\wedge G_{i, j} > G_{i \pm 1, j}$\;
-			\nl NegDiagonalGradient $\gets \frac{2 \pi}{3} < \theta < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \theta < -\frac{\pi}{6}$\;
-			\nl NegDiagonalMax $\gets$ NegDiagonalGradient $\wedge G_{i, j} > G_{i \pm 1, \mp j}$\;
+			\nl HorizontalGradient $\gets -\frac{\pi}{6} < \boldsymbol{\theta}_{i, j} < \frac{\pi}{6} \vee -\pi < \boldsymbol{\theta}_{i, j} < -\frac{5 \pi}{6} \vee \frac{5\pi}{6} < \boldsymbol{\theta}_{i, j} < \pi$\;
+			\nl HorizontalMax $\gets$ HorizontalGradient $\wedge \boldsymbol{G}_{i, j} > \boldsymbol{G}_{i, j \pm 1}$\;
+			\nl PosDiagonalGradient $\gets \frac{\pi}{6} < \boldsymbol{\theta}_{i, j} < \frac{\pi}{3} \vee -\frac{5 \pi}{6} < \boldsymbol{\theta}_{i, j} < -\frac{2 \pi}{3}$\;
+			\nl PosDiagonalMax $\gets$ PosDiagonalGradient $\wedge \boldsymbol{G}_{i, j} > \boldsymbol{G}_{i \pm 1, \pm j}$\;
+			\nl VerticalGradient $\gets \frac{\pi}{3} < \boldsymbol{\theta}_{i, j} < \frac{2 \pi}{3} \vee -\frac{2 \pi}{3} < \boldsymbol{\theta}_{i, j} < -\frac{\pi}{3}$\;
+			\nl VerticalMax $\gets$ VerticalGradient $\wedge \boldsymbol{G}_{i, j} > \boldsymbol{G}_{i \pm 1, j}$\;
+			\nl NegDiagonalGradient $\gets \frac{2 \pi}{3} < \boldsymbol{\theta}_{i, j} < \frac{5 \pi}{6} \vee -\frac{\pi}{3} < \boldsymbol{\theta}_{i, j} < -\frac{\pi}{6}$\;
+			\nl NegDiagonalMax $\gets$ NegDiagonalGradient $\wedge \boldsymbol{G}_{i, j} > \boldsymbol{G}_{i \pm 1, \mp j}$\;
 
 			\nl \If {$\boldsymbol{G}_{i, j} > t_h \wedge ($\textnormal{HorizontalMax} $\vee$ \textnormal{PosDiagonalMax} $\vee$ \textnormal{VerticalMax} $\vee$ \textnormal{NegDiagonalMax}$)$} {
-				\For {$i' \in [-\alpha, \alpha]$} {
-					\For {$j' \in [-\alpha, \alpha]$} {
-						\nl \If{$\boldsymbol{E}_{i + i', j + j'} > t_l$} {
+				\ForEach {$i' \in [-\alpha, \alpha]$} {
+					\ForEach {$j' \in [-\alpha, \alpha]$} {
+						\nl \If{$\boldsymbol{G}_{i + i', j + j'} > t_l$} {
 							\nl $\boldsymbol{E}_{i + i', j + j'} \gets 255$\;
@@
 	}
+	\nl \Return $\boldsymbol{E}$\;
 	\
@@
 
+\subsection{GPU parallelization}
+Placeholder text
+
 \section{Motion Detection Algorithm}
 Our motion detection algorithm compares the edges from two frames temporally separated by a small interval of time and attempts to estimate regions of the frame that exhibit movement or motion.
+This process is then repeated continuously to provide a real-time visualization of the regions of movement in a continuous video stream.
+The motion detection algorithm we present is composed of the following steps:
+\begin{enumerate}
+	\item Given two frames $\boldsymbol{F}_1$ and $\boldsymbol{F}_2$ separated by a time interval of $\Delta t$, compute its edges $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$.
+	\item Create a strict binary difference matrix $\boldsymbol{D}_0$ between $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ whose value is high at locations at which an edge is present in frame but not the other, and low at locations at which both frames have an edge or no edge.
+	\item For each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta$ x $\beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$.
+This step is designed to reduce false positives of motion detection arising from stray movement of the frame (e.g.
+camera shake).
+\end{enumerate}
+More precisely, the matrix $\boldsymbol{D}$ referenced above is defined as
+\[
+\boldsymbol{D} = |\boldsymbol{E}_1 - \boldsymbol{E}_2| =
+\begin{cases}
+	255 & \boldsymbol{E}_{1_{i, j}} \neq \boldsymbol{E}_{2_{i, j}} \\
+	0 & \boldsymbol{E}_{1_{i, j}} = \boldsymbol{E}_{2_{i, j}}
+\end{cases}
+\]
+The computation of $\boldsymbol{D'}$ is determined by Algorithm \ref{difference-matrix-algorithm}.
+\begin{algorithm}[h]
+	\smaller
+	\label{difference-matrix-algorithm}
+	\caption{Determining the difference matrix between two frames, given their edge data: pseudocode for a serial, CPU implementation}
+
+	\KwIn{Edge matrices $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$, both of dimensions $M$ x $N$, and movement tolerance threshold $\beta$}
+	\KwOut{Thresholded difference matrix $\boldsymbol{D}$; any entry $\boldsymbol{D}_{i, j}$ is $255$ if there is a difference, $0$ otherwise}
+
+	\nl Initialize $\boldsymbol{D}_{i, j}$ to 0 $\forall i \in [0, M - 1], \forall j \in [0, N - 1]$\;
+	\nl \ForEach {$i \in [0, M - 1]$} {
+		\nl \ForEach {$j \in [0, N - 1]$} {
+			\nl \If {$\boldsymbol{E}_{1_{i, j}} \neq \boldsymbol{E}_{2_{i, j}}$} {
+				\nl $\boldsymbol{D}_{i, j} \gets 255$\;
+				\ForEach {$i' \in [-\beta, \beta]$} {
+					\ForEach {$j' \in [-\beta, \beta]$} {
+						\nl \If{$(i + i', j + j')$ \textnormal{is within the bounds of} $E_1$ \textnormal{and} $\boldsymbol{E}_{1_{i + i', j + j'}} = \boldsymbol{E}_{2_{i + i', j + j'}}$} {
+							\nl $\boldsymbol{D}_{i, j} \gets 0$\;
+						}
+					}
+				}
+			}
+		}
+	}
+	\nl \Return $\boldsymbol{D}$\;
+	\
+\end{algorithm}
 

### 2015-12-01 documents/paper/paper.tex
Theory details for motion detection algorithm + images

@@
 This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.
+\begin{figure}[]
+	\centering
+	\includegraphics[width=0.45\textwidth]{pcb_edges.jpg}
+	\caption{Output of edge detection algorithm after non-maximum suppression and selective thresholding on the image of Figure \ref{pcb-original}, with parameters $t_l = 70$ and $t_h = 80$}
+    \label{pcb-edges}
+\end{figure}
 \begin{algorithm}[h]
@@
 	\item Create a strict binary difference matrix $\boldsymbol{D}_0$ between $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ whose value is high at locations at which an edge is present in frame but not the other, and low at locations at which both frames have an edge or no edge.
-	\item For each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta$ x $\beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$.
-This step is designed to reduce false positives of motion detection arising from stray movement of the frame (e.g.
-camera shake).
+	\item Apply a parameter-controlled movement tolerance thresholding operation to the matrix $\boldsymbol{D}_0$ to obtain matrix $\boldsymbol{D}$, in order to suppress potential false positives of detected motion from incidental camera movement.
 \end{enumerate}
-More precisely, the matrix $\boldsymbol{D}$ referenced above is defined as
+\par $\boldsymbol{F}_1$ and $\boldsymbol{F}_2$ are obtained by sampling frames from a continuous video stream, a task aided by the OpenCV C library.
+Thus, the time interval $\Delta t$ will be small enough for a real-time algorithm.
+The edges $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ are computed with the edge detection algorithm of Section II.
+Below, we explore, in greater depth, the procedures for building a thresholded diffrence matrix and finally estimating the motion area by analyzing a spatial difference density map.
+
+\subsection{Thresholded Frame-by-Frame Difference Map}
+The preliminary requirement in determining areas of motion from two frames is a systematic method of determining and quantifying the differences between two frames separated by a small $\Delta t$.
+Intuitively, areas of motion within the time interval $\Delta t$ should exist at locations at which the difference between the two frames is large.
+This process is then continuously repeated for every pair of input frames in order to create a real-time motion detector.
+In our algorithm, we quantify differences on a binary basis using edge data from our edge detection algorithm, then determine motion locations from a thresholded density map of the difference matrix $\boldsymbol{D}$.
+\par More formally, we begin with a matrix $\boldsymbol{D}_0$ created from a simple comparison of $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$.
 \[
-\boldsymbol{D} = |\boldsymbol{E}_1 - \boldsymbol{E}_2| =
+\boldsymbol{D}_0 = |\boldsymbol{E}_1 - \boldsymbol{E}_2| =
 \begin{cases}
@@
 \]
-The computation of $\boldsymbol{D'}$ is determined by Algorithm \ref{difference-matrix-algorithm}.
-\begin{algorithm}[h]
+Then, for each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta$ x $\beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$.
+This step is designed to reduce false positives of motion detection arising from stray movement of the frame (e.g.
+camera shake).
+The computation of $\boldsymbol{D}$ is demonstrated in Algorithm \ref{difference-matrix-algorithm}.
+The output matrix is then used as input to building a spatial difference density map, from which we estimate motion locations.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_1.jpg}
+	\caption{Example of difference matrix computation: original initial frame $\boldsymbol{F}_1$ of dimensions $5312$ x $2988$}
+    \label{desk-1-original}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_2.jpg}
+	\caption{Example of difference matrix computation: original initial frame $\boldsymbol{F}_2$ (the item displaced between the two frames is the highlighter at the center)}
+    \label{desk-2-original}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_edges_1.jpg}
+	\caption{Example of difference matrix computation: edges of the initial frame $\boldsymbol{E}_1$ computed with parameters $k = 5$, $\sigma = 5$, $t_l = 15$, $t_h = 25$}
+    \label{desk-1-edges}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_edges_2.jpg}
+	\caption{Example of difference matrix computation: edges of the initial frame $\boldsymbol{E}_2$ computed with parameters $k = 5$, $\sigma = 5$, $t_l = 15$, $t_h = 25$}
+    \label{desk-2-edges}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_difference.jpg}
+	\caption{Example of difference matrix computation: difference matrix $\boldsymbol{D}$ with parameter $\beta = 12$}
+    \label{desk-difference}
+\end{figure}
+\begin{algorithm}[]
 	\smaller
@@
 
-\subsection{Thresholded Frame-by-Frame Difference Map}
-Placeholder text
-
 \subsection{Motion Area Estimation via a Spatial Difference Density Map}

### 2015-12-01 documents/paper/paper.tex
Spatial difference map details in paper

@@
 
-\subsection{GPU parallelization}
-Placeholder text
+\subsection{GPU Parallelization and CUDA Implementation}
+We begin with a parallelized implementation of separable convolution, since the same operation is applied for both the initial Gaussian low pass filter
 
@@
 						\nl \If{$(i + i', j + j')$ \textnormal{is within the bounds of} $E_1$ \textnormal{and} $\boldsymbol{E}_{1_{i + i', j + j'}} = \boldsymbol{E}_{2_{i + i', j + j'}}$} {
-							\nl $\boldsymbol{D}_{i, j} \gets 0$\;
+							\nl $\boldsymbol{D}_{i + i', j + j'} \gets 0$\;
 						}
@@
 \subsection{Motion Area Estimation via a Spatial Difference Density Map}
-Placeholder text
+The difference matrix $\boldsymbol{D}$ determined with Algorithm \ref{difference-matrix-algorithm} is effectively a maximally precise estimate of motion area (e.g., to the resolution of a single pixel, since $\boldsymbol{D}$ is computed on a pixel-by-pixel basis).
+However, in practice, it is more useful to consider general regions of the image around which motion exists.
+For this reason, we build a density map of the matrix $\boldsymbol{D}$ using equal-sized rectangular sections of the input image that approximates the density of the difference per unit area (units pixels$^2$).
+Thus, the original $M$ x $N$ difference matrix $\boldsymbol{D}$ is reduced to a matrix $\boldsymbol{D}'$ of size $m$ x $n$ such that
+$$M = k_1 m$$
+$$N = k_2 n$$
+where $k_1, k_2 \in \mathbb{Z}$.
+Intuitively, choosing a higher $k_1$ and $k_2$ will result in a higher resolution over which the density is evaluated.
+Selection of $k_1$ and $k_2$ should depend on how large of a region over which motion should be estimated.
+The upper limits of $k_1$ and $k_2$ are $M$ and $N$, respectively, at which point $\boldsymbol{D}'$ = $\boldsymbol{D}$.
+Each index $(m, n)$ in $\boldsymbol{D}'$ corresponds to the upper-left corner of a rectangular section of $\boldsymbol{A}$ whose width is $\frac{M}{m}$ and height is $\frac{N}{n}$.
+\par Regions of motion are the rectangular sections whose density exceeds a user-defined threshold $\Gamma$, denoting the number of difference pixels over the total number of pixels considered in the rectangular section.
+For each such region $\boldsymbol{R}$ in $\boldsymbol{A}$ of size $\frac{M}{m}$ x $\frac{N}{n}$,
+$$\textnormal{DifferenceDensity}(\boldsymbol{R}) = \frac{\textnormal{number of pixel differences in } \boldsymbol{R}}{\textnormal{number of pixels in } \boldsymbol{R}}$$
+\[
+\boldsymbol{D}' =
+\begin{cases}
+1 & \textnormal{DifferenceDensity}(\boldsymbol{R}) > \Gamma\\
+0 & \textnormal{otherwise}
+\end{cases}
+\]
+\begin{algorithm}[]
+	\smaller
+	\label{difference-density-matrix-algorithm}
+	\caption{Estimating regions of motion in an image with a difference density matrix: pseudocode for a serial, CPU implementation}
+
+	\KwIn{Thresholded difference matrix $\boldsymbol{D}$, motion threshold $\Gamma$, width of the original image $M$, height of the original image $N$, number of horizontal divisions $m$, and number of vertical divisions $n$ ($m$ and $n$ are the dimensions of $\boldsymbol{D}'$)}
+	\KwOut{The difference density matrix $\boldsymbol{D}'$ of dimensions $m$ x $n$ whose value is 1 if there is motion in that region, and 0 otherwise}
+
+	\nl Initialize $\boldsymbol{D}'$ to 0 $\forall i \in [0, m - 1], \forall j \in [0, n - 1]$\;
+	\nl \ForEach {$i \in [0, m - 1]$} {
+		\nl \ForEach {$j \in [0, n - 1]$} {
+			\nl $x \gets 0$\tcp*{Number of differences in this region}\
+			\nl \ForEach{$i' \in [0, \frac{M}{n}]$} {
+				\nl \ForEach{$j' \in [0, \frac{N}{n}]$} {
+					\nl \If {$\boldsymbol{D}_{i + i', j + j'} \neq 0$} {
+						\nl $x \gets x + 1$\;
+					}
+				}
+			}
+			\nl \If{$\frac{mn}{MN}x > \Gamma$} {
+				\nl $\boldsymbol{D}'_{i, j} \gets 1$\;
+			}
+		}
+	}
+	\nl \Return $\boldsymbol{D}'$\;
+	\
+\end{algorithm}
 
@@
 \section*{Acknowledgment}
-\begin{thebibliography}{1}
+Placeholder text
 
-\bibitem{IEEEhowto:kopka}
-H.~Kopka and P.~W. Daly, \emph{A Guide to \LaTeX}, 3rd~ed.\hskip 1em plus
-  0.5em minus 0.4em\relax Harlow, England: Addison-Wesley, 1999.
-
-\end{thebibliography}
 \end{document}

### 2015-12-01 documents/paper/paper.tex
Add FFT images

@@
 \end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{fft_noisy_image.jpg}
+	\caption{FFT magnitude of unfiltered noisy image at each pixel (log scale)}
+    \label{noisy-image-fft}
+\end{figure}
 To low-pass filter our image, we apply a discrete Gaussian filter. Generally, a Gaussian blur kernel of size $2n + 1$ x $2n + 1$ ($\forall n \in \mathbb{Z}^+$, and with parameter $\sigma$) is given by
@@
     \label{filtered-noisy-image-mesh}
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{fft_filtered_noisy_image.jpg}
+	\caption{FFT magnitude of Gaussian-filtered noisy image at each pixel (log scale)}
+    \label{filtered-noisy-image-fft}
 \end{figure}
@@
 
-	\KwIn{Thresholded difference matrix $\boldsymbol{D}$, motion threshold $\Gamma$, width of the original image $M$, height of the original image $N$, number of horizontal divisions $m$, and number of vertical divisions $n$ ($m$ and $n$ are the dimensions of $\boldsymbol{D}'$)}
+	\KwIn{Thresholded difference matrix $\boldsymbol{D}$, motion threshold $\Gamma$, width of the original image $M$, height of the original image $N$, number of horizontal divisions $m$, and number of vertical divisions $n$}
 	\KwOut{The difference density matrix $\boldsymbol{D}'$ of dimensions $m$ x $n$ whose value is 1 if there is motion in that region, and 0 otherwise}
@@
 		\nl \ForEach {$j \in [0, n - 1]$} {
-			\nl $x \gets 0$\tcp*{Number of differences in this region}\
+			\nl $x \gets 0$\tcp*{Number of differences in this region}
 			\nl \ForEach{$i' \in [0, \frac{M}{n}]$} {
@@
 
-\section*{Acknowledgment}
+\section*{Acknowledgments}
 Placeholder text

### 2015-12-02 cuda/edge-detect/edge_detect.cu
Continued modularization of edge_detect

@@ -33,9 +33,6 @@ __global__ void thresholding_and_suppression(int *dev_output, double *dev_magnit
     if (r > 1 && c > 1 && r < input_height - 1 && c < input_width - 1) {
         double magnitude = dev_magnitude[i];
         if (magnitude > high_threshold) {
-        	// Immediately accept any pixel above the high threshold
-        	dev_output[i] = 255;
-        	
         	// Non-maximum suppression: determine magnitudes in the surrounding pixel and the gradient direction of the current pixel
         	double magnitude_above = dev_magnitude[(r - 1) * input_width + c];
 			double magnitude_below = dev_magnitude[(r + 1) * input_width + c];
@@ -60,6 +57,7 @@ __global__ void thresholding_and_suppression(int *dev_output, double *dev_magnit
 			// Consider a surrounding apron around the current pixel to catch potentially disconnected pixel nodes
 			int apron_size = 2;
 			if (is_vertical_max || is_horizontal_max || is_positive_diagonal_max || is_negative_diagonal_max) {
+				dev_output[i] = 255;
 				for (int m = -apron_size; m <= apron_size; m++) {
 					for (int n = -apron_size; n <= apron_size; n++) {
 						if (r + m > 0 && r + m < input_height && c + n > 0 && c + n < input_width) {
@@ -86,9 +84,6 @@ double serial_thresholding_and_suppression(int *dev_output, int input_width, int
 			if (r > 1 && c > 1 && r < input_height - 1 && c < input_width - 1) {
 				double magnitude = sqrt(pow((double)g_x[i], 2) + pow((double)g_y[i], 2));
 				if (magnitude > high_threshold) {
-					// Immediately accept any pixel above the high threshold
-					dev_output[i] = 255;
-					
 					// Non-maximum suppression: determine magnitudes in the surrounding pixel and the gradient direction of the current pixel
 					double magnitude_above = sqrt(pow((double)g_x[(r - 1) * input_width + c], 2) + pow((double)g_y[(r - 1) * input_width + c], 2));
 					double magnitude_below = sqrt(pow((double)g_x[(r + 1) * input_width + c], 2) + pow((double)g_y[(r + 1) * input_width + c], 2));
@@ -113,6 +108,7 @@ double serial_thresholding_and_suppression(int *dev_output, int input_width, int
[…]

### 2015-12-02 cuda/edge-detect/edge_detect.h

@@ -1,6 +1,6 @@
 #ifndef EDGE_DETECT_H
 #define EDGE_DETECT_H
 
-void edge_detect(int *edges_out, int *image, int image_width, int image_height, int high_threshold, int low_threshold);
+void edge_detect(int *edges_out, int *gx_out, int *gy_out, int *image, int image_width, int image_height, int high_threshold, int low_threshold);
 
 #endif

### 2015-12-02 cuda/motion-tracking/motion_tracking.cu
Fixed host memory leak

@@ -22,7 +22,7 @@ void spatial_difference_density_map(double *density_map, int *difference, int wi
 	int vertical_block_size = height/vertical_divisions;
 	int block_size = horizontal_block_size * vertical_block_size;
 	
-	const int scaling_factor = 1000; // Used to linearly scale density map to units millipixels/pixels^2 (if that makes any sense?)
+	const int scaling_factor = 1000;  // Used to linearly scale density map to units millipixels/pixels^2 (if that makes any sense?)
 	
 	for (int block_x_index = 0; block_x_index < horizontal_divisions; block_x_index++) {
 		for (int block_y_index = 0; block_y_index < vertical_divisions; block_y_index++) {
@@ -112,7 +112,7 @@ double serial_difference_filter(int *difference, int *edges_1, int *edges_2, int
     return time_spent;
 }
 
-void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2, int width, int height, int movement_threshold, int motion_threshold) {
+void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2, int width, int height, int movement_threshold, int motion_threshold, int horizontal_divisions, int vertical_divisions) {
 	// Note: movement_threshold refers to the pixel apron around which the difference filter attempts to look for differences.
 	// Higher movement_threshold == more leniency in how much camera shake is tolerated
 	// motion_threshold refers to the minimum spatial pixel difference density required for a particular segment of the difference to be registered as motion.
@@ -157,28 +157,57 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
     checkCudaErrors(cudaFree(dev_edges_2));
     
     // Determine spatial density map
-    int horizontal_divisions = 12, vertical_divisions = 10;
 	double *density_map = (double *)malloc(horizontal_divisions * vertical_divisions * sizeof(double));
 	spatial_difference_density_map(density_map, difference, width, height, horizontal_divisions, vertical_divisions);
 	
 	// Estimate motion area
[…]

### 2015-12-02 documents/paper/paper.tex
Add details on CUDA implementation of separable convolution

@@
 \usepackage{tabularx}
+\usepackage{listings}
 \usepackage{float}
+\usepackage{color}
 \usepackage{bm}
+
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{gray}{rgb}{0.5,0.5,0.5}
+\definecolor{mauve}{rgb}{0.58,0,0.82}
+
+\lstset{frame=tb,
+	language=C,
+	aboveskip=3mm,
+	belowskip=3mm,
+	showstringspaces=false,
+	columns=flexible,
+	basicstyle={\small\ttfamily},
+	numbers=none,
+	numberstyle=\tiny\color{gray},
+	keywordstyle=\color{blue},
+	commentstyle=\color{dkgreen},
+	stringstyle=\color{mauve},
+	breaklines=true,
+	breakatwhitespace=true,
+	tabsize=3
+}
 
@@
 that $h$ can be represented as the product of two one-dimensional matrices $h_1$ and $h_2$.
-This effectively reduces our two-dimensional convolution of \eqref{2d-naive-convolution} to two separate instances of one-dimensional convolution:
-$$y[m, n] = h_1[n] * (h_2[m] * x[m, n])$$
+This effectively reduces our two-dimensional convolution of Equation \eqref{2d-naive-convolution} to two separate instances of one-dimensional convolution:
+\begin{equation}
+	\label{separable-convolution-definition}
+	y[m, n] = h_1[n] * (h_2[m] * x[m, n])
+\end{equation}
 which is implemented as
-$$y[m, n] = \sum_{i = -\infty}^{\infty}h_1[i] \sum_{j = -\infty}^{\infty}h_2[j] x[m - i, n - j]$$
+\begin{equation}
+	\label{separable-convolution-formula}
+	y[m, n] = \sum_{i = -\infty}^{\infty}h_1[i] \sum_{j = -\infty}^{\infty}h_2[j] x[m - i, n - j]
+\end{equation}
 For the same $M$ x $N$ input and square kernel of size $k$, a separable implementation reduces the computational complexity to $O(MNk)$.
@@
 \subsection{GPU Parallelization and CUDA Implementation}
-We begin with a parallelized implementation of separable convolution, since the same operation is applied for both the initial Gaussian low pass filter
+We begin with a parallelized implementation of separable convolution, since the same convolution operation is applied for both the initial Gaussian low pass filter and the edge detecting Sobel filter.
+Our parallel implementation involves first independently computing the convolution with the horizontal filter ($h_2$ in Equation \eqref{separable-convolution-definition}), then convolving that result with the vertical filter $h_1$.
+In computing the output $y[m, n]$ at every index $(i, j)$, we launch a kernel with 16 threads per block, where the total number of blocks along one dimension is equal to the size of that dimension divided by 16.
+For our 5312 x 2988 test images, this corresponds to 332 blocks in the horizontal direction and 186 blocks in the vertical direction, each with 16 computational threads.
+\begin{lstlisting}[
+	caption=Grid configuration for the separable convolution device kernel,
+	label=grid-configuration
+]
+#define TX 16
+#define TY 16
+
+dim3 block_size(TX, TY);
+int bx_horizontal = horizontal_convolution_width/block_size.x;
+int by_horizontal = horizontal_convolution_height/block_size.y;
+dim3 grid_size_horizontal = dim3(bx_horizontal, by_horizontal);
+int bx_vertical = vertical_convolution_width/block_size.x;
+int by_vertical = vertical_convolution_height/block_size.y;
+dim3 grid_size_vertical = dim3(bx_vertical, by_vertical);
+
+horizontal_convolve<<<grid_size_horizontal, block_size>>>(...);
+vertical_convolve<<<grid_size_vertical, block_size>>>(...);
+\end{lstlisting}
+\par The convolution on the device executes only one serial loop across the dimensions of the filter exactly as it is defined in Equation \eqref{separable-convolution-formula}.
+This sum is then stored in the output array\footnote{In our implementation, we represent a two-dimensional image as a one-dimensional array by directly concatenating the rows of the matrix into a single array.
+Thus, for a matrix of size $M$ x $N$, the index $(i, j)$ would correspond to index $Ni + j$ in our one-dimensional array.} according to the block index, block size, and thread index.
+The implementation is modularized so that any separable filter can be applied to an input matrix.
+\begin{lstlisting}[
+	caption=Implementation of horizontal and vertical separated convolution,
+	label=separable-convolution-implementation
+]
+__global__ void horizontal_convolve(int *d_out, int *x, int *h, int x_width, int x_height, int h_width, int h_height) {
+    const int r = blockIdx.y * blockDim.y + threadIdx.y;
+    const int c = blockIdx.x * blockDim.x + threadIdx.x;
+    const int i = r * (x_width + h_width - 1) + c;
+
+    int sum = 0;
+    for (int j = 0; j < h_width; j++) {
+        int p = x_width*r + c - j;
+        if (c - j >= 0 && c - j < x_width) {
+            sum += h[j] * x[p];
+        }
+    }
+    d_out[i] = sum;
+    __syncthreads();
+}
+
+__global__ void vertical_convolve(int *d_out, int *x, int *h, int x_width, int x_height, int h_width, int h_height, double constant_scalar) {
+	const int r = blockIdx.y * blockDim.y + threadIdx.y;
+	const int c = blockIdx.x * blockDim.x + threadIdx.x;
+	const int i = r * x_width + c;
+
+    int sum = 0;
+    for (int j = 0; j < h_height; j++) {
+        int p = x_width*(r - j) + c;
+        if (r - j >= 0 && r - j < x_height) {
+            sum += h[j] * x[p];
+        }
+    }
+    d_out[i] = (int)(constant_scalar * (double)sum);
+    __syncthreads();
+}
+\end{lstlisting}
 
@@
 \]
+\begin{figure}[]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_difference_density.jpg}
+	\caption{Spatial difference density matrix $\boldsymbol{D}'$ (visualized in three dimensions) of Figure \ref{desk-difference}, as determined by Algorithm \ref{difference-density-matrix-algorithm} with parameters $m = 10$, $n = 6$, and $\Gamma = 0.01$}
+    \label{desk-difference-density}
+\end{figure}
+\begin{figure}[]
+	\centering
+	\includegraphics[width=0.45\textwidth]{desk_motion_area.jpg}
+	\caption{Final estimate of motion area of Figures \ref{desk-1-original} and \ref{desk-2-original} based on the density map of Figure \ref{desk-difference-density}}
+    \label{desk-difference-density}
+\end{figure}
 \begin{algorithm}[]
@@
 
+\section*{Team}
+\begin{itemize}
+	\item \textbf{Emilio Del Vechhio}, Electrical and Computer Engineering `18
+	\item \textbf{Kevin Lin}, Electrical and Computer Engineering `18
+	\item \textbf{Senthil Natarjan}, Electrical and Computer Engineering `17
+\end{itemize}
+
 \section*{Acknowledgments}
-Placeholder text
+\begin{itemize}
+	\item \textbf{CJ Barberan}, Ph.D.
+student, Rice University Electrical and Computer Engineering---\textit{for mentoring the entire team in the conception, development, and extensive debugging of our project}.
+\end{itemize}
 

### 2015-12-03 cuda/benchmark/benchmark.cu
Benchmark script

@@ -1,0 +1,61 @@
+#include <stdio.h>
+#include "../separable-convolution/separable_convolution.h"
+
+
+void compare_naive_separable_convolution(int input_width, int input_height) {
+    // Benchmark serial (CPU) performance of naive convolution versus separated convolution on a fixed filter and randomly-valued input
+    int *out = (int *)malloc((input_width + 2) * (input_height + 2) * sizeof(int));
+    int h[9] = {1, 2, 1, 2, 4, 2, 1, 2, 1};
+    int h_horizontal[3] = {1, 2, 1};
+    int h_vertical[3] = {1, 2, 1};
+    int *x = (int *)malloc(input_width * input_height * sizeof(int));
+    for (int i = 0; i < input_width * input_height; i++) {
+        x[i] = rand();
+    }
+    struct timeval tv1, tv2;
+    gettimeofday(&tv1, NULL);
+    serial_naive_convolve(out, x, h, input_width, input_height, 3, 3);
+    gettimeofday(&tv2, NULL);
+    double naive_computation_time = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
+
+    gettimeofday(&tv1, NULL);
+    serial_separable_convolve(out, x, h_horizontal, h_vertical, input_width, input_height, 3, 3, 1.0)
+    gettimeofday(&tv2, NULL);
+    double separable_computation_time = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
+
+    printf("Serial naive convolution execution time: %f", naive_computation_time);
+    printf("Serial separable convolution execution time: %f", separable_computation_time);
+}
+
+void compare_separable_convolution_speedup(int input_width, int input_height) {
+    // Benchmark parallel speedup of Gaussian filter - serial CPU separated convolution vs. parallel GPU separated convolution
+    int *out = (int *)malloc((input_width + 2) * (input_height + 2) * sizeof(int));
+    int h[9] = {1, 2, 1, 2, 4, 2, 1, 2, 1};
+    int h_horizontal[3] = {1, 2, 1};
+    int h_vertical[3] = {1, 2, 1};
+    int *x = (int *)malloc(input_width * input_height * sizeof(int));
+    for (int i = 0; i < input_width * input_height; i++) {
+        x[i] = rand();
+    }
+    struct timeval tv1, tv2;
[…]

### 2015-12-03 cuda/separable-convolution/separable_convolution.cu
Add serial convolution methods to header file

@@ -77,7 +77,7 @@ double serial_separable_convolve(int *out, int *x, int *horizontal_filter, int *
     return time_spent;
 }
 
-double serial_convolve(int *out, int *x, int *h, int x_width, int x_height, int h_width, int h_height) {
+double serial_naive_convolve(int *out, int *x, int *h, int x_width, int x_height, int h_width, int h_height) {
     struct timeval tv1, tv2;
     gettimeofday(&tv1, NULL);
     for (int m = 0; m < x_height + h_height - 1; m++) {

### 2015-12-03 cuda/separable-convolution/separable_convolution.h

@@ -3,4 +3,8 @@
 
 void separable_convolve(int *output, int *x, int x_width, int x_height, int *horizontal_filter, int *vertical_filter, int kernel_size, double constant_scalar);
 
+double serial_separable_convolve(int *out, int *x, int *horizontal_filter, int *vertical_filter, int x_width, int x_height, int horizontal_filter_width, int vertical_filter_height, double constant_scalar);
+
+double serial_naive_convolve(int *out, int *x, int *h, int x_width, int x_height, int h_width, int h_height);
+
 #endif

### 2015-12-04 cuda/separable-convolution/separable_convolution.cu
Removed timing print statements from separable_convolution

@@ -45,9 +45,7 @@ __global__ void vertical_convolve(int *d_out, int *x, int *h, int x_width, int x
     __syncthreads();
 }
 
-double serial_separable_convolve(int *out, int *x, int *horizontal_filter, int *vertical_filter, int x_width, int x_height, int horizontal_filter_width, int vertical_filter_height, double constant_scalar) {
-    struct timeval tv1, tv2;
-    gettimeofday(&tv1, NULL);
+void serial_separable_convolve(int *out, int *x, int *horizontal_filter, int *vertical_filter, int x_width, int x_height, int horizontal_filter_width, int vertical_filter_height, double constant_scalar) {
     int *horizontal_out = (int *)malloc((x_width + horizontal_filter_width - 1) * x_height * sizeof(int));
     for (int m = 0; m < x_height; m++) {
         for (int n = 0; n < x_width + horizontal_filter_width - 1; n++) {
@@ -71,15 +69,9 @@ double serial_separable_convolve(int *out, int *x, int *horizontal_filter, int *
             out[v_m * (x_width + horizontal_filter_width - 1) + v_n] = (int)(constant_scalar * (double)v_sum);
         }
     }
-    gettimeofday(&tv2, NULL);
-    double time_spent = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
-    printf ("Serial separated convolution execution time: %f seconds\n", time_spent);
-    return time_spent;
 }
 
-double serial_naive_convolve(int *out, int *x, int *h, int x_width, int x_height, int h_width, int h_height) {
-    struct timeval tv1, tv2;
-    gettimeofday(&tv1, NULL);
+void serial_naive_convolve(int *out, int *x, int *h, int x_width, int x_height, int h_width, int h_height) {
     for (int m = 0; m < x_height + h_height - 1; m++) {
         for (int n = 0; n < x_width + h_width - 1; n++) {
             int sum = 0;
@@ -93,10 +85,6 @@ double serial_naive_convolve(int *out, int *x, int *h, int x_width, int x_height
             out[m * (x_width + h_width - 1) + n] = sum;
         }
     }
-    gettimeofday(&tv2, NULL);
-    double time_spent = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
-    printf ("Serial naive convolution execution time: %f seconds\n", time_spent);
[…]

### 2015-12-05 documents/paper/paper.tex
Add details of GPU implementation of edge detection algorithm

@@
 \end{equation}
-For an image of size $M$ x $N$ and kernel of size $k$, a direct implementation would require $O(MNk^2)$ time, which is infeasible for a real-time implementation.
+For an image of size $M \times N$ and kernel of size $k$, a direct implementation would require $O(MNk^2)$ time, which is infeasible for a real-time implementation.
 Instead, it is possible to exploit the separable property of certain convolution matrices $h$, e.g.
@@
 \end{equation}
-For the same $M$ x $N$ input and square kernel of size $k$, a separable implementation reduces the computational complexity to $O(MNk)$.
+For the same $M \times N$ input and square kernel of size $k$, a separable implementation reduces the computational complexity to $O(MNk)$.
 By comparison, an implementation using the FFT would cost $O(MN\log{MN})$.
-In practice, consider the following results\footnote{Results were obtained by measuring the execution time of a MATLAB script on an Intel Core i7 4790K @ 4.7 GHz.} for a convolution of a $3$ x $3$ kernel with an image of size $2500$ x $2500$:
+In practice, consider the separable convolution speed gain evident in the following results\footnote{Results were obtained by measuring the CPU execution time of the convolution algorithms implemented in C on an Intel Core i7 4790K @ 4.7 GHz.} for a convolution of a $3 \times 3$ Gaussian filter kernel with images of varying sizes (visualized in Figure \ref{direct-separated-convolution-graph}).
 \begin{table}[H]
@@
 	\centering
-	\caption{Comparison of Computational Complexity of Different Convolution Implementation Algorithms}
+	\caption{Execution Time of Direct 2D Convolution, $O(MNk^2)$}
 	\label{convolution-complexity-comparison}
 	\begin{tabular}{llll}
-	 & Complexity & Execution time (s) \\
-	 Direct convolution & $O(MNk^2)$ & $3$ \\
-	 Separated convolution & $O(MNk)$ & $3$ \\
-	 FFT & $O(MN\log{MN})$ & $3$
+	 Image dimensions & Pixel count ($\times10^6$) & Execution time (s) \\
+	 $100 \times 100$ & $0.01$ & $0.000489$ \\
+	 $500 \times 500$ & $0.25$ & $0.008784$ \\
+	 $1000 \times 1000$ & $1.00$ & $0.034249$ \\
+	 $1500 \times 1500$ & $2.25$ & $0.076803$ \\
+	 $2000 \times 2000$ & $4.00$ & $0.136222$ \\
+	 $2500 \times 2500$ & $6.25$ & $0.212414$ \\
 	\end{tabular}
 \end{table}
+\begin{table}[H]
+	\small
+	\centering
+	\caption{Execution Time of Separated 2D Convolution, $O(MNk)$}
+	\label{convolution-complexity-comparison}
+	\begin{tabular}{llll}
+	 Image dimensions & Pixel count ($\times10^6$) & Execution time (s) \\
+	 $100 \times 100$ & $0.01$ & $0.000279$ \\
+	 $500 \times 500$ & $0.25$ & $0.004768$ \\
+	 $1000 \times 1000$ & $1.00$ & $0.018969$ \\
+	 $1500 \times 1500$ & $2.25$ & $0.042410$ \\
+	 $2000 \times 2000$ & $4.00$ & $0.075280$ \\
+	 $2500 \times 2500$ & $6.25$ & $0.117134$ \\
+	\end{tabular}
+\end{table}
+\begin{figure}
+	\centering
+	\includegraphics[width=0.45\textwidth]{naive_separable_convolution_execution_time.jpg}
+	\caption{Graphical comparison of execution times of direct and separated 2D convolution as a function of the input pixel count}
+    \label{direct-separated-convolution-graph}
+\end{figure}
 As detailed in later sections, we take advantage of the fast computational complexity of separable convolution in our implementation, as all of the filters we apply are separable into one-dimensional matrices.
@@
 \par
-For example, consider the $5312$ x $2988$ pixel image (taken at a high ISO to deliberately introduce noise) of Figure \ref{noisy-image} and a plot of its grayscale intensity versus the image's spatial dimensions in Figure \ref{noisy-image-mesh}.
+For example, consider the $5312 \times 2988$ pixel image (taken at a high ISO to deliberately introduce noise) of Figure \ref{noisy-image} and a plot of its grayscale intensity versus the image's spatial dimensions in Figure \ref{noisy-image-mesh}.
 It is clear from the mesh that there is much high-frequency noise, which can be removed with a low-pass filter.
@@
 \end{figure}
-To low-pass filter our image, we apply a discrete Gaussian filter. Generally, a Gaussian blur kernel of size $2n + 1$ x $2n + 1$ ($\forall n \in \mathbb{Z}^+$, and with parameter $\sigma$) is given by
+To low-pass filter our image, we apply a discrete Gaussian filter.
+Generally, a Gaussian blur kernel of size $2n + 1 \times 2n + 1$ ($\forall n \in \mathbb{Z}^+$, and with parameter $\sigma$) is given by
 $$h[i, j] = \frac{1}{2\pi \sigma^2}e^{-\frac{(i - k - 1)^2 + (j - k - 1)^2}{2\sigma^2}}$$
@@
 This kernel is implemented separably as
-\[
+\begin{equation}
+	\label{gaussian-filter}
 	h[i, j] = \frac{1}{16}
@@
 	\end{bmatrix}
-\] \par
-Applying a Gaussian filter with parameters $k = 5$ and $\sigma = 5$ to the above image significantly denoises the image without sacrificing edge precision, which can be seen from the spatial intensity plot in Figure \ref{filtered-noisy-image-mesh}.
+\end{equation}
+\par Applying a Gaussian filter with parameters $k = 5$ and $\sigma = 5$ to the above image significantly denoises the image without sacrificing edge precision, which can be seen from the spatial intensity plot in Figure \ref{filtered-noisy-image-mesh}.
 \begin{figure}[H]
@@
 Like our Gaussian kernel, the Sobel operator (in both the $x$ and $y$ directions) is separable:
-\[
+\begin{equation}
+	\label{horizontal-sobel-filter}
 	\boldsymbol{G_x} =
@@
 	\end{bmatrix} * \boldsymbol{A})
-\]
-\[
+\end{equation}
+\begin{equation}
+	\label{vertical-sobel-filter}
 	\boldsymbol{G_y} =
@@
 	\end{bmatrix} * \boldsymbol{A})
-\]
+\end{equation}
 \par Thus, we perform two separable convolutions with the separated Sobel filter on the Gaussian-filtered image to obtain gradients in the horizontal and vertical directions, and determine the magnitude and angle matrices $\boldsymbol{G}$ and $\boldsymbol{\theta}$ from equations \eqref{gradient-magnitude} and \eqref{gradient-angle}.
@@
 	\KwIn{Gradient magnitude matrix $\boldsymbol{G}_{i, j}$ and gradient angle matrix $\boldsymbol{\theta}_{i, j}$ for an image $\boldsymbol{A}_{i, j}$, $\forall i \in [0, M - 1], \forall j \in [0, N - 1]$, $\alpha$ (representing the size of the box around which to apply selective thresholding), $t_h$ (the high threshold), and $t_l$ (the low threshold)}
-	\KwOut{Edge matrix $\boldsymbol{E}$ of dimensions $M$ x $N$ that represent the edges of $\boldsymbol{A}$; any entry $\boldsymbol{E}_{i, j}$ is $255$ if an edge is present, and $0$ if an edge is not present}
+	\KwOut{Edge matrix $\boldsymbol{E}$ of dimensions $M \times N$ that represent the edges of $\boldsymbol{A}$; any entry $\boldsymbol{E}_{i, j}$ is $255$ if an edge is present, and $0$ if an edge is not present}
 
@@
 This sum is then stored in the output array\footnote{In our implementation, we represent a two-dimensional image as a one-dimensional array by directly concatenating the rows of the matrix into a single array.
-Thus, for a matrix of size $M$ x $N$, the index $(i, j)$ would correspond to index $Ni + j$ in our one-dimensional array.} according to the block index, block size, and thread index.
+Thus, for a matrix of size $M \times N$, the index $(i, j)$ would correspond to index $Ni + j$ in our one-dimensional array.} according to the block index, block size, and thread index.
 The implementation is modularized so that any separable filter can be applied to an input matrix.
@@
 \end{lstlisting}
+\par The first stage of our edge detection algorithm uses the separable convolution GPU implementation above to apply a Gaussian low-pass filter in Equation \ref{gaussian-filter} to remove high-frequency noise from the image that might be falsely labeled as an edge.
+Then, the same separable convolution kernel is used to apply the Sobel edge filter of Equations \ref{horizontal-sobel-filter} and \ref{vertical-sobel-filter} to the Gaussian low pass-filtered image in both the horizontal and vertical directions to approximate the horizontal and vertical partial derivatives, respectively.
+In order to reduce redundancy in computation, the final stage of non-maximum suppression and selective thresholding is separated into two tasks, each of which is implemented in parallel:
+\begin{enumerate}
+	\item Calculate the gradient magnitude matrix $\boldsymbol{G}$ and gradient angle matrix $\boldsymbol{\theta}$
+	\item Implement Algorithm \ref{non-maximum-suppression-selective-thresholding-algorithm} with the inputs $\boldsymbol{G}$ and $\boldsymbol{\theta}$ calculated above
+\end{enumerate}
+\begin{lstlisting}[
+	caption=GPU computation of gradient magnitude and angle,
+	label=gradient-magnitude-angle-implementation
+]
+__global__ void gradient_magnitude_and_direction(
+		double *dev_magnitude_output,
+		double *dev_angle_output
+		int input_width,
+		int input_height,
+		int *g_x,
+		int *g_y
+	) {
+	const int r = blockIdx.y * blockDim.y + threadIdx.y;
+	const int c = blockIdx.x * blockDim.x + threadIdx.x;
+	const int i = r * (input_width + 2) + c;
+
+	dev_magnitude_output[i] = sqrt(pow((double)g_x[i], 2) + pow((double)g_y[i], 2));
+	dev_angle_output[i] = atan2((double)g_y[i], (double)g_x[i]);
+}
+\end{lstlisting}
+Following completion of the computation of Listing \ref{gradient-magnitude-angle-implementation}, the result is input to our implementation of Algorithm \ref{non-maximum-suppression-selective-thresholding-algorithm}.
+While each of these two tasks operates in parallel (e.g.
+that each pixel's magnitude/angle or edge property is computed in parallel with the other pixels), the two tasks themselves execute serially.
+This is due to the fact that the non-maximum suppression and selective thresholding procedure relies on the gradient and angle at each pixel to have already been computed when the algorithm begins.
+Nonetheless, despite the fact that such serial separation of tasks reduces redundancies in gradient magnitude/angle computations\footnote{Computational redundancy arises from the fact that the non-maximum suppression and selective thresholding procedure needs access to the magnitude and angle of all the pixels in a box of area $\alpha^2$ centered at the current pixel.
+Calculating these properties directly for each pixel during the algorithm will result in calculation of the same magnitude and angle for the same pixel multiple times.}, further optimization is possible to reduce the amount of time that any thread is idle (which occurs after a particular pixel's gradient magnitude/angle is computed but before its edge property is computed).
+\begin{lstlisting}[
+	caption=GPU non-maximum suppression and selective thresholding,
+	label=non-maximum-suppression-selective-thresholding-implementation
+]
+__global__ void thresholding_and_suppression(
+	int *dev_output,
+	double *dev_magnitude,
+	double *dev_angle,
+	int input_width,
+	int input_height,
+	int *g_x,
+	int *g_y,
+	int high_threshold,
+	int low_threshold
+) {
+	const int r = blockIdx.y * blockDim.y + threadIdx.y;
+	const int c = blockIdx.x * blockDim.x + threadIdx.x;
+    const int i = r * (input_width + 2) + c;
+
+    // First, initialize the current pixel to zero (non-edge)
+    dev_output[i] = 0;
+    // Boundary conditions
+    if (r > 1 && c > 1 && r < input_height - 1 && c < input_width - 1) {
+        double magnitude = dev_magnitude[i];
+        if (magnitude > high_threshold) {
+        	// Non-maximum suppression: determine magnitudes in the surrounding pixel and the gradient direction of the current pixel
+        	double magnitude_above = dev_magnitude[(r - 1) * input_width + c];
+			double magnitude_below = dev_magnitude[(r + 1) * input_width + c];
+			double magnitude_left = dev_magnitude[r * input_width + c - 1];
+			double magnitude_right = dev_magnitude[r * input_width + c + 1];
+			double magnitude_upper_right = dev_magnitude[(r + 1) * input_width + c + 1];
+			double magnitude_upper_left = dev_magnitude[(r + 1) * input_width + c - 1];
+			double magnitude_lower_right = dev_magnitude[(r - 1) * input_width + c + 1];
+			double magnitude_lower_left = dev_magnitude[(r - 1) * input_width + c - 1];
+			double theta = dev_angle[i];
+
+			// Check if the current pixel is a ridge pixel, e.g. maximized in the gradient direction
+			int vertical_check = (M_PI/3.0 < theta && theta < 2.0*M_PI/3.0) || (-2.0*M_PI/3.0 < theta && theta < -M_PI/3.0);
+			int is_vertical_max = vertical_check && magnitude > magnitude_below && magnitude > magnitude_above;
+			int horizontal_check = (-M_PI/6.0 < theta && theta < M_PI/6.0) || (-M_PI < theta && theta < -5.0*M_PI/6.0) || (5*M_PI/6.0 < theta && theta < M_PI);
+			int is_horizontal_max = horizontal_check && magnitude > magnitude_right && magnitude > magnitude_left;
+	        int positive_diagonal_check = (theta > M_PI/6.0 && theta < M_PI/3.0) || (theta < -2.0*M_PI/3.0 && theta > -5.0*M_PI/6.0);
+			int is_positive_diagonal_max = positive_diagonal_check && magnitude > magnitude_upper_right && magnitude > magnitude_lower_left;
+			int negative_diagonal_check = (theta > 2.0*M_PI/3.0 && theta < 5.0*M_PI/6.0) || (theta < -M_PI/6.0 && theta > -M_PI/3.0);
+			int is_negative_diagonal_max = negative_diagonal_check && magnitude > magnitude_lower_right && magnitude > magnitude_upper_left;
+
+			// Consider a surrounding apron around the current pixel to catch potentially disconnected pixel nodes
+			int apron_size = 2;
+			if (is_vertical_max || is_horizontal_max || is_positive_diagonal_max || is_negative_diagonal_max) {
+				dev_output[i] = 255;
+				for (int m = -apron_size; m <= apron_size; m++) {
+					for (int n = -apron_size; n <= apron_size; n++) {
+						if (r + m > 0 && r + m < input_height && c + n > 0 && c + n < input_width) {
+							if (dev_magnitude[(r + m) * input_width + c + n] > low_threshold) {
+								dev_output[(r + m) * input_width + c + n] = 255;
+							}
+						}
+					}
+				}
+			}
+		}
+    }
+}
+\end{lstlisting}
 
@@
 \]
-Then, for each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta$ x $\beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$.
+Then, for each location $(i, j)$ in $\boldsymbol{D}_0$, a thresholded difference matrix $\boldsymbol{D}$ is created by considering a box of constant size $\beta \times \beta$ around that pixel and suppressing the value of the surrounding pixels to 0 if its value matches that at $\boldsymbol{D}_{0_{i, j}}$.
 This step is designed to reduce false positives of motion detection arising from stray movement of the frame (e.g.
@@
 For this reason, we build a density map of the matrix $\boldsymbol{D}$ using equal-sized rectangular sections of the input image that approximates the density of the difference per unit area (units pixels$^2$).
-Thus, the original $M$ x $N$ difference matrix $\boldsymbol{D}$ is reduced to a matrix $\boldsymbol{D}'$ of size $m$ x $n$ such that
+Thus, the original $M \times N$ difference matrix $\boldsymbol{D}$ is reduced to a matrix $\boldsymbol{D}'$ of size $m \times n$ such that
 $$M = k_1 m$$
@@
 
+\subsection{GPU Parallelization and CUDA Implementation}
+Placeholder text
+
 \section{GPU Speedup Results}

### 2015-12-05 cuda/motion-tracking/motion_tracking.cu
Initial attempt: parallelization of density map

@@ -39,6 +39,22 @@ void spatial_difference_density_map(double *density_map, int *difference, int wi
 	}
 }
 
+__global__ void gpu_spatial_difference_density_map(double *density_map, int *difference, int width, int height, int horizontal_divisions, int vertical_divisions) {
+	const int r = blockIdx.y * blockDim.y + threadIdx.y;
+	const int c = blockIdx.x * blockDim.x + threadIdx.x;
+	const int i = r * width + c;
+	
+	int horizontal_block_size = width/horizontal_divisions;
+	int vertical_block_size = height/vertical_divisions;
+	int block_size = horizontal_block_size * vertical_block_size;
+	
+	const int scaling_factor = 1000;
+	
+	if (difference[i] == 255) {
+		density_map[(int)(vertical_divisions*r/(double)height) * width + (int)(horizontal_divisions*c/(double)width)] += scaling_factor/(double)block_size;
+	}
+}
+
 void motion_area_estimate(int *motion_area, double *density_map, int width, int height, int horizontal_divisions, int vertical_divisions, double threshold) {
 	int horizontal_block_size = width/horizontal_divisions;
 	int vertical_block_size = height/vertical_divisions;
@@ -119,10 +135,12 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
 	// Lower motion_threshold == more sensitive in picking up motion
 	
     // Allocate space on device
-    int *dev_edges_1, *dev_edges_2, *dev_out;
-    checkCudaErrors(cudaMalloc(&dev_out, width*height*sizeof(int)));
+    int *dev_edges_1, *dev_edges_2, *dev_difference;
     checkCudaErrors(cudaMalloc(&dev_edges_1, width*height*sizeof(int)));
     checkCudaErrors(cudaMalloc(&dev_edges_2, width*height*sizeof(int)));
+    checkCudaErrors(cudaMalloc(&dev_difference, width*height*sizeof(int)));
+    double *dev_density;
+    checkCudaErrors(cudaMalloc(&dev_density, horizontal_divisions*vertical_divisions*sizeof(double)));
 
     // Copy host arrays to device
     checkCudaErrors(cudaMemcpy(dev_edges_1, edges_1, width*height*sizeof(int), cudaMemcpyHostToDevice));
@@ -144,29 +162,34 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
[…]

### 2015-12-05 opencv-python/edge_detect.py
Edge detection and universal image folder

@@ -1,0 +1,24 @@
+import cv2
+from matplotlib import pyplot as plt
+from os import listdir, getcwd
+from os.path import join, isfile
+import datetime as dt
+
+images_path = join(getcwd(), '../images')
+images = [f for f in listdir(images_path) if isfile(join(images_path, f))]
+
+def edge_detect(image_path, display=False):
+    img = cv2.imread(image_path, 0)
+    start_time = dt.datetime.now()
+    edges = cv2.Canny(img, 100, 200)
+    end_time = dt.datetime.now()
+    time = (end_time-start_time).microseconds
+
+    if display:
+        plt.subplot(121), plt.imshow(img, cmap='gray')
+        plt.title('Original Image'), plt.xticks([]), plt.yticks([])
+        plt.subplot(122), plt.imshow(edges, cmap='gray')
+        plt.title('Edge Image'), plt.xticks([]), plt.yticks([])
+        plt.show()
+
+    return edges, time

### 2015-12-05 documents/paper/paper.tex

@@
 We take full advantage of the computational benefits of GPUs by implementing edge detection and motion detection algorithms in CUDA C.
+\par In this paper, we first detail the theory for our edge detection and motion detection algorithms.
+Then, we describe the GPU code implementation of those algorithms with NVIDIA CUDA.
+Finally, we present speedup results achieved by our CUDA implementation with respect to a reference serial, CPU implmentation.
 
@@
 \subsection{GPU Parallelization and CUDA Implementation}
-Placeholder text
+Our CUDA implementation of our motion detection algorithm is similar to that for our implementation of separable convolution and edge detection.
+Using the same 16 threads per block, we divide the input image into the appropriate number of blocks such that the thresholded difference computation at each index $(i, j)$ of Algorithm \ref{difference-matrix-algorithm} has its own thread.
+Thus, each element of $\boldsymbol{D}$ is calculated in parallel.
+\begin{lstlisting}[
+	caption=GPU computation of thresholded difference matrix,
+	label=difference-filter-implementation
+]
+__global__ void difference_filter(int *dev_out, int *edges_1, int *edges_2, int width, int height, int threshold) {
+	const int r = blockIdx.y * blockDim.y + threadIdx.y;
+	const int c = blockIdx.x * blockDim.x + threadIdx.x;
+	const int i = r * width + c;
+
+    // Set it to 0 initially
+    dev_out[i] = 0;
+    int crop_size = 7;
+    if (r > crop_size && c > crop_size && r < height - crop_size && c < width - crop_size && edges_1[i] != edges_2[i]) {
+        // Set to 255 if there is a pixel mismatch
+        dev_out[i] = 255;
+        for (int x_apron = -threshold; x_apron <= threshold; x_apron++) {
+            for (int y_apron = -threshold; y_apron <= threshold; y_apron++) {
+                // Ensure the requested index is within bounds of image
+                if (c + x_apron > 0 && r + y_apron > 0 && c + x_apron < width && r + y_apron < height) {
+                    // Check if there is a matching pixel in the apron, within the threshold
+                    if (edges_1[(r + y_apron) * width + c + x_apron] == edges_2[i]) {
+                        // Set it back to 0 if a corresponding pixel exists within the vicinity of the match
+                        dev_out[i] = 0;
+                    }
+                }
+            }
+        }
+    }
+}
+\end{lstlisting}
 

### 2015-12-05 Object Detect - OpenCV/facedetect.py
started face detect with HAAR classifiers

@@ -1,0 +1,31 @@
+import sys
+import cv2
+
+#cascade = open('/Users/Senthil/Downloads/FaceDetect-master/haarcascade_frontalface_default.xml')
+faceCascade = cv2.CascadeClassifier('/Users/Senthil/Downloads/FaceDetect-master/haarcascade_frontalface_default.xml')
+
+videoFeed = cv2.VideoCapture(0)
+
+while True:
+    ret, frame = videoFeed.read()
+    grayImg = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
+    face = faceCascade.detectMultiScale(
+
+        grayImg,
+        scaleFactor = 1.2,
+        minNeighbors = 5,
+        minSize = (30,30),
+        flags = cv2.CASCADE_SCALE_IMAGE
+
+    )
+
+    for (x, y, w, h) in face:
+        cv2.rectangle(frame, (x, y), (x+w, y+h), (0, 255, 0), 2)
+
+    cv2.imshow('Video', frame)
+
+    if cv2.waitKey(1) & 0xFF == ord('q'):
+        break
+
+videoFeed.release()
+cv2.destroyAllWindows()

### 2015-12-05 opencv-python/edge_detect.py
benchmarking in python

@@ -1,11 +1,7 @@
 import cv2
 from matplotlib import pyplot as plt
-from os import listdir, getcwd
-from os.path import join, isfile
 import datetime as dt
 
-images_path = join(getcwd(), '../images')
-images = [f for f in listdir(images_path) if isfile(join(images_path, f))]
 
 def edge_detect(image_path, display=False):
     img = cv2.imread(image_path, 0)

### 2015-12-05 Object Detect - OpenCV/facedetect.py
face detect with still images

@@ -1,31 +1,62 @@
 import sys
 import cv2
 
-#cascade = open('/Users/Senthil/Downloads/FaceDetect-master/haarcascade_frontalface_default.xml')
-faceCascade = cv2.CascadeClassifier('/Users/Senthil/Downloads/FaceDetect-master/haarcascade_frontalface_default.xml')
+cascade = '/Users/Senthil/Documents/elec-301/Object Detect - OpenCV/haarcascade_frontalface_default.xml'
+faceCascade = cv2.CascadeClassifier(cascade)
 
-videoFeed = cv2.VideoCapture(0)
+image1 = '/Users/Senthil/Documents/elec-301/images/senthil1.jpg'
 
-while True:
-    ret, frame = videoFeed.read()
-    grayImg = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
+def faceDetectWebcam():
+
+    videoFeed = cv2.VideoCapture(0)
+
+    while True:
+        ret, frame = videoFeed.read()
+        grayImg = cv2.cvtColor(frame, cv2.COLOR_BGR2GRAY)
+        face = faceCascade.detectMultiScale(
+
+            grayImg,
+            scaleFactor = 1.2,
+            minNeighbors = 5,
+            minSize = (30,30),
+            flags = cv2.CASCADE_SCALE_IMAGE
+
+        )
+
+        for (x, y, w, h) in face:
+            cv2.rectangle(frame, (x, y), (x+w, y+h), (0, 255, 0), 2)
+
+        cv2.imshow('Video', frame)
+
+        if cv2.waitKey(1) & 0xFF == ord('q'):
+            break
+
+    videoFeed.release()
+    cv2.destroyAllWindows()
+
+
+def faceDetectTiming(images, display = False):
+    image = cv2.imread(images)
+    grayImg = cv2.cvtColor(image, cv2.COLOR_BGR2GRAY)
+
     face = faceCascade.detectMultiScale(
-
-        grayImg,
-        scaleFactor = 1.2,
-        minNeighbors = 5,
-        minSize = (30,30),
-        flags = cv2.CASCADE_SCALE_IMAGE
-
-    )
+            grayImg,
+            scaleFactor = 1.2,
+            minNeighbors = 5,
+            minSize = (30,30),
+            flags = cv2.CASCADE_SCALE_IMAGE
+        )
 
     for (x, y, w, h) in face:
[…]

### 2015-12-05 Object Detect - OpenCV/facedetect.py
added timing to still face detect

@@ -1,10 +1,11 @@
 import sys
 import cv2
+import datetime as dt
 
 cascade = '/Users/Senthil/Documents/elec-301/Object Detect - OpenCV/haarcascade_frontalface_default.xml'
 faceCascade = cv2.CascadeClassifier(cascade)
 
-image1 = '/Users/Senthil/Documents/elec-301/images/senthil1.jpg'
+image1 = '/Users/Senthil/Documents/elec-301/images/friends1.jpg'
 
 def faceDetectWebcam():
 
@@ -35,17 +36,27 @@ def faceDetectWebcam():
     cv2.destroyAllWindows()
 
 
-def faceDetectTiming(images, display = False):
+#train classifiers
+
+def faceDetectTiming(images):
     image = cv2.imread(images)
+
+    start_time = dt.datetime.now()
+
     grayImg = cv2.cvtColor(image, cv2.COLOR_BGR2GRAY)
 
     face = faceCascade.detectMultiScale(
             grayImg,
-            scaleFactor = 1.2,
+            scaleFactor = 1.1,
             minNeighbors = 5,
             minSize = (30,30),
             flags = cv2.CASCADE_SCALE_IMAGE
         )
+
+    end_time = dt.datetime.now()
+    time = (end_time-start_time).microseconds
+
+    print("detected in:" + " " + str(time) + "ms")
 
     for (x, y, w, h) in face:
         cv2.rectangle(image, (x, y), (x+w, y+h), (0, 255, 0), 2)

### 2015-12-05 opencv-python/benchmark.py
motion detection

@@ -1,10 +1,12 @@
 from os import listdir, getcwd
 from os.path import join, isfile
 import edge_detect as edge
+import cv2
 
 images_path = join(getcwd(), '../images')
 images = [f for f in listdir(images_path) if isfile(join(images_path, f))]
 
-for img in images:
-    edges, time = edge.edge_detect(join(images_path, img))
-    print img+":", time
+for img_path in images:
+    img = cv2.imread(join(images_path, img_path), 0)
+    edges, time = edge.edge_detect(img)
+    print img_path+":", time

### 2015-12-05 opencv-python/edge_detect.py

@@ -3,8 +3,7 @@ from matplotlib import pyplot as plt
 import datetime as dt
 
 
-def edge_detect(image_path, display=False):
-    img = cv2.imread(image_path, 0)
+def edge_detect(img, display=False):
     start_time = dt.datetime.now()
     edges = cv2.Canny(img, 100, 200)
     end_time = dt.datetime.now()

### 2015-12-05 opencv-python/motion_detection.py

@@ -1,0 +1,25 @@
+import numpy as np
+import cv2
+from matplotlib import pyplot as plt
+import edge_detect as edge
+
+cap = cv2.VideoCapture(0)
+
+if not(cap.isOpened()):
+    cap.open()
+
+last_frame = np.zeros((480, 640), np.uint8)
+last_frame_edge = edge.edge_detect(last_frame)[0]
+while(cap.isOpened()):
+    ret, frame = cap.read()
+    frame_edge = edge.edge_detect(frame)[0]
+    # cv2.imshow('frame', frame)  # normal picture
+    cv2.imshow('edge', frame_edge)  # edge detection picture
+    cv2.imshow('motion', frame_edge-last_frame_edge)
+    last_frame = frame
+    last_frame_edge = frame_edge
+    if cv2.waitKey(1) & 0xFF == ord('q'):
+        break
+
+cap.release()
+cv2.destroyAllWindows()

### 2015-12-06 cuda/motion-tracking/motion_tracking.cu
Fixed parallelized density map implementation

@@ -17,18 +17,18 @@ using namespace cv;
 #define TY 8
 
 
-void spatial_difference_density_map(double *density_map, int *difference, int width, int height, int horizontal_divisions, int vertical_divisions) {
+void serial_spatial_difference_density_map(double *density_map, int *difference, int width, int height, int horizontal_divisions, int vertical_divisions) {
 	int horizontal_block_size = width/horizontal_divisions;
 	int vertical_block_size = height/vertical_divisions;
 	int block_size = horizontal_block_size * vertical_block_size;
 	
 	const int scaling_factor = 1000;  // Used to linearly scale density map to units millipixels/pixels^2 (if that makes any sense?)
 	
-	for (int block_x_index = 0; block_x_index < horizontal_divisions; block_x_index++) {
-		for (int block_y_index = 0; block_y_index < vertical_divisions; block_y_index++) {
+	for (int block_x_index = 0; block_x_index < horizontal_divisions - 1; block_x_index++) {
+		for (int block_y_index = 0; block_y_index < vertical_divisions - 1; block_y_index++) {
 			int num_differences = 0;
-			for (int x = (block_x_index - 1) * horizontal_block_size; x < block_x_index * horizontal_block_size; x++) {
-				for (int y = (block_y_index - 1) * vertical_block_size; y < block_y_index * vertical_block_size; y++) {
+			for (int x = block_x_index * horizontal_block_size; x < (block_x_index + 1) * horizontal_block_size; x++) {
+				for (int y = block_y_index * vertical_block_size; y < (block_y_index + 1) * vertical_block_size; y++) {
 					if (x > 0 && y > 0 && x < width && y < height && difference[y * width + x] == 255) {
 						num_differences++;
 					}
@@ -39,19 +39,18 @@ void spatial_difference_density_map(double *density_map, int *difference, int wi
 	}
 }
 
-__global__ void gpu_spatial_difference_density_map(double *density_map, int *difference, int width, int height, int horizontal_divisions, int vertical_divisions) {
-	const int r = blockIdx.y * blockDim.y + threadIdx.y;
-	const int c = blockIdx.x * blockDim.x + threadIdx.x;
-	const int i = r * width + c;
[…]

### 2015-12-06 cuda/motion-tracking/motion_tracking.cu
Parallelized generation of motion area image

@@ -13,8 +13,8 @@
 
 using namespace cv;
 
-#define TX 8
-#define TY 8
+#define TX 16
+#define TY 16
 
 
 void serial_spatial_difference_density_map(double *density_map, int *difference, int width, int height, int horizontal_divisions, int vertical_divisions) {
@@ -54,7 +54,24 @@ __global__ void spatial_difference_density_map(double *density_map, int *differe
 	}
 }
 
-void motion_area_estimate(int *motion_area, double *density_map, int width, int height, int horizontal_divisions, int vertical_divisions, double threshold) {
+__global__ void motion_area_estimate(int *motion_area, double *density_map, int width, int height, int horizontal_divisions, int vertical_divisions, double threshold) {
+	int r = blockIdx.y * blockDim.y + threadIdx.y;
+	int c = blockIdx.x * blockDim.x + threadIdx.x;
+	int i = r * width + c;
+	
+	int density_map_index = (int)(vertical_divisions*r/(double)height) * horizontal_divisions + (int)(horizontal_divisions*c/(double)width);
+	
+	int horizontal_block_size = width/horizontal_divisions;
+	int vertical_block_size = height/vertical_divisions;
+	
+	if (density_map[density_map_index] >= threshold) {
+		motion_area[i] = 255;
+	} else {
+		motion_area[i] = 0;
+	}
+}
+
+void serial_motion_area_estimate(int *motion_area, double *density_map, int width, int height, int horizontal_divisions, int vertical_divisions, double threshold) {
 	int horizontal_block_size = width/horizontal_divisions;
 	int vertical_block_size = height/vertical_divisions;
 	
@@ -129,10 +146,11 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
 	// Lower motion_threshold == more sensitive in picking up motion
 	
     // Allocate space on device
-    int *dev_edges_1, *dev_edges_2, *dev_difference;
+    int *dev_edges_1, *dev_edges_2, *dev_difference, *dev_motion_area;
     checkCudaErrors(cudaMalloc(&dev_edges_1, width*height*sizeof(int)));
     checkCudaErrors(cudaMalloc(&dev_edges_2, width*height*sizeof(int)));
     checkCudaErrors(cudaMalloc(&dev_difference, width*height*sizeof(int)));
+    checkCudaErrors(cudaMalloc(&dev_motion_area, width*height*sizeof(int)));
     
[…]

### 2015-12-06 cuda/benchmark/benchmark.cu
Change benchmark to use standardized test images

@@ -77,17 +77,21 @@ void compare_separable_convolution_speedup(const char **images, int num_images) 
 }
 
 int main() {
-	const char *images[6];
-	images[0] = "../images/nvidia_100_100.jpg";
-	images[1] = "../images/nvidia_500_500.jpg";
-	images[2] = "../images/nvidia_1000_1000.jpg";
-	images[3] = "../images/nvidia_1500_1500.jpg";
-	images[4] = "../images/nvidia_2000_2000.jpg";
-	images[5] = "../images/nvidia_2500_2500.jpg";
+    const char *images[9];
+    images[0] = "../../images/city_100_100.jpg";
+    images[1] = "../../images/city_500_500.jpg";
+    images[2] = "../../images/city_1000_1000.jpg";
+    images[3] = "../../images/city_2000_2000.jpg";
+    images[4] = "../../images/city_3000_3000.jpg";
+    images[5] = "../../images/city_4000_4000.jpg";
+    images[6] = "../../images/city_5000_5000.jpg";
+    images[7] = "../../images/city_6000_6000.jpg";
+    images[8] = "../../images/city_7500_7500.jpg";
 	
-	printf("==========SERIAL NAIVE VS SEPARABLE CONVOLUTION COMPARISON==========\n");
-    compare_naive_separable_convolution(images, 6);
+    printf("==========SERIAL NAIVE VS SEPARABLE CONVOLUTION COMPARISON==========\n");
+    compare_naive_separable_convolution(images, 9);
     printf("==========CPU VS GPU SEPARABLE CONVOLUTION SPEEDUP==========\n");
-    compare_separable_convolution_speedup(images, 6);
+    compare_separable_convolution_speedup(images, 9);
+
     return 0;
 }

### 2015-12-06 documents/paper/paper.tex
Add some speedup results

@@
 Generally, GPGPU is used to exploit the improved multithreaded performance and raw floating-point computational ability of GPUs over CPUs.
-For example, on modern hardware, an NVIDIA GeForce GTX 970 (1664 CUDA cores) exhibits peak single-precision floating point performance of nearly 3500 GFLOPS (floating-point operations per second), while an Intel Core i7 4790K (8C, 16T) achieves 100 GFLOPS.
+For example, on modern hardware, an NVIDIA GeForce GTX 970 (1664 CUDA cores) exhibits peak single-precision floating point performance of nearly 3500 GFLOPS (floating-point operations per second), while an Intel Core i7 4790K (4C, 8T) achieves 100 GFLOPS.
 Our goal is to demonstrate the performance of GPU computing by solving a handful of existing problems in computer vision with CUDA: namely, edge detection and motion detection.
@@
 \section{GPU Speedup Results}
-Placeholder text
+Our code for separable convolution, computation of the image's gradient magnitude and angle, non-maximum suppression and selective thresholding, and computation of the thresholded difference and spatial difference density matrices has both serial CPU and parallel CUDA implementations.
+Here, we present speedup results for each of these procedures on various image sizes obtained on a system with the following hardware:
+\begin{itemize}
+	\item Intel Core i7 4790K @ 4.7 GHz (4C, 8T)
+	\item 16 GB of DDR3 memory @ 1600 MHz
+	\item NVIDIA GeForce GTX 750 (512 CUDA cores, 1437 MHz core clock, 6500 MHz memory clock, 2 GB video memory)
+\end{itemize}
+
+\subsection{Separable convolution}
+The below results were obtained by convolving the same image, resized to various image dimensions, with a separated Gaussian filter of kernel size $k = 3$.
+\begin{table}[H]
+	\small
+	\centering
+	\caption{Speedup of Separable Convolution}
+	\label{separable-convolution-speedup}
+	\begin{tabular}{llllll}
+	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
+	 $100 \times 100$ & $0.000191$ & $0.054855$ & $0.003482$ \\
+	 $500 \times 500$ & $0.004748$ & $0.002728$ & $1.740469$ \\
+	 $1000 \times 1000$ & $0.018850$ & $0.008431$ & $2.235796$ \\
+	 $2000 \times 2000$ & $0.074475$ & $0.034503$ & $2.158508$ \\
+	 $3000 \times 3000$ & $0.167459$ & $0.072181$ & $2.319987$ \\
+	 $4000 \times 4000$ & $0.299840$ & $0.136641$ & $2.194363$ \\
+	 $5000 \times 5000$ & $0.465225$ & $0.194512$ & $2.391755$ \\
+	 $6000 \times 6000$ & $0.673266$ & $0.276730$ & $2.432935$ \\
+	 $7500 \times 7500$ & $1.052294$ & $0.439262$ & $2.395595$ \\
+	\end{tabular}
+\end{table}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{separable_convolution_speedup_graph.jpg}
+	\caption{Visualization of separable convolution speedup versus pixel count}
+    \label{separable-convolution-speedup-graph}
+\end{figure}
+
+\subsection{Non-maximum suppression and selective thresholding}
+
+\subsection{Thresholded spatial difference density matrix}
 

### 2015-12-06 cuda/edge-detect/edge_detect.cu
Modularized GPU edge detection methods

@@ -72,14 +72,12 @@ __global__ void thresholding_and_suppression(int *dev_output, double *dev_magnit
     }
 }
 
-double serial_thresholding_and_suppression(int *dev_output, int input_width, int input_height, int *g_x, int *g_y, int high_threshold, int low_threshold) {
-	struct timeval tv1, tv2;
-	gettimeofday(&tv1, NULL);
+void serial_thresholding_and_suppression(int *output, int input_width, int input_height, int *g_x, int *g_y, int high_threshold, int low_threshold) {
 	for (int r = 0; r < input_height; r++) {
 		for (int c = 0; c < input_width; c++) {
 			int i = r * input_width + c;
 			// First, initialize the current pixel to zero (non-edge)
-			dev_output[i] = 0;
+			output[i] = 0;
 			// Boundary conditions
 			if (r > 1 && c > 1 && r < input_height - 1 && c < input_width - 1) {
 				double magnitude = sqrt(pow((double)g_x[i], 2) + pow((double)g_y[i], 2));
@@ -108,12 +106,12 @@ double serial_thresholding_and_suppression(int *dev_output, int input_width, int
 					// Consider a surrounding apron around the current pixel to catch potentially disconnected pixel nodes
 					int apron_size = 2;
 					if (is_vertical_max || is_horizontal_max || is_positive_diagonal_max || is_negative_diagonal_max) {
-						dev_output[i] = 255;
+						output[i] = 255;
 						for (int m = -apron_size; m <= apron_size; m++) {
 							for (int n = -apron_size; n <= apron_size; n++) {
 								if (r + m > 0 && r + m < input_height && c + n > 0 && c + n < input_width) {
 									if (sqrt(pow((double)g_x[(r + m) * input_width + c + n], 2) + pow((double)g_y[(r + m) * input_width + c + n], 2)) > low_threshold) {
-										dev_output[(r + m) * input_width + c + n] = 255;
+										output[(r + m) * input_width + c + n] = 255;
 									}
 								}
 							}
@@ -123,10 +121,12 @@ double serial_thresholding_and_suppression(int *dev_output, int input_width, int
 			}
 		}
 	}
-	gettimeofday(&tv2, NULL);
-	double time_spent = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
[…]

### 2015-12-06 cuda/edge-detect/edge_detect.h

@@ -3,4 +3,8 @@
 
 void edge_detect(int *edges_out, int *gx_out, int *gy_out, int *image, int image_width, int image_height, int high_threshold, int low_threshold);
 
+void gradient_magnitude_angle_thresholding_and_suppresion(double *dev_magnitude, double *dev_angle, int sobel_out_width, int sobel_out_height, int *dev_gx, int *dev_gy, int *dev_edges, int *edges_out, int high_threshold, int low_threshold, dim3 grid_size, dim3 block_size);
+
+void serial_thresholding_and_suppression(int *output, int input_width, int input_height, int *g_x, int *g_y, int high_threshold, int low_threshold);
+
 #endif

### 2015-12-06 cuda/benchmark/benchmark.cu
Non-maximum suppression and selective thresholding benchmark

@@ -4,8 +4,13 @@
 #include <opencv2/highgui/highgui.hpp>
 #include <iostream>
 #include "../separable-convolution/separable_convolution.h"
+#include "../edge-detect/edge_detect.h"
+#include "../helper/helper_cuda.h"
 
 using namespace cv;
+
+#define TX 16
+#define TY 16
 
 
 void compare_naive_separable_convolution(const char **images, int num_images) {
@@ -76,6 +81,77 @@ void compare_separable_convolution_speedup(const char **images, int num_images) 
 	}
 }
 
+void non_maximum_suppression_selective_thresholding_speedup(const char **images, int num_images) {
+	int high_threshold = 70;
+	int low_threshold = 50;
+	
+	for (int image_index = 0; image_index < num_images; image_index++) {
+		Mat raw_image = imread(images[image_index], 0);
+		int *image = (int *)malloc(raw_image.cols * raw_image.rows * sizeof(int));
+		for (int i = 0; i < raw_image.rows; i++) {
+			for (int j = 0; j < raw_image.cols; j++) {
+				image[i * raw_image.cols + j] = raw_image.at<uchar>(i, j);
+			}
+		}
+		
+		int *gx_out = (int *)malloc((raw_image.cols + 4) * (raw_image.rows + 4) * sizeof(int));
+		int *gy_out = (int *)malloc((raw_image.cols + 4) * (raw_image.rows + 4) * sizeof(int));
+		int *edges_out = (int *)malloc((raw_image.cols + 4) * (raw_image.rows + 4) * sizeof(int));
+		
+		int kernel_size = 3;
+		int sobel_out_width = raw_image.cols + kernel_size - 1;
+		int sobel_out_height = raw_image.rows + kernel_size - 1;
+		dim3 block_size(TX, TY);
+		int bx = sobel_out_width/block_size.x;
+		int by = sobel_out_height/block_size.y;
+		dim3 grid_size = dim3(bx, by);
+		
+		// Horizontal direction
+		int gx_horizontal[3] = {1, 0, -1};
+		int gx_vertical[3] = {1, 2, 1};
+		separable_convolve(gx_out, image, sobel_out_width, sobel_out_height, gx_horizontal, gx_vertical, 3, 1);
+		
+		// Vertical direction
+		int gy_horizontal[3] = {1, 2, 1};
+		int gy_vertical[3] = {1, 0, -1};
[…]

### 2015-12-06 cuda/edge-detect/edge_detect.cu
Code cleanup

@@ -152,7 +152,6 @@ void edge_detect(int *edges_out, int *gx_out, int *gy_out, int *image, int image
 	separable_convolve(gy_out, image, sobel_out_width, sobel_out_height, gy_horizontal, gy_vertical, 3, 1);
 
 	// Magnitude and thresholding
-//	static int *serial_edges = (int *)malloc(sobel_out_width * sobel_out_height * sizeof(int));
 	int *dev_edges, *dev_gx, *dev_gy;
 	double *dev_magnitude, *dev_angle;
 	
@@ -165,18 +164,7 @@ void edge_detect(int *edges_out, int *gx_out, int *gy_out, int *image, int image
 	checkCudaErrors(cudaMemcpy(dev_gx, gx_out, sobel_out_width * sobel_out_height * sizeof(int), cudaMemcpyHostToDevice));
 	checkCudaErrors(cudaMemcpy(dev_gy, gy_out, sobel_out_width * sobel_out_height * sizeof(int), cudaMemcpyHostToDevice));
 	
-	// Serial comparison
-//	printf("=====THRESHOLDING AND NON-MAXIMUM SUPPRESSION=====\n");
-//	double serial_computation_time = serial_thresholding_and_suppression(serial_edges, sobel_out_width, sobel_out_height, gx_out, gy_out, high_threshold, low_threshold);
-	
-	// Parallelization
-	struct timeval tv1, tv2;
-	gettimeofday(&tv1, NULL);
 	gradient_magnitude_angle_thresholding_and_suppresion(dev_magnitude, dev_angle, sobel_out_width, sobel_out_height, dev_gx, dev_gy, dev_edges, edges_out, high_threshold, low_threshold, grid_size, block_size);
-	gettimeofday(&tv2, NULL);
-	double parallel_computation_time = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
-//	printf("Parallel thresholding and non-maximum suppression execution time: %f seconds\n", parallel_computation_time);
-//	printf("Estimated parallelization speedup: %f\n", serial_computation_time/parallel_computation_time);
 
 	// Free GPU memory
 	checkCudaErrors(cudaFree(dev_edges));

### 2015-12-06 cuda/motion-tracking/motion_tracking.cu

@@ -60,10 +60,7 @@ __global__ void motion_area_estimate(int *motion_area, double *density_map, int 
 	int i = r * width + c;
 	
 	int density_map_index = (int)(vertical_divisions*r/(double)height) * horizontal_divisions + (int)(horizontal_divisions*c/(double)width);
-	
-	int horizontal_block_size = width/horizontal_divisions;
-	int vertical_block_size = height/vertical_divisions;
-	
+
 	if (density_map[density_map_index] >= threshold) {
 		motion_area[i] = 255;
 	} else {

### 2015-12-06 opencv-object-detect/facedetect.py
Add Daft Punk mask

@@ -1,11 +1,13 @@
 import sys
 import cv2
 import datetime as dt
+import numpy
 
-cascade = '/Users/Senthil/Documents/elec-301/Object Detect - OpenCV/haarcascade_frontalface_default.xml'
+cascade = '/Users/Senthil/Documents/elec-301/opencv-object-detect/haarcascade_frontalface_default.xml'
 faceCascade = cv2.CascadeClassifier(cascade)
 
-image1 = '/Users/Senthil/Documents/elec-301/images/friends1.jpg'
+image1 = '/Users/Senthil/Documents/elec-301/images/birthday1.jpg'
+daftPunk = cv2.imread('/Users/Senthil/Downloads/DaftPunk.png', cv2.IMREAD_UNCHANGED)
 
 def faceDetectWebcam():
 
@@ -25,7 +27,12 @@ def faceDetectWebcam():
         )
 
         for (x, y, w, h) in face:
-            cv2.rectangle(frame, (x, y), (x+w, y+h), (0, 255, 0), 2)
+
+            recog = cv2.resize(daftPunk, (w, h), interpolation=cv2.INTER_AREA)
+            #print(recog.shape, recog.shape[0], recog.shape[1], w, h)
+            for c in range(0,3):
+                frame[y:y+h, x:x+w, c] = recog[:,:,c] * (recog[:,:,3]/255.0) +  frame[y:y+h, x:x+w, c] * (1.0 - recog[:,:,3]/255.0)
+            #cv2.rectangle(frame, (x, y), (x+w, y+h), (0, 255, 0), 2)
 
         cv2.imshow('Video', frame)
 
@@ -37,10 +44,10 @@ def faceDetectWebcam():
 
 
 #train classifiers
+#train to recognize rich b's face, and apply a daft punk mask
 
 def faceDetectTiming(images):
-    image = cv2.imread(images)
-
+    image = cv2.imread(images, cv2.IMREAD_UNCHANGED)
     start_time = dt.datetime.now()
 
     grayImg = cv2.cvtColor(image, cv2.COLOR_BGR2GRAY)
@@ -55,10 +62,14 @@ def faceDetectTiming(images):
 
     end_time = dt.datetime.now()
     time = (end_time-start_time).microseconds
-
     print("detected in:" + " " + str(time) + "ms")
 
     for (x, y, w, h) in face:
[…]

### 2015-12-06 opencv-object-detect/facedetect.py
face detect cleanup

@@ -9,6 +9,7 @@ faceCascade = cv2.CascadeClassifier(cascade)
 image1 = '/Users/Senthil/Documents/elec-301/images/birthday1.jpg'
 daftPunk = cv2.imread('/Users/Senthil/Downloads/DaftPunk.png', cv2.IMREAD_UNCHANGED)
 
+
 def faceDetectWebcam():
 
     videoFeed = cv2.VideoCapture(0)
@@ -19,20 +20,20 @@ def faceDetectWebcam():
         face = faceCascade.detectMultiScale(
 
             grayImg,
-            scaleFactor = 1.2,
-            minNeighbors = 5,
-            minSize = (30,30),
-            flags = cv2.CASCADE_SCALE_IMAGE
+            scaleFactor=1.2,
+            minNeighbors=5,
+            minSize=(30,30),
+            flags=cv2.CASCADE_SCALE_IMAGE
 
         )
 
         for (x, y, w, h) in face:
 
             recog = cv2.resize(daftPunk, (w, h), interpolation=cv2.INTER_AREA)
-            #print(recog.shape, recog.shape[0], recog.shape[1], w, h)
-            for c in range(0,3):
-                frame[y:y+h, x:x+w, c] = recog[:,:,c] * (recog[:,:,3]/255.0) +  frame[y:y+h, x:x+w, c] * (1.0 - recog[:,:,3]/255.0)
-            #cv2.rectangle(frame, (x, y), (x+w, y+h), (0, 255, 0), 2)
+            # print(recog.shape, recog.shape[0], recog.shape[1], w, h)
+            for c in range(0, 3):
+                frame[y:y+h, x:x+w, c] = recog[:, :, c] * (recog[:, :, 3]/255.0) + frame[y:y+h, x:x+w, c] * (1.0 - recog[:, :, 3]/255.0)
+            # cv2.rectangle(frame, (x, y), (x+w, y+h), (0, 255, 0), 2)
 
         cv2.imshow('Video', frame)
 
@@ -43,8 +44,8 @@ def faceDetectWebcam():
     cv2.destroyAllWindows()
 
 
-#train classifiers
-#train to recognize rich b's face, and apply a daft punk mask
+# train classifiers
+# train to recognize rich b's face, and apply a daft punk mask
 
 def faceDetectTiming(images):
     image = cv2.imread(images, cv2.IMREAD_UNCHANGED)
[…]

### 2015-12-06 opencv-object-detect/facedetect.py
added daft punk image to repo

@@ -7,7 +7,7 @@ cascade = '/Users/Senthil/Documents/elec-301/opencv-object-detect/haarcascade_fr
 faceCascade = cv2.CascadeClassifier(cascade)
 
 image1 = '/Users/Senthil/Documents/elec-301/images/birthday1.jpg'
-daftPunk = cv2.imread('/Users/Senthil/Downloads/DaftPunk.png', cv2.IMREAD_UNCHANGED)
+daftPunk = cv2.imread('/Users/Senthil/Documents/elec-301/opencv-object-detect/DaftPunk.png', cv2.IMREAD_UNCHANGED)
 
 
 def faceDetectWebcam():

### 2015-12-06 cuda/benchmark/benchmark.cu
Motion area estimation benchmark

@@ -4,6 +4,7 @@
 #include <opencv2/highgui/highgui.hpp>
 #include <iostream>
 #include "../separable-convolution/separable_convolution.h"
+#include "../motion-tracking/motion_tracking.h"
 #include "../edge-detect/edge_detect.h"
 #include "../helper/helper_cuda.h"
 
@@ -152,6 +153,44 @@ void non_maximum_suppression_selective_thresholding_speedup(const char **images,
 	}
 }
 
+void motion_area_estimation_speedup(const char **images, int num_images) {
+	// Uses random matrices for the difference
+	double movement_threshold = 5.0;
+	int motion_threshold = 4;
+	int horizontal_divisions = 5;
+	int vertical_divisions = 5;
+	
+	struct timeval tv1, tv2;
+	
+	for (int image_index = 0; image_index < num_images; image_index++) {
+		Mat raw_image = imread(images[image_index], 0);
+		int *motion_area = (int *)malloc(raw_image.cols * raw_image.rows * sizeof(int));
+		int *difference = (int *)malloc(raw_image.cols * raw_image.rows * sizeof(int));
+		int *edges_1 = (int *)malloc(raw_image.cols * raw_image.rows * sizeof(int));
+		int *edges_2 = (int *)malloc(raw_image.cols * raw_image.rows * sizeof(int));
+		
+		for (int i = 0; i < raw_image.cols * raw_image.rows; i++) {
+			edges_1[i] = rand() % 2;
+			edges_2[i] = rand() % 2;
+		}
+		
+		gettimeofday(&tv1, NULL);
+		motion_detect(motion_area, difference, edges_1, edges_2, raw_image.cols, raw_image.rows, movement_threshold, motion_threshold, horizontal_divisions, vertical_divisions);
+		gettimeofday(&tv2, NULL);
+		double parallel_computation_time = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
+		
+		gettimeofday(&tv1, NULL);
+		serial_motion_detect(motion_area, difference, edges_1, edges_2, raw_image.cols, raw_image.rows, movement_threshold, motion_threshold, horizontal_divisions, vertical_divisions);
+		gettimeofday(&tv2, NULL);
+		double serial_computation_time = (double) (tv2.tv_usec - tv1.tv_usec) / 1000000 + (double) (tv2.tv_sec - tv1.tv_sec);
+		
+		printf("Test image: %s\n", images[image_index]);
[…]

### 2015-12-06 cuda/motion-tracking/motion_tracking.cu
Motion detection modularization

@@ -136,6 +136,14 @@ void serial_difference_filter(int *difference, int *edges_1, int *edges_2, int w
     }
 }
 
+void serial_motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2, int width, int height, int movement_threshold, int motion_threshold, int horizontal_divisions, int vertical_divisions) {
+    double *density_map = (double *)calloc(horizontal_divisions * vertical_divisions, sizeof(double));    
+    serial_difference_filter(difference, edges_1, edges_2, width, height, movement_threshold);
+	serial_spatial_difference_density_map(density_map, difference, width, height, horizontal_divisions, vertical_divisions);
+	serial_motion_area_estimate(motion_area, density_map, width, height, horizontal_divisions, vertical_divisions, motion_threshold);
+	free(density_map);
+}
+
 void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2, int width, int height, int movement_threshold, int motion_threshold, int horizontal_divisions, int vertical_divisions) {
 	// Note: movement_threshold refers to the pixel apron around which the difference filter attempts to look for differences.
 	// Higher movement_threshold == more leniency in how much camera shake is tolerated
@@ -189,110 +197,4 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
 	
 	// Free allocated host memory
 	free(density_map);
-}
-
-int main() {
-	// Config constants
-	double movement_threshold = 5.0;  // Camera shake tolerance
-	int motion_threshold = 4;  // Threshold above which motion is registered
-	int edge_detect_high_threshold = 70;
-	int edge_detect_low_threshold = 50;
-	int horizontal_divisions = 12;
-	int vertical_divisions = 10;
-	
-	// Initialize windows
-	namedWindow("Input", WINDOW_NORMAL);
-	namedWindow("Difference", WINDOW_NORMAL);
-	namedWindow("Edges", WINDOW_NORMAL);
-	namedWindow("Motion", WINDOW_NORMAL);
-	
-	// Initialize video stream
-	VideoCapture cap(0);
-	Mat temp_image;
-	cap >> temp_image;
-	
-	// Allocate host memory
[…]

### 2015-12-06 cuda/motion-tracking/motion_tracking.h

@@ -1,0 +1,8 @@
+#ifndef MOTION_TRACKING_H
+#define MOTION_TRACKING_H
+
+void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2, int width, int height, int movement_threshold, int motion_threshold, int horizontal_divisions, int vertical_divisions);
+
+void serial_motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2, int width, int height, int movement_threshold, int motion_threshold, int horizontal_divisions, int vertical_divisions);
+
+#endif

### 2015-12-06 cuda/main/main.cu
Main file for running real-time motion detection

@@ -1,0 +1,119 @@
+#include <stdio.h>
+#include <sys/time.h>
+#include <opencv2/core/core.hpp>
+#include <opencv2/imgproc/imgproc.hpp>
+#include <opencv2/highgui/highgui.hpp>
+#include <iostream>
+#include "../edge-detect/edge_detect.h"
+#include "../motion-tracking/motion_tracking.h"
+#include "../separable-convolution/separable_convolution.h"
+#include "../helper/helper_cuda.h"
+
+using namespace cv;
+
+
+int main() {
+	// Config constants
+	double movement_threshold = 5.0;  // Camera shake tolerance
+	int motion_threshold = 4;  // Threshold above which motion is registered
+	int edge_detect_high_threshold = 70;
+	int edge_detect_low_threshold = 50;
+	int horizontal_divisions = 12;
+	int vertical_divisions = 10;
+	
+	// Initialize windows
+	namedWindow("Input", WINDOW_NORMAL);
+	namedWindow("Difference", WINDOW_NORMAL);
+	namedWindow("Edges", WINDOW_NORMAL);
+	namedWindow("Motion", WINDOW_NORMAL);
+	
+	// Initialize video stream
+	VideoCapture cap(0);
+	Mat temp_image;
+	cap >> temp_image;
+	
+	// Allocate host memory
+	// Do it all here so as to prevent new allocations happening during the infinite loop below
+	int *edges_1 = (int *)malloc((temp_image.cols + 4) * (temp_image.rows + 4) * sizeof(int));
+	int *edges_2 = (int *)malloc((temp_image.cols + 4) * (temp_image.rows + 4) * sizeof(int));
+	int *gx_out = (int *)malloc((temp_image.cols + 4) * (temp_image.rows + 4) * sizeof(int));
+	int *gy_out = (int *)malloc((temp_image.cols + 4) * (temp_image.rows + 4) * sizeof(int));
+	int *x_1 = (int *)malloc(temp_image.cols * temp_image.rows * sizeof(int));
+	int *x_2 = (int *)malloc(temp_image.cols * temp_image.rows * sizeof(int));
+	int *gaussian_out_1 = (int *)malloc((temp_image.cols + 2) * (temp_image.rows + 2) * sizeof(int));
+	int *gaussian_out_2 = (int *)malloc((temp_image.cols + 2) * (temp_image.rows + 2) * sizeof(int));
[…]

### 2015-12-06 documents/paper/paper.tex
Speedup results for non-maximum suppression thresholding and motion area estimation

@@
 	\item 16 GB of DDR3 memory @ 1600 MHz
-	\item NVIDIA GeForce GTX 750 (512 CUDA cores, 1437 MHz core clock, 6500 MHz memory clock, 2 GB video memory)
+	\item NVIDIA GeForce GTX 750 (512 CUDA cores, 1653 MHz core clock, 6510 MHz memory clock, 2 GB video memory)
 \end{itemize}
 
-\subsection{Separable convolution}
+\subsection{Separable Convolution}
 The below results were obtained by convolving the same image, resized to various image dimensions, with a separated Gaussian filter of kernel size $k = 3$.
@@
 
-\subsection{Non-maximum suppression and selective thresholding}
+\subsection{Non-maximum Suppression and Selective Thresholding}
+As discussed previously, our CUDA implementation of non-maximum suppression and selective thresholding is split into two parallel tasks that run one after the other: computation of the gradient magnitude and angle, followed by the actual edge selection algorithm.
+The computations for the speedup numbers below reflect the full procedure of calculating the gradient's magnitude and angle, selecting edges, and finally copying the results to host memory.
+\begin{table}[H]
+	\small
+	\centering
+	\caption{Speedup of Non-maximum Suppression and Selective Thresholding}
+	\label{non-maximum-suppression-selective-thresholding-speedup}
+	\begin{tabular}{llllll}
+	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
+	 $100 \times 100$ & $0.001701$ & $0.000332$ & $5.123494$ \\
+	 $500 \times 500$ & $0.027121$ & $0.009488$ & $2.858453$ \\
+	 $1000 \times 1000$ & $0.083744$ & $0.028291$ & $2.960093$ \\
+	 $2000 \times 2000$ & $0.237126$ & $0.104160$ & $2.276555$ \\
+	 $3000 \times 3000$ & $0.436223$ & $0.207502$ & $2.102259$ \\
+	 $4000 \times 4000$ & $0.608094$ & $0.341184$ & $1.782305$ \\
+	 $5000 \times 5000$ & $0.823099$ & $0.484530$ & $1.698758$ \\
+	 $6000 \times 6000$ & $1.020302$ & $0.668001$ & $1.527396$ \\
+	\end{tabular}
+\end{table}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{non_maximum_suppression_selective_thresholding_speedup_graph.jpg}
+	\caption{Visualization of non-maximum suppression and selective thresholding speedup versus pixel count}
+    \label{non-maximum-suppression-selective-thresholding-speedup-graph}
+\end{figure}
 
-\subsection{Thresholded spatial difference density matrix}
+\subsection{Thresholded Difference Density Matrix and Motion Area Estimation}
+The below results reflect the speedup in the complete procedure of calculating the difference matrix $\boldsymbol{D}_0$, determining the thresholded difference matrix $\boldsymbol{D}$, building a difference density matrix $\boldsymbol{D}'$, and finally generating an image representing the estimated motion area.
+\begin{table}[H]
+	\small
+	\centering
+	\caption{Speedup of Motion Area Estimation}
+	\label{motion-area-speedup}
+	\begin{tabular}{llllll}
+	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
+	 $100 \times 100$ & $0.004406$ & $0.002055$ & $2.144039$ \\
+	 $500 \times 500$ & $0.116421$ & $0.028079$ & $4.146195$ \\
+	 $1000 \times 1000$ & $0.490470$ & $0.107598$ & $4.558356$ \\
+	 $2000 \times 2000$ & $1.863727$ & $0.424997$ & $4.385271$ \\
+	 $3000 \times 3000$ & $4.334172$ & $0.952842$ & $4.548679$ \\
+	 $4000 \times 4000$ & $7.730778$ & $1.778392$ & $4.347061$ \\
+	 $5000 \times 5000$ & $12.174515$ & $2.660347$ & $4.576288$ \\
+	 $6000 \times 6000$ & $17.628128$ & $3.865616$ & $4.560238$ \\
+	 $7500 \times 7500$ & $28.354028$ & $6.118255$ & $4.634333$ \\
+	\end{tabular}
+\end{table}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{motion_area_estimation_speedup_graph.jpg}
+	\caption{Visualization of motion area estimation speedup versus pixel count}
+    \label{motion-area-estimation-speedup-graph}
+\end{figure}
+
+\subsection{Real-time Motion Detection Performance}
 
 \section{Conclusion}
-Placeholder text
+TODO: Some discussion of the results is warranted
 

### 2015-12-07 documents/paper/paper.tex
Details on difference density matrix CUDA implementation

@@
 	\caption{Final estimate of motion area of Figures \ref{desk-1-original} and \ref{desk-2-original} based on the density map of Figure \ref{desk-difference-density}}
-    \label{desk-difference-density}
+    \label{desk-motion-area}
 \end{figure}
@@
         }
+    }
+}
+\end{lstlisting}
+The construction of the difference density matrix is parallelized by considering the value of the thresholded difference matrix $\boldsymbol{D}'$ at each index $(r, c)$, mapping that index to the appropriate index in the difference density matrix $(i, j)$ where $i < m \wedge j < n$, then adding to the value of the array at $(i, j)$ at each encounter of a difference pixel in $\boldsymbol{D}'$.
+Our CUDA implementation therefore relies on the fact that the density array is initialized to 0 before the kernel is launched; this is accomplished in practice with a \texttt{calloc} operation on the host, followed by copying that array to GPU memory.
+\begin{lstlisting}[
+	caption=GPU computation of the difference density matrix,
+	label=difference-density-implementation
+]
+__global__ void spatial_difference_density_map(
+	double *density_map,
+	int *difference,
+	int width,
+	int height,
+	int horizontal_divisions,
+	int vertical_divisions
+) {
+	int r = blockIdx.y * blockDim.y + threadIdx.y;
+	int c = blockIdx.x * blockDim.x + threadIdx.x;
+	int i = r * width + c;
+
+	int horizontal_block_size = width/horizontal_divisions;
+	int vertical_block_size = height/vertical_divisions;
+	int block_size = horizontal_block_size * vertical_block_size;
+
+	const int scaling_factor = 1000;
+	if (difference[i] != 0) {
+		int i = (int)(vertical_divisions * r/(double)height);
+		int j = (int)(horizontal_divisions * c/(double)width);
+		density_map[i * horizontal_divisions + j] += scaling_factor/(double)block_size;
 	}
@@
 \subsection{Real-time Motion Detection Performance}
+Our real-time motion detection implementation continuously repeats the following procedure in an infinite loop:
+\begin{enumerate}
+	\item Read two frames $\boldsymbol{F}_1$ and $\boldsymbol{F}_2$ from the camera in succession
+	\item Apply a Gaussian low-pass filter of kernel size $k = 3$ to both frames
+	\item Apply a Sobel edge filter in both directions
+	\item Calculate the edges $\boldsymbol{E}_1$ and $\boldsymbol{E}_2$ from the image with non-maximum suppression and selective thresholding
+	\item Calculate the difference matrix $\boldsymbol{D}$ and difference density matrix $\boldsymbol{D}'$
+	\item Draw an image representing the portions of the image with motion
+	\item Update a live preview of $\boldsymbol{F}_1$, $\boldsymbol{E}_1$, $\boldsymbol{D}'$, and the motion area estimate
+\end{enumerate}
+Table \ref{real-time-fps-statistics} shows basic statistics on the frame rate results achieved on a $640 \times 480$ continuous video stream, running on the same hardware as the benchmarks in the previous section.
+The frame rate was evaluated on a 10-second video stream in a well-lit environment with motion of a moderate number of edge pixels ($< 10\%$ of the total number of pixels in the frame) in front of the camera.
+\begin{table}[H]
+	\small
+	\centering
+	\caption{Real-time Frame Rate Statistics}
+	\label{real-time-fps-statistics}
+	\begin{tabular}{ll}
+	 Average FPS & $11.01$ \\
+	 Median FPS & $10.72$ \\
+	 Min FPS & $9.68$ \\
+	 Max FPS & $12.02$ \\
+	 Standard deviation & $0.37$ \\
+	\end{tabular}
+\end{table}
 
@@
 TODO: Some discussion of the results is warranted
+
+\section{Source Code}
+The full source code for this project, including the CUDA and Python implementations of all of our algorithms, is available at \texttt{https://github.com/LINKIWI/elec-301}.
 

### 2015-12-07 cuda/edge-detect/edge_detect.cu
Code cleanup: removed redundant timing statements

@@ -140,13 +140,11 @@ void edge_detect(int *edges_out, int *gx_out, int *gy_out, int *image, int image
 	dim3 grid_size = dim3(bx, by);
 	
 	// Horizontal direction
-//	printf("=====HORIZONTAL PARTIAL DIFFERENTIATION=====\n");
 	int gx_horizontal[3] = {1, 0, -1};
 	int gx_vertical[3] = {1, 2, 1};
 	separable_convolve(gx_out, image, sobel_out_width, sobel_out_height, gx_horizontal, gx_vertical, 3, 1);
 	
 	// Vertical direction
-//	printf("=====VERTICAL PARTIAL DIFFERENTIATION=====\n");
 	int gy_horizontal[3] = {1, 2, 1};
 	int gy_vertical[3] = {1, 0, -1};
 	separable_convolve(gy_out, image, sobel_out_width, sobel_out_height, gy_horizontal, gy_vertical, 3, 1);

### 2015-12-07 cuda/main/main.cu

@@ -68,7 +68,6 @@ int main() {
 		}
 	
 		// Gaussian filter
-//		printf("=====GAUSSIAN FILTER=====\n");
 		int horizontal_filter[3] = {1, 2, 1};
 		int vertical_filter[3] = {1, 2, 1};
 		int kernel_size = 3;
@@ -78,7 +77,7 @@ int main() {
 		int gaussian_out_width = image_1.cols + kernel_size - 1;
 		int gaussian_out_height = image_1.rows + kernel_size - 1;
 		
-		// Edge detect
+		// Edge detection
 		int sobel_out_width = gaussian_out_width + kernel_size - 1;
 		int sobel_out_height = gaussian_out_height + kernel_size - 1;
 		edge_detect(edges_1, gx_out, gy_out, gaussian_out_1, gaussian_out_width, gaussian_out_height, edge_detect_high_threshold, edge_detect_low_threshold);

### 2015-12-07 cuda/motion-tracking/motion_tracking.cu

@@ -166,9 +166,6 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
     checkCudaErrors(cudaMemcpy(dev_edges_1, edges_1, width*height*sizeof(int), cudaMemcpyHostToDevice));
     checkCudaErrors(cudaMemcpy(dev_edges_2, edges_2, width*height*sizeof(int), cudaMemcpyHostToDevice));
 
-//    int *serial_output = (int *)malloc(width * height * sizeof(int));
-//    serial_difference_filter(difference, edges_1, edges_2, width, height, movement_threshold);
-
     // Initialize grid
 	dim3 block_size(TX, TY);
 	int bx = width/block_size.x;
@@ -181,10 +178,8 @@ void motion_detect(int *motion_area, int *difference, int *edges_1, int *edges_2
 
     // Determine spatial density map
 	spatial_difference_density_map<<<grid_size, block_size>>>(dev_density, dev_difference, width, height, horizontal_divisions, vertical_divisions);
-//	serial_spatial_difference_density_map(density_map, difference, width, height, horizontal_divisions, vertical_divisions);
 	
 	// Estimate motion area
-//	serial_motion_area_estimate(motion_area, density_map, width, height, horizontal_divisions, vertical_divisions, motion_threshold);
 	motion_area_estimate<<<grid_size, block_size>>>(dev_motion_area, dev_density, width, height, horizontal_divisions, vertical_divisions, motion_threshold);
 	checkCudaErrors(cudaMemcpy(motion_area, dev_motion_area, width * height * sizeof(int), cudaMemcpyDeviceToHost));
 	

### 2015-12-07 cuda/separable-convolution/separable_convolution.cu

@@ -95,9 +95,6 @@ void separable_convolve(int *output, int *x, int x_width, int x_height, int *hor
     // Allocate space for host and device arrays
     static int *dev_horizontal_out, *dev_vertical_out;  // Results of the horizontal and vertical convolutions on the input array
     static int *dev_horizontal_filter, *dev_vertical_filter, *dev_x;  // Horizontal filter, vertical filter, and input array
-    //int *output = (int *)malloc((x_width * x_height + horizontal_filter_width + vertical_filter_height) * sizeof(int));
-    //int *serial_output = (int *)malloc((x_width * x_height + horizontal_filter_width + vertical_filter_height) * sizeof(int));
-//    static int serial_output[170000000];
 
     // Horizontal filter, followed by vertical filter
     int horizontal_convolution_width = x_width + horizontal_filter_width - 1;
@@ -117,10 +114,6 @@ void separable_convolve(int *output, int *x, int x_width, int x_height, int *hor
     checkCudaErrors(cudaMemcpy(dev_vertical_filter, vertical_filter, vertical_filter_height*sizeof(int), cudaMemcpyHostToDevice));
     checkCudaErrors(cudaMemcpy(dev_x, x, x_width*x_height*sizeof(int), cudaMemcpyHostToDevice));
 
-    // Start with a serial convolution for comparison
-    // double serial_computation_time = serial_convolve(serial_output, x, filter, x_width, x_height, horizontal_filter_width, vertical_filter_height);
-//    double serial_computation_time = serial_separable_convolve(serial_output, x, horizontal_filter, vertical_filter, x_width, x_height, horizontal_filter_width, vertical_filter_height, constant_scalar);
-
     // Initialize grid
     dim3 block_size(TX, TY);
 	int bx_horizontal = horizontal_convolution_width/block_size.x;
@@ -129,10 +122,6 @@ void separable_convolve(int *output, int *x, int x_width, int x_height, int *hor
 	int bx_vertical = vertical_convolution_width/block_size.x;
 	int by_vertical = vertical_convolution_height/block_size.y;
 	dim3 grid_size_vertical = dim3(bx_vertical, by_vertical);
-    
-	// Start a timer and do the two convolutions
-    struct timeval tv1, tv2;
-    gettimeofday(&tv1, NULL);
     
[…]

### 2015-12-07 opencv-object-detect/facedetect.py
removed absolute paths

@@ -1,13 +1,15 @@
 import sys
 import cv2
 import datetime as dt
+import os
 import numpy
 
-cascade = '/Users/Senthil/Documents/elec-301/opencv-object-detect/haarcascade_frontalface_default.xml'
+cascade = 'haarcascade_frontalface_default.xml'
 faceCascade = cv2.CascadeClassifier(cascade)
 
-image1 = '/Users/Senthil/Documents/elec-301/images/birthday1.jpg'
-daftPunk = cv2.imread('/Users/Senthil/Documents/elec-301/opencv-object-detect/DaftPunk.png', cv2.IMREAD_UNCHANGED)
+image1 = os.path.join(os.path.dirname(__file__), 'images/birthday1.jpg')
+
+daftPunk = cv2.imread('DaftPunk.png', cv2.IMREAD_UNCHANGED)
 
 
 def faceDetectWebcam():
@@ -81,5 +83,5 @@ def faceDetectTiming(images):
         cv2.destroyAllWindows()
 
 
-# faceDetectTiming(image1)
-faceDetectWebcam()
+faceDetectTiming(image1)
+#faceDetectWebcam()

### 2015-12-07 opencv-object-detect/facedetect.py
Relative paths for face detection

@@ -7,7 +7,7 @@ import numpy
 cascade = 'haarcascade_frontalface_default.xml'
 faceCascade = cv2.CascadeClassifier(cascade)
 
-image1 = os.path.join(os.path.dirname(__file__), 'images/birthday1.jpg')
+image1 = os.path.join(os.path.dirname(__file__), 'birthday.jpg')
 
 daftPunk = cv2.imread('DaftPunk.png', cv2.IMREAD_UNCHANGED)
 

### 2015-12-07 documents/paper/paper.tex
Motivation and conclusion

@@
 Our goal is to demonstrate the performance of GPU computing by solving a handful of existing problems in computer vision with CUDA: namely, edge detection and motion detection.
+
+\subsection{Motivation}
+GPU parallelization is a powerful tool for accelerating image processing tasks.
+The motivation for our CUDA motion detection algorithm stems from the ease of creating a GPU implementation, due in large part to the simplicity of the CUDA APIs, and the diverse applications of motion detection.
+For example, motion detection is applied in real-time traffic analysis, vision-based security, and consumer technology (e.g.
+motion-triggered home automation solutions).
+Further, since our computer vision algorithm calculates the location of motion in a frame based on detected edges, our approach is both more accurate and supplemented with more specific data.
+Therefore, we view an opportunity use CUDA to explore the speedup possible by using a GPU to accelerate this common task in computer vision.
 
@@
 ]
-__global__ void difference_filter(int *dev_out, int *edges_1, int *edges_2, int width, int height, int threshold) {
+__global__ void difference_filter(
+	int *dev_out,
+	int *edges_1,
+	int *edges_2,
+	int width,
+	int height,
+	int threshold
+) {
 	const int r = blockIdx.y * blockDim.y + threadIdx.y;
@@
 \end{lstlisting}
-The construction of the difference density matrix is parallelized by considering the value of the thresholded difference matrix $\boldsymbol{D}'$ at each index $(r, c)$, mapping that index to the appropriate index in the difference density matrix $(i, j)$ where $i < m \wedge j < n$, then adding to the value of the array at $(i, j)$ at each encounter of a difference pixel in $\boldsymbol{D}'$.
+The construction of the difference density matrix as in Algorithm \ref{difference-density-matrix-algorithm} is parallelized by considering the value of the thresholded difference matrix $\boldsymbol{D}'$ at each index $(r, c)$, mapping that index to the appropriate index in the difference density matrix $(i, j)$ where $i < m \wedge j < n$, then adding to the value of the array at $(i, j)$ at each encounter of a difference pixel in $\boldsymbol{D}'$.
 Our CUDA implementation therefore relies on the fact that the density array is initialized to 0 before the kernel is launched; this is accomplished in practice with a \texttt{calloc} operation on the host, followed by copying that array to GPU memory.
@@
 		density_map[i * horizontal_divisions + j] += scaling_factor/(double)block_size;
+	}
+}
+\end{lstlisting}
+The generation of the motion area estimation image is relatively straightforward: we map the index $(i, j)$ of $\boldsymbol{D}'$ back to an index $(r, c)$ where $r < M \wedge c < N$, and set the value at that index high or low accordingly.
+\begin{lstlisting}[
+	caption=GPU generation of the motion area estimation image,
+	label=motion-area-estimation-implementation
+]
+__global__ void motion_area_estimate(
+	int *motion_area,
+	double *density_map,
+	int width,
+	int height,
+	int horizontal_divisions,
+	int vertical_divisions,
+	double threshold
+) {
+	int r = blockIdx.y * blockDim.y + threadIdx.y;
+	int c = blockIdx.x * blockDim.x + threadIdx.x;
+	int i = r * width + c;
+
+	int density_map_index = (int)(vertical_divisions*r/(double)height) * horizontal_divisions + (int)(horizontal_divisions*c/(double)width);
+
+	if (density_map[density_map_index] >= threshold) {
+		motion_area[i] = 255;
+	} else {
+		motion_area[i] = 0;
 	}
@@
 \end{table}
-\begin{figure}[H]
+\begin{figure}[h]
 	\centering
@@
 The frame rate was evaluated on a 10-second video stream in a well-lit environment with motion of a moderate number of edge pixels ($< 10\%$ of the total number of pixels in the frame) in front of the camera.
+All procedures (as described in this paper) were executed in parallel on the GPU.
 \begin{table}[H]
@@
 \section{Conclusion}
-TODO: Some discussion of the results is warranted
+Our CUDA implementation of both the edge detection and motion detection algorithms demonstrate that parallelized, GPU computation results in significant speedups compared to a serial, CPU implementation.
+In all benchmarked cases (separable convolution with a Gaussian filter, edge detection via non-maximum suppression and selective thresholding, and motion area estimation from a difference density map), we find that the GPU implementation, running on a mid-range graphics card, demonstrated anywhere between a 1.5x to 4.6x speedup over a relatively high-performance, overclocked CPU.
+Further, for each algorithm, we find that the speedup asymptotically approaches a general range as the input size increases.
+\par It is important to note the edge case for the separable convolution algorithm on images of small input sizes (namely, near the dimensions $100 \times 100$, or $10000$ total pixels).
+For this case, we observe a remarkably underwhelming speedup ratio of 0.003482, suggesting that a CPU implementation is nearly 300 times as fast as a GPU implementation (Table \ref{separable-convolution-speedup}).
+We speculate that, because the input size is very small (especially in comparison to the larger versions of the same image, which exceed 55 million total pixels in size), the additional overhead caused by launching the GPU kernel (e.g.
+allocating device memory, copying the host arrays onto the device, and copying the device arrays back to the host following completion of the computation) causes the GPU code to take much longer to run than the more directly implemented CPU code.
+This reasoning is backed by the fact that, at a $250000$ total pixel count (a $500 \times 500$ image), the speedup ($1.740469$) is still much lower than the asymptotic value it approaches as the pixel count exceeds about 1 million ($2.235796$).
+\par However, an opposite trend appears to be in play for the speedup of the non-maximum suppression and selective thresholding algorithm (Table \ref{non-maximum-suppression-selective-thresholding-speedup}), where we observe a high speedup for low input sizes and a lower speedup for high input sizes.
+We speculate that this is most likely due to the overhead of running two separate parallel tasks in sequence, both of which require the same amount of memory.
+As a result of our implementation, not only does a large input size result in a greater percentage of the time that any thread is idle, but it also doubles the total memory footprint of the algorithm.
+This is evident from the significantly higher memory consumption of this procedure compared to the others; an input size of 36 million pixels nearly maxes out the 2 GB of available GPU memory, while the other algorithms consume no more than a few hundred megabytes on the largest benchmarking image we use.
+Despite this trend, the speedup is still greater than 1 for large input sizes, suggesting that a GPU implementation is still consistently faster than its CPU counterpart.
 
-\section{Source Code}
+\section*{Source Code}
 The full source code for this project, including the CUDA and Python implementations of all of our algorithms, is available at \texttt{https://github.com/LINKIWI/elec-301}.

### 2015-12-07 opencv-python/motion_detection.py
motion estimation altorithm implemented

@@ -3,21 +3,50 @@ import cv2
 from matplotlib import pyplot as plt
 import edge_detect as edge
 
-cap = cv2.VideoCapture(0)
+# define constants
+M = 10  # number of columns
+N = 8  # number of rows
+T = 0.01  # Threshold
 
+# start video capture
+cap = cv2.VideoCapture(0)
 if not(cap.isOpened()):
     cap.open()
 
+# initialize last frame and edges of last frame
 last_frame = np.zeros((480, 640), np.uint8)
 last_frame_edge = edge.edge_detect(last_frame)[0]
+
+# motion detection algorithm
+def detect_motion(frame, last_frame):
+    height, width = np.shape(frame)
+    d = last_frame - frame
+    motion = np.zeros(((N, M), np.uint8))
+    for i in range(M):
+        for j in range(N):
+            x = 0
+            for i_prime in range(height/M):
+                for j_prime in range(width/N):
+                    if d[i*M+i_prime, j*N+j_prime] != 0:
+                        x += 1
+            if M*N*x/(height*width*1.0) > T:
+                motion[i, j] = 1
+    return motion
+
+# main loop
 while(cap.isOpened()):
     ret, frame = cap.read()
     frame_edge = edge.edge_detect(frame)[0]
-    # cv2.imshow('frame', frame)  # normal picture
-    cv2.imshow('edge', frame_edge)  # edge detection picture
-    cv2.imshow('motion', frame_edge-last_frame_edge)
+    # cv2.imshow('normal', frame)  # normal video
+    cv2.imshow('edge', frame_edge)  # edge detection video
+    motion = detect_motion(frame_edge, last_frame_edge)
+    cv2.imshow('motion', cv2.resize(motion, (480, 640)))  # motion detection video
+
+    # update frames
     last_frame = frame
     last_frame_edge = frame_edge
+
+    # listen for exit
     if cv2.waitKey(1) & 0xFF == ord('q'):
         break
 

### 2015-12-07 opencv-python/benchmark.py
changed timing for benchmark

@@ -2,11 +2,17 @@ from os import listdir, getcwd
 from os.path import join, isfile
 import edge_detect as edge
 import cv2
+import datetime as dt
 
 images_path = join(getcwd(), '../images')
 images = [f for f in listdir(images_path) if isfile(join(images_path, f))]
 
 for img_path in images:
     img = cv2.imread(join(images_path, img_path), 0)
-    edges, time = edge.edge_detect(img)
+
+    start_time = dt.datetime.now()
+    edges = edge.edge_detect(img)
+    end_time = dt.datetime.now()
+
+    time = (end_time-start_time).microseconds
     print img_path+":", time

### 2015-12-07 opencv-python/edge_detect.py

@@ -1,13 +1,10 @@
 import cv2
 from matplotlib import pyplot as plt
-import datetime as dt
 
 
 def edge_detect(img, display=False):
-    start_time = dt.datetime.now()
+
     edges = cv2.Canny(img, 100, 200)
-    end_time = dt.datetime.now()
-    time = (end_time-start_time).microseconds
 
     if display:
         plt.subplot(121), plt.imshow(img, cmap='gray')
@@ -16,4 +13,4 @@ def edge_detect(img, display=False):
         plt.title('Edge Image'), plt.xticks([]), plt.yticks([])
         plt.show()
 
-    return edges, time
+    return edges

### 2015-12-07 opencv-python/motion_detection.py

@@ -17,10 +17,11 @@ if not(cap.isOpened()):
 last_frame = np.zeros((480, 640), np.uint8)
 last_frame_edge = edge.edge_detect(last_frame)[0]
 
+
 # motion detection algorithm
-def detect_motion(frame, last_frame):
+def detect_motion(edge, last_edge):
     height, width = np.shape(frame)
-    d = last_frame - frame
+    d = last_edge - edge
     motion = np.zeros(((N, M), np.uint8))
     for i in range(M):
         for j in range(N):
@@ -34,10 +35,11 @@ def detect_motion(frame, last_frame):
     return motion
 
 # main loop
-while(cap.isOpened()):
+while cap.isOpened():
     ret, frame = cap.read()
-    frame_edge = edge.edge_detect(frame)[0]
-    # cv2.imshow('normal', frame)  # normal video
+    frame_edge = edge.edge_detect(frame)
+
+    cv2.imshow('normal', frame)  # normal video
     cv2.imshow('edge', frame_edge)  # edge detection video
     motion = detect_motion(frame_edge, last_frame_edge)
     cv2.imshow('motion', cv2.resize(motion, (480, 640)))  # motion detection video

### 2015-12-07 opencv-object-detect/cuda-facedetect.cu
started face recognizer with CUDA

@@ -1,0 +1,23 @@
+#include <stdio.h>
+#include <iostream>
+#include <opencv2/core/core.hpp>
+#include <opencv2/imgproc/imgproc.hpp>
+#include <opencv2/highgui/highgui.hpp>
+#include <opencv2/objdetect/objdetect.hpp>
+
+using namespace cv;
+using namespace std;
+using namespace cv::cuda;
+
+#include <opencv2/gpu/gpu.hpp>
+#include <vector>
+
+int main()
+{
+    printf("\n Hello World\n");
+    return EXIT_SUCCESS;
+}
+
+
+
+

### 2015-12-08 opencv-python/motion_detection.py
motion detection working

@@ -4,9 +4,9 @@ from matplotlib import pyplot as plt
 import edge_detect as edge
 
 # define constants
-M = 10  # number of columns
-N = 8  # number of rows
-T = 0.01  # Threshold
+M = 8  # number of columns
+N = 10  # number of rows
+T = 0.005  # Threshold
 
 # start video capture
 cap = cv2.VideoCapture(0)
@@ -20,18 +20,18 @@ last_frame_edge = edge.edge_detect(last_frame)[0]
 
 # motion detection algorithm
 def detect_motion(edge, last_edge):
-    height, width = np.shape(frame)
+    height, width, depth = np.shape(frame)
     d = last_edge - edge
-    motion = np.zeros(((N, M), np.uint8))
+    motion = np.zeros((M, N), np.uint8)
     for i in range(M):
         for j in range(N):
             x = 0
             for i_prime in range(height/M):
                 for j_prime in range(width/N):
-                    if d[i*M+i_prime, j*N+j_prime] != 0:
+                    if d[i*height/M+i_prime, j*width/N+j_prime] != 0:
                         x += 1
             if M*N*x/(height*width*1.0) > T:
-                motion[i, j] = 1
+                motion[i, j] = 255
     return motion
 
 # main loop
@@ -39,10 +39,10 @@ while cap.isOpened():
     ret, frame = cap.read()
     frame_edge = edge.edge_detect(frame)
 
-    cv2.imshow('normal', frame)  # normal video
+    # cv2.imshow('normal', frame)  # normal video
     cv2.imshow('edge', frame_edge)  # edge detection video
     motion = detect_motion(frame_edge, last_frame_edge)
-    cv2.imshow('motion', cv2.resize(motion, (480, 640)))  # motion detection video
+    cv2.imshow('motion', cv2.resize(motion, (640, 480)))  # motion detection video
 
     # update frames
     last_frame = frame

### 2015-12-08 opencv-object-detect/CudaTest.cu
working with sample CUDA face detect programs

@@ -1,0 +1,323 @@
+#if defined _MSC_VER && _MSC_VER >= 1400
+#pragma warning(disable : 4100)
+#endif
+
+
+#include <iostream>
+#include <iomanip>
+#include "opencv2/objdetect/objdetect.hpp"
+#include "opencv2/highgui/highgui.hpp"
+#include "opencv2/imgproc/imgproc.hpp"
+#include "opencv2/cudaobjdetect.hpp"
+#include "opencv2/cudaimgproc.hpp"
+#include "opencv2/cudawarping.hpp"
+
+//#include "/home/ubuntu/Downloads/opencv/samples/gpu/tickmeter.hpp"
+
+using namespace std;
+using namespace cv;
+using namespace cv::cuda;
+
+
+static void help()
+{
+    cout << "Usage: ./cascadeclassifier_gpu \n\t--cascade <cascade_file>\n\t(<image>|--video <video>|--camera <camera_id>)\n"
+            "Using OpenCV version " << CV_VERSION << endl << endl;
+}
+
+
+static void convertAndResize(const Mat& src, Mat& gray, Mat& resized, double scale)
+{
+    if (src.channels() == 3)
+    {
+        cv::cvtColor( src, gray, COLOR_BGR2GRAY );
+    }
+    else
+    {
+        gray = src;
+    }
+
+    Size sz(cvRound(gray.cols * scale), cvRound(gray.rows * scale));
+
+    if (scale != 1)
+    {
+        cv::resize(gray, resized, sz);
+    }
+    else
+    {
+        resized = gray;
+    }
+}
+
+static void convertAndResize(const GpuMat& src, GpuMat& gray, GpuMat& resized, double scale)
+{
+    if (src.channels() == 3)
+    {
+        cv::cuda::cvtColor( src, gray, COLOR_BGR2GRAY );
+    }
+    else
+    {
+        gray = src;
+    }
+
+    Size sz(cvRound(gray.cols * scale), cvRound(gray.rows * scale));
+
+    if (scale != 1)
+    {
+        cv::cuda::resize(gray, resized, sz);
+    }
+    else
+    {
+        resized = gray;
+    }
+}
+
+
+static void matPrint(Mat &img, int lineOffsY, Scalar fontColor, const string &ss)
+{
+    int fontFace = FONT_HERSHEY_DUPLEX;
+    double fontScale = 0.8;
+    int fontThickness = 2;
[…]

### 2015-12-08 opencv-object-detect/CudaTest2.cu

@@ -1,0 +1,40 @@
+#include <opencv2/opencv.hpp>
+#include <opencv2/cudaobjdetect.hpp>
+#include <iostream>
+
+using namespace std;
+using namespace cv;
+
+int main(int argc, char* argv[])
+{
+    Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml");
+
+    if (debugCascade.empty()){
+        cerr << "Could not load model!" << endl;
+        return -1;
+    }
+
+    Mat original = imread("/home/ubuntu/project/opencv-object-detect/birthday.jpg");
+    Mat gray = imread("/home/ubuntu/project/opencv-object-detect/birthday.jpg", CV_LOAD_IMAGE_GRAYSCALE);
+    Mat work_img = original.clone();
+
+    // Preprocessing
+    cvtColor(original, work_img, COLOR_RGB2GRAY);
+    equalizeHist(work_img, work_img);
+
+    // Detection
+    cuda::GpuMat inputGPU(work_img);
+    cuda::GpuMat faces;
+    debugCascade->detectMultiScale(inputGPU, faces);
+    vector<Rect> objects;
+    debugCascade->convert(faces, objects);
+
+    for(int i=0;i<(int)objects.size();++i)
+    {
+         rectangle(original, objects[i], Scalar(255));
+    }  
+
+    imshow("detections", original);
+    moveWindow("detections", 100, 100);
+    waitKey(0);
+}

### 2015-12-08 opencv-object-detect/cuda-facedetect.cu

@@ -9,15 +9,99 @@ using namespace cv;
 using namespace std;
 using namespace cv::cuda;
 
+#include <opencv2/cudaobjdetect.hpp>
 #include <opencv2/gpu/gpu.hpp>
 #include <vector>
+#include "fstream"
+#include <opencv2/ml/ml.hpp>
+#include <opencv2/core/gpumat.hpp>
+#include <opencv2/video/tracking.hpp>
+#include <opencv2/core/cuda.hpp>
 
-int main()
+//cv::cuda::CascadeClassifier_CUDA cascade_gpu;
+
+Ptr<cuda::CascadeClassifier> cascade_gpu = cuda::CascadeClassifier::create("haarcascade_frontalface_default.xml");
+
+//vector<Rect> detect_faces(Mat& image)
+void detect_faces(Mat& image)
 {
-    printf("\n Hello World\n");
-    return EXIT_SUCCESS;
+	vector<Rect> res;
+	bool findLargestObject = true;
+	bool filterRects = true;
+	int detections_num;
+	Mat faces_downloaded;
+	Mat im(image.size(), CV_8UC1);
+	GpuMat facesBuf_gpu;
+	if(image.channels()==3)
+	{
+		cvtColor(image, im, CV_BGR2GRAY);
+	}
+	else
+	{
+		image.copyTo(im);
+	}
+	GpuMat gray_gpu(im);
+	
+//	cascade_gpu.visualizeInPlace = false;
+//	findLargestObject=cascade_gpu->FindLargestObject;
+//	detections_num = cascade_gpu->detectMultiScale(gray_gpu, facesBuf_gpu, 1.2, (filterRects || findLargestObject) ? 4 : 0,Size(image.cols/4,image.rows/4));
+//	detections_num = cascade_gpu->detectMultiScale(gray_gpu, facesBuf_gpu);
+//	if(detections_num==0){return res;}
+	
+//	facesBuf_gpu.colRange(0, detections_num).download(faces_downloaded);
+//	Rect *faceRects = faces_downloaded.ptr<Rect>();
+
+//	for(int i = 0; i<detections_num; i++)
+//	{
+//		res.push_back(faceRects[i]);
+//	}
+	
+	cascade_gpu->detectMultiScale(gray_gpu, facesBuf_gpu);
+	std::vector<Rect> faces;
+	cascade_gpu->convert(facesBuf_gpu, faces);
+
[…]

### 2015-12-08 opencv-python/motion_detection.py
fully implemented motin detection

@@ -4,24 +4,43 @@ from matplotlib import pyplot as plt
 import edge_detect as edge
 
 # define constants
-M = 8  # number of columns
+M = 6  # number of columns
 N = 10  # number of rows
-T = 0.005  # Threshold
+T = 0.01  # Threshold
+B = 12  # movement tolerance
+
 
 # start video capture
 cap = cv2.VideoCapture(0)
 if not(cap.isOpened()):
     cap.open()
 
+# create gaussian filter size 3x3
+kernel = np.ones((3, 3), np.float32)/9
+
 # initialize last frame and edges of last frame
-last_frame = np.zeros((480, 640), np.uint8)
-last_frame_edge = edge.edge_detect(last_frame)[0]
+last_frame_edge = np.zeros((480, 640), np.uint8)
+
+
+def difference(frame1, frame2):
+    height, width = np.shape(frame1)
+    diff = np.zeros((height, width), np.uint8)
+    for i in range(height):
+        for j in range(width):
+            if frame1[i, j] != frame2[i, j]:
+                diff[i, j] = 255
+                for i_prime in range(-B, B):
+                    for j_prime in range(-B, B):
+                        if i+i_prime < height and i+i_prime >= 0 and j+j_prime < width and j+j_prime >= 0:
+                            if frame1[i+i_prime, j+j_prime] == frame2[i+i_prime, j+j_prime]:
+                                diff[i+i_prime, j+j_prime] = 0
+    return diff
 
 
 # motion detection algorithm
 def detect_motion(edge, last_edge):
     height, width, depth = np.shape(frame)
-    d = last_edge - edge
+    d = difference(edge, last_edge)
     motion = np.zeros((M, N), np.uint8)
     for i in range(M):
         for j in range(N):
@@ -37,15 +56,16 @@ def detect_motion(edge, last_edge):
 # main loop
 while cap.isOpened():
     ret, frame = cap.read()
-    frame_edge = edge.edge_detect(frame)
+    frame_filt = cv2.filter2D(frame, -1, kernel)
+    frame_edge = edge.edge_detect(frame_filt)
 
     # cv2.imshow('normal', frame)  # normal video
-    cv2.imshow('edge', frame_edge)  # edge detection video
+    cv2.imshow('edge', frame_edge)  # edge detection video)
     motion = detect_motion(frame_edge, last_frame_edge)
     cv2.imshow('motion', cv2.resize(motion, (640, 480)))  # motion detection video
[…]

### 2015-12-08 opencv-object-detect/CudaTest.cu
CudaTests are working

@@ -236,7 +236,7 @@ int main(int argc, const char *argv[])
         if (useGPU)
         {
             cascade_gpu->setFindLargestObject(findLargestObject);
-            cascade_gpu->setScaleFactor(1.2);
+            cascade_gpu->setScaleFactor(1.1);
             cascade_gpu->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
 
             cascade_gpu->detectMultiScale(resized_gpu, facesBuf_gpu);
@@ -245,7 +245,7 @@ int main(int argc, const char *argv[])
         else
         {
             Size minSize = cascade_gpu->getClassifierSize();
-            cascade_cpu.detectMultiScale(resized_cpu, faces, 1.2,
+            cascade_cpu.detectMultiScale(resized_cpu, faces, 1.1,
                                          (filterRects || findLargestObject) ? 4 : 0,
                                          (findLargestObject ? CASCADE_FIND_BIGGEST_OBJECT : 0)
                                             | CASCADE_SCALE_IMAGE,

### 2015-12-08 opencv-object-detect/CudaTest2.cu

@@ -7,7 +7,8 @@ using namespace cv;
 
 int main(int argc, char* argv[])
 {
-    Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml");
+//    Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/project/opencv-object-detect/haarcascade_frontalface_default.xml");
+	Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml");
 
     if (debugCascade.empty()){
         cerr << "Could not load model!" << endl;

### 2015-12-08 opencv-object-detect/CudaTest.cu
started timing benchmarks

@@ -11,6 +11,8 @@
 #include "opencv2/cudaobjdetect.hpp"
 #include "opencv2/cudaimgproc.hpp"
 #include "opencv2/cudawarping.hpp"
+#include <stdio.h>
+#include <sys/time.h>
 
 //#include "/home/ubuntu/Downloads/opencv/samples/gpu/tickmeter.hpp"
 
@@ -208,6 +210,7 @@ int main(int argc, const char *argv[])
 
     /* parameters */
     bool useGPU = true;
+   // bool useGPU = false;
     double scaleFactor = 1.0;
     bool findLargestObject = false;
     bool filterRects = true;
@@ -232,6 +235,9 @@ int main(int argc, const char *argv[])
 
 //        TickMeter tm;
   //      tm.start();
+	
+	struct timeval tstart, tend;
+	gettimeofday(&tstart, NULL);	
 
         if (useGPU)
         {
@@ -256,6 +262,10 @@ int main(int argc, const char *argv[])
         {
             rectangle(resized_cpu, faces[i], Scalar(255));
         }
+	
+	gettimeofday(&tend, NULL);
+	double runtime = (double) (tend.tv_usec - tstart.tv_usec) / 100000 + (double) (tend.tv_sec - tstart.tv_sec);
+	printf("Runtime: %f seconds\n", runtime);	
 
     //    tm.stop();
       //  double detectionTime = tm.getTimeMilli();

### 2015-12-09 opencv-object-detect/CudaTest2.cu
added timing for CudaTest.cu

@@ -1,13 +1,13 @@
 #include <opencv2/opencv.hpp>
 #include <opencv2/cudaobjdetect.hpp>
 #include <iostream>
+#include <sys/time.h>
 
 using namespace std;
 using namespace cv;
 
 int main(int argc, char* argv[])
 {
-//    Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/project/opencv-object-detect/haarcascade_frontalface_default.xml");
 	Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml");
 
     if (debugCascade.empty()){
@@ -15,24 +15,40 @@ int main(int argc, char* argv[])
         return -1;
     }
 
-    Mat original = imread("/home/ubuntu/project/opencv-object-detect/birthday.jpg");
-    Mat gray = imread("/home/ubuntu/project/opencv-object-detect/birthday.jpg", CV_LOAD_IMAGE_GRAYSCALE);
+    Mat original = imread("/home/ubuntu/project/opencv-object-detect/senthil1000.jpg");
+    Mat gray = imread("/home/ubuntu/project/opencv-object-detect/senthil1000.jpg", CV_LOAD_IMAGE_GRAYSCALE);
     Mat work_img = original.clone();
 
     // Preprocessing
     cvtColor(original, work_img, COLOR_RGB2GRAY);
     equalizeHist(work_img, work_img);
 
+    bool findLargestObject = false;
+    bool filterRects = true;
+
+    struct timeval tstart, tend;
+    gettimeofday(&tstart, NULL);	
+    
     // Detection
     cuda::GpuMat inputGPU(work_img);
     cuda::GpuMat faces;
+    debugCascade->setFindLargestObject(findLargestObject);
+    debugCascade->setScaleFactor(1.2);
+    debugCascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
     debugCascade->detectMultiScale(inputGPU, faces);
+
+    gettimeofday(&tend, NULL);
+    double runtime = (double) (tend.tv_usec - tstart.tv_usec) / 100000 + (double) (tend.tv_sec - tstart.tv_sec);
+    printf("GPU Runtime: %f seconds\n", runtime);
+    printf("end %f \n", (double)(tend.tv_usec));
[…]

### 2015-12-09 opencv-object-detect/CudaTest2.cu
added cpu implementation and timing

@@ -2,15 +2,25 @@
 #include <opencv2/cudaobjdetect.hpp>
 #include <iostream>
 #include <sys/time.h>
+#include "opencv2/objdetect/objdetect.hpp"
+#include "opencv2/highgui/highgui.hpp"
+#include "opencv2/imgproc/imgproc.hpp"
+#include "opencv2/cudaimgproc.hpp"
+#include "opencv2/cudawarping.hpp"
+#include <iomanip>
 
 using namespace std;
 using namespace cv;
+//using namespace cv::cuda;
 
 int main(int argc, char* argv[])
 {
-	Ptr<cuda::CascadeClassifier> debugCascade = cuda::CascadeClassifier::create("/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml");
+	string cascadeName = "/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml";
+	Ptr<cuda::CascadeClassifier> cascade = cuda::CascadeClassifier::create(cascadeName);
+	cv::CascadeClassifier cascadeCPU;
+	cascadeCPU.load(cascadeName);
 
-    if (debugCascade.empty()){
+    if (cascade.empty() || cascadeCPU.empty()){
         cerr << "Could not load model!" << endl;
         return -1;
     }
@@ -26,26 +36,46 @@ int main(int argc, char* argv[])
     bool findLargestObject = false;
     bool filterRects = true;
 
-    struct timeval tstart, tend;
-    gettimeofday(&tstart, NULL);	
+    struct timeval tstartC, tendC, tstartG, tendG;	
     
-    // Detection
+    // GPU Detection
+    gettimeofday(&tstartG, NULL);
     cuda::GpuMat inputGPU(work_img);
     cuda::GpuMat faces;
-    debugCascade->setFindLargestObject(findLargestObject);
-    debugCascade->setScaleFactor(1.2);
-    debugCascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
-    debugCascade->detectMultiScale(inputGPU, faces);
+    cascade->setFindLargestObject(findLargestObject);
+    cascade->setScaleFactor(1.2);
+    cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
+    cascade->detectMultiScale(inputGPU, faces);
 
-    gettimeofday(&tend, NULL);
[…]

### 2015-12-09 opencv-object-detect/CudaTest2.cu
tweaked timing parameters

@@ -25,8 +25,8 @@ int main(int argc, char* argv[])
         return -1;
     }
 
-    Mat original = imread("/home/ubuntu/project/opencv-object-detect/senthil1000.jpg");
-    Mat gray = imread("/home/ubuntu/project/opencv-object-detect/senthil1000.jpg", CV_LOAD_IMAGE_GRAYSCALE);
+    Mat original = imread("/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg");
+    Mat gray = imread("/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg", CV_LOAD_IMAGE_GRAYSCALE);
     Mat work_img = original.clone();
 
     // Preprocessing
@@ -47,14 +47,25 @@ int main(int argc, char* argv[])
     cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
     cascade->detectMultiScale(inputGPU, faces);
 
-    gettimeofday(&tendG, NULL);
-    double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
-    printf("GPU Runtime: %f seconds\n", runtimeG);
+//    gettimeofday(&tendG, NULL);
+//    double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
+//    printf("GPU Runtime: %f seconds\n", runtimeG);
     //printf("end %f \n", (double)(tend.tv_usec));
     //printf("start %f \n", (double)(tstart.tv_usec));
 
     vector<Rect> objects;
     cascade->convert(faces, objects);
+   
+    for(int i=0;i<(int)objects.size();++i)
+    {
+         rectangle(original, objects[i], Scalar(0, 255, 0), 3);
+    }  
+  
+    imshow("detections", original);
+
+    gettimeofday(&tendG, NULL);
+    double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
+    printf("GPU Runtime: %f seconds\n", runtimeG);
 
     //CPU Detection
     gettimeofday(&tstartC, NULL);
@@ -66,8 +77,16 @@ int main(int argc, char* argv[])
                                          (filterRects || findLargestObject) ? 4 : 0,
                                          (findLargestObject ? CASCADE_FIND_BIGGEST_OBJECT : 0)
[…]

### 2015-12-09 opencv-object-detect/CudaTest2.cu
cleaned up path refereces

@@ -25,8 +25,9 @@ int main(int argc, char* argv[])
         return -1;
     }
 
-    Mat original = imread("/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg");
-    Mat gray = imread("/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg", CV_LOAD_IMAGE_GRAYSCALE);
+    string imageName = "/home/ubuntu/project/opencv-object-detect/senthil1000.jpg";
+    Mat original = imread(imageName);
+    Mat gray = imread(imageName, CV_LOAD_IMAGE_GRAYSCALE);
     Mat work_img = original.clone();
 
     // Preprocessing

### 2015-12-09 opencv-object-detect/CudaTest2.cu
resave jic

@@ -25,7 +25,7 @@ int main(int argc, char* argv[])
         return -1;
     }
 
-    string imageName = "/home/ubuntu/project/opencv-object-detect/senthil1000.jpg";
+    string imageName = "/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg";
     Mat original = imread(imageName);
     Mat gray = imread(imageName, CV_LOAD_IMAGE_GRAYSCALE);
     Mat work_img = original.clone();
@@ -48,9 +48,9 @@ int main(int argc, char* argv[])
     cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
     cascade->detectMultiScale(inputGPU, faces);
 
-//    gettimeofday(&tendG, NULL);
-//    double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
-//    printf("GPU Runtime: %f seconds\n", runtimeG);
+   // gettimeofday(&tendG, NULL);
+   // double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
+   // printf("GPU Runtime: %f seconds\n", runtimeG);
     //printf("end %f \n", (double)(tend.tv_usec));
     //printf("start %f \n", (double)(tstart.tv_usec));
 
@@ -71,7 +71,6 @@ int main(int argc, char* argv[])
     //CPU Detection
     gettimeofday(&tstartC, NULL);
     cv::Mat inputCPU(work_img);
-    //cv::Mat facesCPU;
     vector<Rect> facesCPU;
     Size minsize = cascade->getClassifierSize();
     cascadeCPU.detectMultiScale(inputCPU, facesCPU, 1.1,

### 2015-12-09 opencv-object-detect/CudaTest2.cu
made more modular, added command line arguments

@@ -11,7 +11,6 @@
 
 using namespace std;
 using namespace cv;
-//using namespace cv::cuda;
 
 int main(int argc, char* argv[])
 {
@@ -25,7 +24,8 @@ int main(int argc, char* argv[])
         return -1;
     }
 
-    string imageName = "/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg";
+    //string imageName = "/home/ubuntu/project/opencv-object-detect/Senthil2000.jpg";
+    string imageName = string(argv[2]);
     Mat original = imread(imageName);
     Mat gray = imread(imageName, CV_LOAD_IMAGE_GRAYSCALE);
     Mat work_img = original.clone();
@@ -38,15 +38,75 @@ int main(int argc, char* argv[])
     bool filterRects = true;
 
     struct timeval tstartC, tendC, tstartG, tendG;	
-    
+  
+    if (string(argv[1]) == "--GPU")
+	{
+
+		gettimeofday(&tstartG, NULL);
+    		cuda::GpuMat inputGPU(work_img);
+    		cuda::GpuMat faces;
+    		cascade->setFindLargestObject(findLargestObject);
+    		cascade->setScaleFactor(1.2);
+    		cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
+    		cascade->detectMultiScale(inputGPU, faces);
+
+   		// gettimeofday(&tendG, NULL);
+   		// double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
+   		// printf("GPU Runtime: %f seconds\n", runtimeG);
+   		 //printf("end %f \n", (double)(tend.tv_usec));
+   		 //printf("start %f \n", (double)(tstart.tv_usec));
+	
+    		vector<Rect> objects;
+    		cascade->convert(faces, objects);
+	
+    		for(int i=0;i<(int)objects.size();++i)
+    		{
+    		     rectangle(original, objects[i], Scalar(0, 255, 0), 3);
+    		}
+
+    		imshow("detections", original);
+
+    		gettimeofday(&tendG, NULL);
+    		double runtimeG = (double) (tendG.tv_usec - tstartG.tv_usec) / 100000 + (double) (tendG.tv_sec - tstartG.tv_sec);
+    		printf("GPU Runtime: %f seconds\n", runtimeG);
+
+	}
+
[…]

### 2015-12-09 opencv-object-detect/CudaTest2.cu
Benchmarking code for facial recognition

@@ -8,27 +8,229 @@
 #include "opencv2/cudaimgproc.hpp"
 #include "opencv2/cudawarping.hpp"
 #include <iomanip>
+#include <cuda_runtime.h>
 
 using namespace std;
 using namespace cv;
 
+//global call of cascade
+
+void recog(string imageName, string mode)
+{
+
+	string cascadeName = "/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml";
+        Ptr<cuda::CascadeClassifier> cascade = cuda::CascadeClassifier::create(cascadeName);
+        cv::CascadeClassifier cascadeCPU;
+        cascadeCPU.load(cascadeName);
+	
+	if (cascade.empty() || cascadeCPU.empty()){
+                cerr << "Could not load model!" << endl;
+               // return -1;
+        }
+	
+	Mat original = imread(imageName);
+    	Mat gray = imread(imageName, CV_LOAD_IMAGE_GRAYSCALE);
+    	Mat work_img = original.clone();
+    	Mat copy = original.clone();
+
+	//pre processing
+	cvtColor(original, work_img, COLOR_RGB2GRAY);
+    	equalizeHist(work_img, work_img);
+
+    	bool findLargestObject = false;
+    	bool filterRects = true;
+
+	if (mode == "--GPU")
+        {
+
+                cuda::GpuMat inputGPU(work_img);
+                cuda::GpuMat faces;
+		faces.create(1, 100000, cv::DataType<cv::Rect>::type); ///
+                cascade->setFindLargestObject(findLargestObject);
+                cascade->setScaleFactor(1.2);
+                cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
+                cascade->detectMultiScale(inputGPU, faces);
+
+                vector<Rect> objects;
+                cascade->convert(faces, objects);
+
+                for(int i=0;i<(int)objects.size();++i)
+                {
+                     rectangle(original, objects[i], Scalar(0, 255, 0), 3);
+                }
+
+                imshow("detections", original);
+		//moveWindow("detections", 100, 100);
+    		//waitKey(0);
+        }
+
+	else if (mode == "--CPU")
+        {
+                
+                cv::Mat inputCPU(work_img);
+                vector<Rect> facesCPU;
[…]

### 2015-12-09 opencv-object-detect/CudaTestKevin.cu

@@ -1,0 +1,239 @@
+#include <opencv2/opencv.hpp>
+#include <opencv2/cudaobjdetect.hpp>
+#include <iostream>
+#include <sys/time.h>
+#include "opencv2/objdetect/objdetect.hpp"
+#include "opencv2/highgui/highgui.hpp"
+#include "opencv2/imgproc/imgproc.hpp"
+#include "opencv2/cudaimgproc.hpp"
+#include "opencv2/cudawarping.hpp"
+#include <iomanip>
+#include <cuda_runtime.h>
+
+using namespace std;
+using namespace cv;
+
+//global call of cascade
+
+void recog(string imageName, string mode)
+{
+	//string cascadeName = "/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml";
+        string cascadeName = "haarcascade_frontalface_alt2.xml";
+        Ptr<cuda::CascadeClassifier> cascade = cuda::CascadeClassifier::create(cascadeName);
+        cv::CascadeClassifier cascadeCPU;
+        cascadeCPU.load(cascadeName);
+
+        if (cascade.empty() || cascadeCPU.empty()){
+                cerr << "Could not load model!" << endl;
+               // return -1;
+        }
+
+        Mat original = imread(imageName);
+        Mat gray = imread(imageName, CV_LOAD_IMAGE_GRAYSCALE);
+        Mat work_img = original.clone();
+        Mat copy = original.clone();
+
+        //pre processing
+        cvtColor(original, work_img, COLOR_RGB2GRAY);
+        equalizeHist(work_img, work_img);
+
+        bool findLargestObject = false;
+        bool filterRects = true;
+
+	if (mode == "--GPU")
+        {
+
+                cuda::GpuMat inputGPU(work_img);
+                cuda::GpuMat faces;
+                faces.create(10000000, 10000000, cv::DataType<cv::Rect>::type); ///
+                cascade->setFindLargestObject(findLargestObject);
+                cascade->setScaleFactor(1.2);
+                cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
+                cascade->detectMultiScale(inputGPU, faces);
+
+                vector<Rect> objects;
+                cascade->convert(faces, objects);
+
+                for(int i=0;i<(int)objects.size();++i)
+                {
+                     rectangle(original, objects[i], Scalar(0, 255, 0), 3);
+                }
+
+                imshow("detections", original);
[…]

### 2015-12-11 documents/paper/paper.tex
Addressed error in algorithm
corrected the spelling of my name (because Kevin is not a bro)
added python section into intro

@@
 Our goal is to demonstrate the performance of GPU computing by solving a handful of existing problems in computer vision with CUDA: namely, edge detection and motion detection.
+
+\subsection{OpenCV and Python}
+While CUDA is an excelent language for maximizing performance, it could be infeasible due to inexperience with C, time constraints, or lack of an NVIDIA GPU.
+To address these constraints, we also implemented our algorithms in Python using OpenCV.
+Python was chosen because it is very easy for inexperienced programmers to use and OpenCV was chosen because it offers many common image processing algorithms in a variety of languages.
 
@@
 			\nl $x \gets 0$\tcp*{Number of differences in this region}
-			\nl \ForEach{$i' \in [0, \frac{M}{n}]$} {
+			\nl \ForEach{$i' \in [0, \frac{M}{m}]$} {
 				\nl \ForEach{$j' \in [0, \frac{N}{n}]$} {
-					\nl \If {$\boldsymbol{D}_{i + i', j + j'} \neq 0$} {
+					\nl \If {$\boldsymbol{D}_{\frac{Mi}{m} + i', \frac{Nj}{n} + j'} \neq 0$} {
 						\nl $x \gets x + 1$\;
@@
 \end{lstlisting}
+
+\subsection{High Level Python Implementation}
+Python is a widely used big-data programming language because of its extensive library support in such areas and its ease of use.
 
@@
 	\item \textbf{Kevin Lin}, Electrical and Computer Engineering `18
-	\item \textbf{Senthil Natarjan}, Electrical and Computer Engineering `17
+	\item \textbf{Senthil Natarajan}, Electrical and Computer Engineering `17
 \end{itemize}

### 2015-12-11 opencv-python/motion_detection.py

@@ -29,8 +29,8 @@ def difference(frame1, frame2):
         for j in range(width):
             if frame1[i, j] != frame2[i, j]:
                 diff[i, j] = 255
-                for i_prime in range(-B, B):
-                    for j_prime in range(-B, B):
+                for i_prime in range(-B, B+1):
+                    for j_prime in range(-B, B+1):
                         if i+i_prime < height and i+i_prime >= 0 and j+j_prime < width and j+j_prime >= 0:
                             if frame1[i+i_prime, j+j_prime] == frame2[i+i_prime, j+j_prime]:
                                 diff[i+i_prime, j+j_prime] = 0

### 2015-12-11 documents/paper/speedup_graphs.m
Update speedup graphs + MATLAB script for plots

@@ -1,0 +1,17 @@
+pixel_count = [100^2, 500^2, 1000^2, 2000^2, 3000^2, 4000^2, 5000^2, 6000^2, 7500^2]./10^6;
+separable_convolution_speedup = [0.001521 2.306766 2.932381 3.238829 3.319243 3.375912 3.480884 3.503627 3.509777];
+edge_classification_speedup = [97.544669 10.350657 9.318748 7.847971 6.959467 6.129282 5.709230 5.230187, -1];
+motion_area_estimation_speedup = [5.084746 7.820414 8.895548 8.923449 9.547545 9.456960 8.677697 8.805193 4.264629];
+
+figure, scatter(pixel_count, separable_convolution_speedup);
+xlabel('Pixel count (millions of pixels)');
+ylabel('Speedup');
+ylim([1, 12]);
+figure, scatter(pixel_count, edge_classification_speedup);
+xlabel('Pixel count (millions of pixels)');
+ylabel('Speedup');
+ylim([1, 12]);
+figure, scatter(pixel_count, motion_area_estimation_speedup);
+xlabel('Pixel count (millions of pixels)');
+ylabel('Speedup');
+ylim([1, 12]);

### 2015-12-11 documents/paper/paper.tex
Update speedup numbers, title, acknowledgements, add facial recognition section

@@
 
-\title{GPU-Accelerated Parallelization of Edge-Based Motion Detection Algorithms with the NVIDIA CUDA Framework}
+\title{GPU Acceleration of Machine Learning-Aided Facial Recognition and Edge-Based Motion Detection with NVIDIA CUDA}
 
@@
 Therefore, we view an opportunity use CUDA to explore the speedup possible by using a GPU to accelerate this common task in computer vision.
+
+\section{Facial Recognition Algorithm}
+Placeholder text
+
+\subsection{Placeholder}
+Placeholder text
+
+\subsection{Placeholder}
+Placeholder text
+
+\subsection{GPU Parallelization and CUDA Implementation}
+Placeholder text
 
@@
 \begin{itemize}
-	\item Intel Core i7 4790K @ 4.7 GHz (4C, 8T)
-	\item 16 GB of DDR3 memory @ 1600 MHz
-	\item NVIDIA GeForce GTX 750 (512 CUDA cores, 1653 MHz core clock, 6510 MHz memory clock, 2 GB video memory)
+	\item Intel Core i5 4200U @ 2.6 GHz (2C, 4T)
+	\item 8 GB of DDR3 memory @ 1600 MHz
+	\item NVIDIA GeForce GTX 860M (1152 CUDA cores, 1029 MHz core clock, 2500 MHz memory clock, 2 GB GGDR5 video memory)
 \end{itemize}
@@
 	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
-	 $100 \times 100$ & $0.000191$ & $0.054855$ & $0.003482$ \\
-	 $500 \times 500$ & $0.004748$ & $0.002728$ & $1.740469$ \\
-	 $1000 \times 1000$ & $0.018850$ & $0.008431$ & $2.235796$ \\
-	 $2000 \times 2000$ & $0.074475$ & $0.034503$ & $2.158508$ \\
-	 $3000 \times 3000$ & $0.167459$ & $0.072181$ & $2.319987$ \\
-	 $4000 \times 4000$ & $0.299840$ & $0.136641$ & $2.194363$ \\
-	 $5000 \times 5000$ & $0.465225$ & $0.194512$ & $2.391755$ \\
-	 $6000 \times 6000$ & $0.673266$ & $0.276730$ & $2.432935$ \\
-	 $7500 \times 7500$ & $1.052294$ & $0.439262$ & $2.395595$ \\
+	 $100 \times 100$ & $0.000349$ & $0.229505$ & $0.001521$ \\
+	 $500 \times 500$ & $0.009001$ & $0.003902$ & $2.306766$ \\
+	 $1000 \times 1000$ & $0.034910$ & $0.011905$ & $2.932381$ \\
+	 $2000 \times 2000$ & $0.138583$ & $0.042788$ & $3.238829$ \\
+	 $3000 \times 3000$ & $0.311335$ & $0.093797$ & $3.319243$ \\
+	 $4000 \times 4000$ & $0.559149$ & $0.165629$ & $3.375912$ \\
+	 $5000 \times 5000$ & $0.864850$ & $0.248457$ & $3.480884$ \\
+	 $6000 \times 6000$ & $1.253738$ & $0.357840$ & $3.503627$ \\
+	 $7500 \times 7500$ & $1.960140$ & $0.558480$ & $3.509777$ \\
 	\end{tabular}
@@
 	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
-	 $100 \times 100$ & $0.001701$ & $0.000332$ & $5.123494$ \\
-	 $500 \times 500$ & $0.027121$ & $0.009488$ & $2.858453$ \\
-	 $1000 \times 1000$ & $0.083744$ & $0.028291$ & $2.960093$ \\
-	 $2000 \times 2000$ & $0.237126$ & $0.104160$ & $2.276555$ \\
-	 $3000 \times 3000$ & $0.436223$ & $0.207502$ & $2.102259$ \\
-	 $4000 \times 4000$ & $0.608094$ & $0.341184$ & $1.782305$ \\
-	 $5000 \times 5000$ & $0.823099$ & $0.484530$ & $1.698758$ \\
-	 $6000 \times 6000$ & $1.020302$ & $0.668001$ & $1.527396$ \\
+	 $100 \times 100$ & $0.033848$ & $0.000347$ & $97.544669$ \\
+	 $500 \times 500$ & $0.070105$ & $0.006773$ & $10.350657$ \\
+	 $1000 \times 1000$ & $0.213390$ & $0.022899$ & $9.318748$ \\
+	 $2000 \times 2000$ & $0.616050$ & $0.078498$ & $7.847971$ \\
+	 $3000 \times 3000$ & $1.114336$ & $0.160118$ & $6.959467$ \\
+	 $4000 \times 4000$ & $1.627116$ & $0.265466$ & $6.129282$ \\
+	 $5000 \times 5000$ & $2.215267$ & $0.388015$ & $5.709230$ \\
+	 $6000 \times 6000$ & $2.764201$ & $0.528509$ & $5.230187$ \\
 	\end{tabular}
@@
 	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
-	 $100 \times 100$ & $0.004406$ & $0.002055$ & $2.144039$ \\
-	 $500 \times 500$ & $0.116421$ & $0.028079$ & $4.146195$ \\
-	 $1000 \times 1000$ & $0.490470$ & $0.107598$ & $4.558356$ \\
-	 $2000 \times 2000$ & $1.863727$ & $0.424997$ & $4.385271$ \\
-	 $3000 \times 3000$ & $4.334172$ & $0.952842$ & $4.548679$ \\
-	 $4000 \times 4000$ & $7.730778$ & $1.778392$ & $4.347061$ \\
-	 $5000 \times 5000$ & $12.174515$ & $2.660347$ & $4.576288$ \\
-	 $6000 \times 6000$ & $17.628128$ & $3.865616$ & $4.560238$ \\
-	 $7500 \times 7500$ & $28.354028$ & $6.118255$ & $4.634333$ \\
+	 $100 \times 100$ & $0.008400$ & $0.001652$ & $5.084746$ \\
+	 $500 \times 500$ & $0.214381$ & $0.027413$ & $7.820414$ \\
+	 $1000 \times 1000$ & $0.866711$ & $0.097432$ & $8.895548$ \\
+	 $2000 \times 2000$ & $3.468732$ & $0.388721$ & $8.923449$ \\
+	 $3000 \times 3000$ & $8.320342$ & $0.871464$ & $9.547545$ \\
+	 $4000 \times 4000$ & $14.841659$ & $1.569390$ & $9.456960$ \\
+	 $5000 \times 5000$ & $22.094181$ & $2.546088$ & $8.677697$ \\
+	 $6000 \times 6000$ & $31.911879$ & $3.624211$ & $8.805193$ \\
+	 $7500 \times 7500$ & $55.772766$ & $13.077988$ & $4.264629$ \\
 	\end{tabular}
@@
 \par It is important to note the edge case for the separable convolution algorithm on images of small input sizes (namely, near the dimensions $100 \times 100$, or $10000$ total pixels).
-For this case, we observe a remarkably underwhelming speedup ratio of 0.003482, suggesting that a CPU implementation is nearly 300 times as fast as a GPU implementation (Table \ref{separable-convolution-speedup}).
+For this case, we observe a remarkably underwhelming speedup ratio of 0.001521, suggesting that a CPU implementation is nearly 700 times as fast as a GPU implementation (Table \ref{separable-convolution-speedup}).
 We speculate that, because the input size is very small (especially in comparison to the larger versions of the same image, which exceed 55 million total pixels in size), the additional overhead caused by launching the GPU kernel (e.g.
 allocating device memory, copying the host arrays onto the device, and copying the device arrays back to the host following completion of the computation) causes the GPU code to take much longer to run than the more directly implemented CPU code.
-This reasoning is backed by the fact that, at a $250000$ total pixel count (a $500 \times 500$ image), the speedup ($1.740469$) is still much lower than the asymptotic value it approaches as the pixel count exceeds about 1 million ($2.235796$).
+This reasoning is backed by the fact that, at a $250000$ total pixel count (a $500 \times 500$ image), the speedup ($2.306766$) is still much lower than the asymptotic value it approaches as the pixel count exceeds about 1 million ($2.932381$).
 \par However, an opposite trend appears to be in play for the speedup of the non-maximum suppression and selective thresholding algorithm (Table \ref{non-maximum-suppression-selective-thresholding-speedup}), where we observe a high speedup for low input sizes and a lower speedup for high input sizes.
@@
 \section*{Source Code}
-The full source code for this project, including the CUDA and Python implementations of all of our algorithms, is available at \texttt{https://github.com/LINKIWI/elec-301}.
+The full source code for this project, including CUDA, Python, and MATLAB implementations of all of our algorithms, is open sourced and available at \texttt{https://github.com/LINKIWI/cuda-computer-vision}.
 
@@
 \begin{itemize}
-	\item \textbf{CJ Barberan}, Ph.D.
-student, Rice University Electrical and Computer Engineering---\textit{for mentoring the entire team in the conception, development, and extensive debugging of our project}.
+	\item \textbf{CJ Barberan}, ECE Ph.D. student, Rice University---\textit{for mentoring the entire team in the conception, development, and extensive debugging of our project}.
+	\item \textbf{Richard Baraniuk}, Victor E. Cameron Professor of ECE, Rice University---\textit{for his Fall 2015 instruction of the signal processing course that inspired this project}.
 \end{itemize}

### 2015-12-11 documents/paper/paper.tex
Edge-based motion detection block diagram

@@
 \subsection{GPU Computation and the NVIDIA CUDA Framework}
-Formally, NVIDIA CUDA is a parallel computing platform and programming model that exposes familiar C-based APIs for parallelized computations.
+Formally, CUDA (Compute Unified Device Architecture) is a parallel computing platform and programming model that exposes familiar C-based APIs for parallelized computations.
 CUDA is NVIDIA's platform for general-purpose computing on graphics processing units (commonly, GPGPU or GP$^2$U), or the use of a GPU for computations traditionally handled by the CPU.
@@
 Therefore, we view an opportunity use CUDA to explore the speedup possible by using a GPU to accelerate this common task in computer vision.
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{high_level_block_diagram.png}
+	\caption{High-level block diagram describing our implementation of a real-time edge-based motion detector}
+	\label{motion-detection-block-diagram}
+\end{figure*}
 

### 2015-12-11 documents/paper/paper.tex
added python subsection to motion detection section

--- 2015-12-11 documents/paper/paper.tex
@@
 Python was chosen because it is very easy for inexperienced programmers to use and OpenCV was chosen because it offers many common image processing algorithms in a variety of languages.
+Unfortunately using OpenCV limits us to the library's implementation, does not currently support the use of GPU parallelization in Python.
+To take advantage of the GPU and parallelization in Python we would have to use C wrappers, which defeats the purpose of our implementation in Python.
 
@@
 \subsection{High Level Python Implementation}
-Python is a widely used big-data programming language because of its extensive library support in such areas and its ease of use.
+The Python implementation of Algorithm 3 is very straightforward and looks almost exactly like the pseudo-code, making it easy to implement even for those with minimal programming experience.
+\begin{lstlisting}[
+	caption=Python implementation of the motion area estimation image,
+	label=python-motion-area-estimattion-implementation
+]
+def detect_motion(e1, e2):
+    height, width, depth = np.shape(e1)
+    d = difference(e1, e2)
+    d_prime = np.zeros((M, N), np.uint8)
+    for i in range(M):
+        for j in range(N):
+            x = 0
+            for i_prime in range(height/M):
+                for j_prime in range(width/N):
+                    if d[i*height/M+i_prime, j*width/N+j_prime] != 0:
+                        x += 1
+            if M*N*x/(height*width*1.0) > T:
+                d_prime[i, j] = 255
+    return d_prime
+\end{lstlisting}
 

### 2015-12-11 documents/paper/paper.tex

@@
 Therefore, we view an opportunity use CUDA to explore the speedup possible by using a GPU to accelerate this common task in computer vision.
+
+\begin{figure*}
+	\centering
+	\includegraphics[width=0.8\textwidth]{high_level_block_diagram.png}
+	\caption{High-level block diagram describing our implementation of a real-time edge-based motion detector}
+	\label{motion-detection-block-diagram}
+\end{figure*}
+
+\section{Facial Recognition Algorithm}
+Placeholder text
+
+\subsection{Placeholder}
+Placeholder text
+
+\subsection{Placeholder}
+Placeholder text
+
+\subsection{GPU Parallelization and CUDA Implementation}
+Placeholder text
 

### 2015-12-13 opencv-object-detect/CudaTestKevin.cu
code cleanup

@@ -45,7 +45,7 @@ void recog(string imageName, string mode)
 
                 cuda::GpuMat inputGPU(work_img);
                 cuda::GpuMat faces;
-                faces.create(10000000, 10000000, cv::DataType<cv::Rect>::type); ///
+                faces.create(1, 10000, cv::DataType<cv::Rect>::type); ///
                 cascade->setFindLargestObject(findLargestObject);
                 cascade->setScaleFactor(1.2);
                 cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
@@ -101,18 +101,6 @@ int main(int argc, char* argv[])
         images[6] = "Senthil5000.jpg";
         images[7] = "Senthil6000.jpg";
         images[8] = "Senthil7500.jpg";
-/*
-	
-	images[0] = "/media/kiwi/Windows 10/Users/Kevin Lin/Work Folders/OneDrive/Homework/Rice University/ELEC 301 - Signals and Systems/Project/main/opencv-object-detect/senthil100.jpg";
-	images[1] = "/media/kiwi/Windows 10/Users/Kevin Lin/Work Folders/OneDrive/Homework/Rice University/ELEC 301 - Signals and Systems/Project/main/opencv-object-detect/senthil500.jpg";
-	images[2] = "/media/kiwi/Windows 10/Users/Kevin Lin/Work Folders/OneDrive/Homework/Rice University/ELEC 301 - Signals and Systems/Project/main/opencv-object-detect/senthil1000.jpg";
-	images[3] = "/media/kiwi/Windows 10/Users/Kevin Lin/Work Folders/OneDrive/Homework/Rice University/ELEC 301 - Signals and Systems/Project/main/opencv-object-detect/Senthil2000.jpg";
-	images[4] = "/media/kiwi/Windows 10/Users/Kevin Lin/Work Folders/OneDrive/Homework/Rice University/ELEC 301 - Signals and Systems/Project/main/opencv-object-detect/Senthil3000.jpg";
-	images[5] = "/media/kiwi/Windows 10/Users/Kevin Lin/Work Folders/OneDrive/Homework/Rice University/ELEC 301 - Signals and Systems/Project/main/opencv-object-detect/Senthil4000.jpg";
[…]

### 2015-12-14 documents/paper/paper.tex
Added facial recognition section + some Python + future work in conclusion

--- 2015-12-11 documents/paper/paper.tex
@@
 
-\lstset{frame=tb,
-	language=C,
+\lstset{
+	frame=tb,
 	aboveskip=3mm,
@@
 
-\title{GPU Acceleration of Machine Learning-Aided Facial Recognition and Edge-Based Motion Detection with NVIDIA CUDA}
+\title{GPU Acceleration of Edge-Based Motion Detection and Machine Learning-Aided Facial Recognition with NVIDIA CUDA}
 
@@
 In this paper, we explore a GPU-based implementation of a simplified adaptation of existing edge detection algorithms fast enough to operate on frames of a continuous video stream in real-time.
-We also demonstrate practical applications of edge detection, including an edge-based method for motion detection estimation.
-Finally, we demonstrate the speedups of over 2x we achieve with GPU parallelism, as compared to a reference serial CPU-based implementation.
+We also demonstrate a practical application of edge detection--an edge-based method for motion detection estimation.
+Additionally, we explore the GPU-CPU speedup of existing OpenCV GPU computation libraries, namely, for facial recognition algorithms.
+Finally, we demonstrate the speedups as high as 10x we achieve with GPU parallelism, as compared to a reference serial CPU-based implementation.
 \end{abstract}
@@
 Further, SDKs like the NVIDIA CUDA framework provide developers easy APIs to take advantage of the parallel computing power of GPUs.
-We take full advantage of the computational benefits of GPUs by implementing edge detection and motion detection algorithms in CUDA C.
-\par In this paper, we first detail the theory for our edge detection and motion detection algorithms.
-Then, we describe the GPU code implementation of those algorithms with NVIDIA CUDA.
-Finally, we present speedup results achieved by our CUDA implementation with respect to a reference serial, CPU implmentation.
+We take full advantage of the computational benefits of GPUs by implementing edge detection and motion detection algorithms in CUDA C, and making use of existing CUDA libraries for our facial recognition algorithm.
+\par In this paper, we first detail the theory for our edge detection, motion detection, and facial recognition algorithms in Sections \ref{edge-detection-section}, \ref{motion-detection-section}, and \ref{facial-recognition-section}, respectively.
+At the end of Sections \ref{edge-detection-section} and \ref{motion-detection-section}, we describe our GPU code implementation of these algorithms with NVIDIA CUDA.
+At the end of Section \ref{facial-recognition-section}, we comment on the performance we achieve with a prebuilt, CUDA-based OpenCV GPU computation library, as opposed to that we achieve with a custom CUDA implementation as in Sections \ref{edge-detection-section} and \ref{motion-detection-section}.
+We present speedup results achieved with our CUDA implementation with respect to a reference serial, CPU implmentation in Section \ref{results}.
+Finally, we conclude in Section \ref{conclusion}.
 
@@
 For example, on modern hardware, an NVIDIA GeForce GTX 970 (1664 CUDA cores) exhibits peak single-precision floating point performance of nearly 3500 GFLOPS (floating-point operations per second), while an Intel Core i7 4790K (4C, 8T) achieves 100 GFLOPS.
-Our goal is to demonstrate the performance of GPU computing by solving a handful of existing problems in computer vision with CUDA: namely, edge detection and motion detection.
+Our goal is to demonstrate the performance of GPU computing by solving a handful of existing problems in computer vision with CUDA: namely, edge detection, motion detection, and facial recognition.
 
 \subsection{OpenCV and Python}
-While CUDA is an excelent language for maximizing performance, it could be infeasible due to inexperience with C, time constraints, or lack of an NVIDIA GPU.
-To address these constraints, we also implemented our algorithms in Python using OpenCV.
-Python was chosen because it is very easy for inexperienced programmers to use and OpenCV was chosen because it offers many common image processing algorithms in a variety of languages.
+While CUDA is an excellent language for maximizing performance, it could be infeasible due to inexperience with C, time constraints, or lack of an NVIDIA GPU.
+To address these constraints, we also implemented our algorithms in Python using the Python OpenCV library.
+We chose Python because it is very easy for inexperienced programmers to use, and we chose OpenCV because it offers implementations of many common image processing algorithms.
+Unfortunately, using OpenCV limits us to the library's implementation, which does not currently support the use of GPU parallelization in Python without using C wrappers.
+For each of the algorithms in this paper, we present code snippets that provide a sample CPU-based Python implementation alongside a GPU-based parallel CUDA implementation.
 
@@
 GPU parallelization is a powerful tool for accelerating image processing tasks.
-The motivation for our CUDA motion detection algorithm stems from the ease of creating a GPU implementation, due in large part to the simplicity of the CUDA APIs, and the diverse applications of motion detection.
+The motivation for our CUDA motion detection and facial recognition algorithms stems from their diverse applications and the ease of coding a GPU implementation, which is due in large part to the simplicity of CUDA APIs and OpenCV libraries.
 For example, motion detection is applied in real-time traffic analysis, vision-based security, and consumer technology (e.g.
@@
 Further, since our computer vision algorithm calculates the location of motion in a frame based on detected edges, our approach is both more accurate and supplemented with more specific data.
-Therefore, we view an opportunity use CUDA to explore the speedup possible by using a GPU to accelerate this common task in computer vision.
+Facial recognition is applied in real-time security, smart photographic analysis, and military purposes, including target-based missile guiding systems.
+Therefore, we view an opportunity use CUDA to explore the speedup possible by using a GPU to accelerate these common tasks in computer vision.
 
@@
 
-\section{Facial Recognition Algorithm}
-Placeholder text
-
-\subsection{Placeholder}
-Placeholder text
-
-\subsection{Placeholder}
-Placeholder text
-
-\subsection{GPU Parallelization and CUDA Implementation}
-Placeholder text
-
 \section{Edge Detection Algorithm}
+\label{edge-detection-section}
 Edge detection is the process of determining the locations of boundaries that separate objects, or sections of objects, in an image.
@@
 	caption=Grid configuration for the separable convolution device kernel,
-	label=grid-configuration
+	label=grid-configuration,
+	language=C
 ]
@@
 	caption=Implementation of horizontal and vertical separated convolution,
-	label=separable-convolution-implementation
+	label=separable-convolution-implementation,
+	language=C
 ]
@@
 	caption=GPU computation of gradient magnitude and angle,
-	label=gradient-magnitude-angle-implementation
+	label=gradient-magnitude-angle-implementation,
+	language=C
 ]
@@
 	caption=GPU non-maximum suppression and selective thresholding,
-	label=non-maximum-suppression-selective-thresholding-implementation
+	label=non-maximum-suppression-selective-thresholding-implementation,
+	language=C
 ]
@@
 
+\subsection{High-Level Python Implementation}
+The Python OpenCV library exposes an API for the entire Canny edge detection algorithm.
+The Canny algorithm is similar to, but more nuanced than, our algorithm, substituting our selective thresholding procedure for hysteresis.
+However, the end results are very similar.
+As is evident from Listing \ref{python-edge-detection-implementation}, the procedure is very straightforward.
+\begin{lstlisting}[
+	caption=Python implementation of Canny edge detection with OpenCV,
+	label=python-edge-detection-implementation,
+	language=Python
+]
+LOW_THRESHOLD = 100
+HIGH_THRESHOLD = 200
+edges = cv2.Canny(img, LOW_THRESHOLD, HIGH_THRESHOLD)
+\end{lstlisting}
+
 \section{Motion Detection Algorithm}
+\label{motion-detection-section}
 Our motion detection algorithm compares the edges from two frames temporally separated by a small interval of time and attempts to estimate regions of the frame that exhibit movement or motion.
@@
 	caption=GPU computation of thresholded difference matrix,
-	label=difference-filter-implementation
+	label=difference-filter-implementation,
+	language=C
 ]
@@
 	caption=GPU computation of the difference density matrix,
-	label=difference-density-implementation
+	label=difference-density-implementation,
+	language=C
 ]
@@
 	caption=GPU generation of the motion area estimation image,
-	label=motion-area-estimation-implementation
+	label=motion-area-estimation-implementation,
+	language=C
 ]
@@
 
-\subsection{High Level Python Implementation}
-Python is a widely used big-data programming language because of its extensive library support in such areas and its ease of use.
+\subsection{High-Level Python Implementation}
+The Python implementation of Algorithm \ref{difference-density-matrix-algorithm} is very straightforward and looks almost exactly like the pseudo-code, making it easy to implement even for those with minimal programming experience.
+\begin{lstlisting}[
+	caption=Python implementation of the motion area estimation image,
+	label=python-motion-area-estimattion-implementation,
+	language=Python
+]
+import numpy as np
+
+def detect_motion(e1, e2):
+    height, width, depth = np.shape(e1)
+    d = difference(e1, e2)
+    d_prime = np.zeros((M, N), np.uint8)
+    for i in range(M):
+        for j in range(N):
+            x = 0
+            for i_prime in range(height/M):
+                for j_prime in range(width/N):
+                    if d[i*height/M+i_prime, j*width/N+j_prime] != 0:
+                        x += 1
+            if M*N*x/(height*width*1.0) > T:
+                d_prime[i, j] = 255
+    return d_prime
+\end{lstlisting}
+
+\section{Facial Recognition Algorithm}
+\label{facial-recognition-section}
+Our facial recognition implementation is based off of the HAAR features-based cascade classifiers method proposed by Viola-Jones in 2001.
+It's split into two phases: training and detection.
+The algorithm is as follows:
+\begin{enumerate}
+	\item Extract HAAR features of a face from a set of positive and negative training images. 
+	\item Use adaptive boosting machine learning to train a multi-stage, or cascade, classifier against positive and negative HAAR features. 
+	\item Apply a final cascade classifier to a loaded image or real time video frame. 
+\end{enumerate}
+Due to the resource capacity necessary to train a full facial recognition classifier, we used a pre-trained frontal face cascade classifier and used OpenCV's object detection library, which encapsulates the algorithm above, and whose theory detailed in the following sections.
+
+\subsection{HAAR Feature Extraction}
+In mathematics analysis, Haar wavelets are a sequence of rescaled, square-like functions.
+In image processing, Haar features are so named for their resemblance to Haar wavelets.
+Haar features are regularities identified in all human faces, by comparing light and dark gradients across the face.
+These are split into edge (two-rectangle) features, line (three-rectangle) features, and diagonal (four-rectangle) features.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{haar_features.png}
+	\caption{Example application of HAAR features on a face}
+    \label{haar-features-face-example}
+\end{figure}
+\par Each feature is obtained by comparing the pixel intensity in the white boxes against the pixel intensity in the dark boxes for a wide sample of sub-windows across the entire image.
+A 24x24 pixel image window can produce over 160,000 features that need to be classified, in some sense, “oversampling” the face.
+\par The rectangle features can be computed by splitting the image into sub-windows called integral images.
+An integral image at a point $i'(x, y)$ is the sum of all pixels to the top and to the left of that point.
+\[
+i'(x, y) = \sum_{m \leq x, n \leq y}{i(m, n)}
+\]
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{integral_image.png}
+	\caption{Integral image on a face}
+    \label{integral-image-face}
+\end{figure}
+Since an integral image includes all previously-computed integral images contained in its window, in the image of Figure \ref{integral-image-face}, the region A can be computed as
+\[
+A = i'(4) + i'(1) - i'(2) - i'(3)
+\]
+Any such rectangular sum can be computed using four array references.
+
+\subsection{Cascade Classifier Training}
+It's far too resource intensive to evaluate over 160,000 images for each 24x24 pixel window.
+However, we can take advantage of the fact that only a very small fraction of extracted sub-windows in an uncropped image will correspond to accurate facial features.
+To do so, we use a cascaded variation of the Adaptive Boosting (or AdaBoost) machine learning algorithm.
+AdaBoost aims to determine the optimal binary threshold of HAAR features which best separate negative and positive samples.
+\begin{algorithm}[]
+	\smaller
+	\label{adaboost-classifier-algorithm}
+	\caption{AdaBoost classifier training}
+
+	\KwIn{Set of training images $I$, where each image $i \in I$ is a two-dimensional matrix representing the image, and is labeled as either a positive or negative sample image; $m$, the number of positive samples; $l$, the number of negative samples; $F$, the set of features extracted from images in $I$; $\theta_f, \forall f \in F$, denoting a threshold for weakly classifying a feature match}
+	\KwOut{Strong classifier function $h(x)$ whose value is $1$ if the image $x$ contains a face, and $0$ otherwise}
+
+	\nl Initialize $w$ to a mapping from images to image weights\;
+	\nl Initialize $t$ to a mapping from images to the constant $1$ or $0$\;
+	\nl \ForEach {$i \in I$} {
+		\nl \If {$i$ \textnormal{is a positive training sample}} {
+			\nl $t_i \gets 1$\;
+			\nl $w_i \gets \frac{1}{2m}$\;
+		}
+		\nl \Else {
+			\nl $t_i \gets 0$\;
+			\nl $w_i \gets \frac{1}{2l}$\;
+		}
+	}
+	\nl \ForEach {$f \in F$} {
+	\nl $g_f \gets \textnormal{ trained weak classifier  } \begin{cases}1 & \textnormal{value of } f < \theta_f \\ 0 & \textnormal{else}\end{cases}$\;
+		\nl \ForEach {$i \in I$} {
+			\nl $w' \gets w \textnormal{ normalized such that } \sum_{k \in I} w_k = 1$\;
+			\nl $\varepsilon_f \gets \textnormal{the minimum value of the error } \sum_{k \in I} w'_k |g_f(i) - t_i|$\;
+			\nl $\beta_f \gets \textnormal{weight inversely proportional to } \varepsilon_f$\;
+			\nl $w_i \gets \beta_f w'_i$\;
+		}
+	}
+	\nl $h(x) \gets \begin{cases}1 & \sum_{f \in F} g_f(x) \log{\frac{1}{\beta_f}} \geq \frac{1}{2} \sum_{f \in F} \log{\frac{1}{\beta_f}} \\ 0 & \textnormal{else}\end{cases}$\;
+	\nl \Return $h(x)$\;
+	\
+\end{algorithm}
+\par From one pass at the AdaBoost algorithm, it is clear that running hundreds of thousands of features on millions of samples will be too computationally intensive.
+However, most features in a general, uncropped image do not correspond to a face.
+This allows us to set up a cascaded variant of the AdaBoost algorithm.
+\par In cascading, all the features are grouped into several stages, where each stage checks for a certain amount of features.
+It's set up such that the most common facial features are grouped into the earliest cascade stages.
+This allows us to discard most non-facial regions early on and only spend successive computations on potentially positive regions.
+If a region fails during a stage, it is discarded, the cascade training stops, and then resets at the next potentially positive region.
+This vastly improves resource and time efficiency of classifier training.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{cascade_framework.png}
+	\caption{Cascade training framework}
+    \label{cascade-framework}
+\end{figure}
+
+\subsection{Application of Classsifier to Facial Recognition}
+The final output of the cascade classifier training is a classifier XML file that contains the results of the training.
+OpenCV then provides an easy module to apply that XML file to real time face detection.
+We just need to import the file, pre-process our images (or frames from real-time video stream), and then run the OpenCV object detection functions.
+\par We load the trained classifier, load our image, and then the object detect functions returns the position of any found faces as a Rectangle with identifiers (x, y, width, height).
+With the returned positions, we can draw a basic rectangle of the detected facial regions or even perhaps take it a step further and apply Daft Punk masks to any detected faces.
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{richb.jpg}
+	\caption{Sample input image to the classification}
+    \label{richb}
+\end{figure}
+\begin{figure}[H]
+	\centering
+	\includegraphics[width=0.45\textwidth]{richb_face_detected.png}
+	\caption{Detected faces from the image of Figure \ref{richb} using the trained Frontal Face HAAR classifier}
+    \label{richb-face-detected}
+\end{figure}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{daft_punk_masks.png}
+	\caption{Sample application of facial recognition: scale-invariant Daft Punk masks applied to detected facial regions in real-time}
+    \label{daft-punk-masks}
+\end{figure}
+
+\subsection{GPU Parallelization and CUDA Implementation}
+The CUDA implementation in OpenCV's object detection framework takes advantage of the inherent parallelization available in pre-processing the images and traversing all of the image sub-windows.
+\par Unlike previously with the edge detection and motion detection, where there is no singularly accepted CUDA implementation of the aforementioned algorithms, with facial recognition, OpenCV's library has been heavily optimized over the years to provide that functionality.
+\par It would an interesting case study then to abstract the functionalities of their object detection libraries, and compare the speedups attained from using a heavily optimized 3rd party library against the speedups which we obtained from programs written by us from scratch.
+\par OpenCV's CUDA implementation is similar to its serial CPU implementation, except that the cascade classifier is loaded in a CUDA optimized format, and the images are loaded and processed in CUDA-optimized matrices formats created by OpenCV.
 
 \section{GPU Speedup Results}
+\label{results}
 Our code for separable convolution, computation of the image's gradient magnitude and angle, non-maximum suppression and selective thresholding, and computation of the thresholded difference and spatial difference density matrices has both serial CPU and parallel CUDA implementations.
+Both the CPU and GPU implementations of the entire facial recognition algorithm are encapsulated in an OpenCV library internally implemented with CUDA.
 Here, we present speedup results for each of these procedures on various image sizes obtained on a system with the following hardware:
@@
     \label{motion-area-estimation-speedup-graph}
+\end{figure}
+
+\subsection{Facial Recognition}
+The below results reflect the speedup achieved using OpenCV's object detection framework.
+\begin{table}[H]
+	\small
+	\centering
+	\caption{Speedup of Facial Recongition}
+	\label{facial-recognition-speedup}
+	\begin{tabular}{llllll}
+	 Image dimensions & CPU time (s) & GPU time (s) & Speedup \\
+	 $100 \times 100$ & $0.067883$ & $0.061132$ & $1.110433$ \\
+	 $500 \times 500$ & $0.226192$ & $0.094812$ & $2.385690$ \\
+	 $1000 \times 1000$ & $0.811302$ & $0.199005$ & $4.076792$ \\
+	 $2000 \times 2000$ & $2.849201$ & $0.562820$ & $5.062366$ \\
+	 $3000 \times 3000$ & $6.236422$ & $1.218879$ & $5.116523$ \\
+	 $4000 \times 4000$ & $10.794892$ & $2.142449$ & $5.038576$ \\
+	\end{tabular}
+\end{table}
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{facial_recognition_speedup_graph.jpg}
+	\caption{Visualization of facial recognition speedup versus pixel count}
+    \label{facial-recognition-speedup-graph}
 \end{figure}
@@
 \section{Conclusion}
+\label{conclusion}
 Our CUDA implementation of both the edge detection and motion detection algorithms demonstrate that parallelized, GPU computation results in significant speedups compared to a serial, CPU implementation.
 In all benchmarked cases (separable convolution with a Gaussian filter, edge detection via non-maximum suppression and selective thresholding, and motion area estimation from a difference density map), we find that the GPU implementation, running on a mid-range graphics card, demonstrated anywhere between a 1.5x to 4.6x speedup over a relatively high-performance, overclocked CPU.
-Further, for each algorithm, we find that the speedup asymptotically approaches a general range as the input size increases.
+Furthermore, we observe a similar speedup trend when comparing existing OpenCV CPU and GPU CUDA implementations of HAAR-based facial recognition.
+In all cases, we find that the speedup asymptotically approaches a general range as the input size increases.
 \par It is important to note the edge case for the separable convolution algorithm on images of small input sizes (namely, near the dimensions $100 \times 100$, or $10000$ total pixels).
@@
 Despite this trend, the speedup is still greater than 1 for large input sizes, suggesting that a GPU implementation is still consistently faster than its CPU counterpart.
+\par While we were successful in exploiting the parallel architecture of GPUs to achieve significant speedups, there are still many areas in GPU computation we have yet to explore.
+Our CUDA code for the edge detection and motion detection are direct implementations of the algorithms (albeit parallelized on each pixel in the image).
+We did not attempt any degree of memory optimization beyond ensuring that no memory leaks occur on either the host or device when running the real-time program.
+We also did not attempt to minimize the amount of time any given GPU thread is idle; this delay (after a particular thread has finished its computation task but has not yet been assigned another computation task) is most likely the largest bottleneck in our parallel performance.
+Future improvements to our existing work would lie primarily in optimizing these parameters, thereby increasing performance.
+\par A future goal is exploring multi-GPU computation.
+Our implementation, through parallelized, is optimized to run only on a single GPU.
+Multi-GPU computation would add another layer of parallelism that could theoretically increase the throughput of our computation by the number of GPUs across which the work is divided.
+NVIDIA's existing multi-GPU computational platform is marketed as Scalable Link Interface (SLI), allowing for the simultaneous use of 2 to 4 GPUs to render real-time graphics.
+In context of the work we present here, a possible future goal is to explore the performance and memory efficiency gain from Split Frame Rendering (SFR), by allowing each of our parallel algorithms to operate simultaneously on a division of the input image; or Alternate Frame Rendering (AFR), to allow for the simultaneous computation on multiple sets of 2 frames.
 
 \section*{Source Code}
-The full source code for this project, including CUDA, Python, and MATLAB implementations of all of our algorithms, is open sourced and available at \texttt{https://github.com/LINKIWI/cuda-computer-vision}.
-
-\section*{Team}
+The full source code for this project, including CUDA, Python, and MATLAB implementations of all of our algorithms, is open sourced and available at \texttt{github.com/LINKIWI/cuda-computer-vision}.
+
+\section*{Authors}
 \begin{itemize}

### 2015-12-16 opencv-object-detect/CudaTest2.cu
cleanup

@@ -1,309 +1,0 @@
-#include <opencv2/opencv.hpp>
-#include <opencv2/cudaobjdetect.hpp>
-#include <iostream>
-#include <sys/time.h>
-#include "opencv2/objdetect/objdetect.hpp"
-#include "opencv2/highgui/highgui.hpp"
-#include "opencv2/imgproc/imgproc.hpp"
-#include "opencv2/cudaimgproc.hpp"
-#include "opencv2/cudawarping.hpp"
-#include <iomanip>
-#include <cuda_runtime.h>
-
-using namespace std;
-using namespace cv;
-
-//global call of cascade
-
-void recog(string imageName, string mode)
-{
-
-	string cascadeName = "/home/ubuntu/Downloads/opencv/data/haarcascades_cuda/haarcascade_frontalface_alt2.xml";
-        Ptr<cuda::CascadeClassifier> cascade = cuda::CascadeClassifier::create(cascadeName);
-        cv::CascadeClassifier cascadeCPU;
-        cascadeCPU.load(cascadeName);
-	
-	if (cascade.empty() || cascadeCPU.empty()){
-                cerr << "Could not load model!" << endl;
-               // return -1;
-        }
-	
-	Mat original = imread(imageName);
-    	Mat gray = imread(imageName, CV_LOAD_IMAGE_GRAYSCALE);
-    	Mat work_img = original.clone();
-    	Mat copy = original.clone();
-
-	//pre processing
-	cvtColor(original, work_img, COLOR_RGB2GRAY);
-    	equalizeHist(work_img, work_img);
-
-    	bool findLargestObject = false;
-    	bool filterRects = true;
-
-	if (mode == "--GPU")
-        {
-
-                cuda::GpuMat inputGPU(work_img);
-                cuda::GpuMat faces;
-		faces.create(1, 100000, cv::DataType<cv::Rect>::type); ///
-                cascade->setFindLargestObject(findLargestObject);
-                cascade->setScaleFactor(1.2);
-                cascade->setMinNeighbors((filterRects || findLargestObject) ? 4 : 0);
-                cascade->detectMultiScale(inputGPU, faces);
-
-                vector<Rect> objects;
-                cascade->convert(faces, objects);
-
-                for(int i=0;i<(int)objects.size();++i)
-                {
-                     rectangle(original, objects[i], Scalar(0, 255, 0), 3);
-                }
-
[…]

### 2015-12-16 opencv-object-detect/cuda-facedetect.cu

@@ -1,107 +1,0 @@
-#include <stdio.h>
-#include <iostream>
-#include <opencv2/core/core.hpp>
-#include <opencv2/imgproc/imgproc.hpp>
-#include <opencv2/highgui/highgui.hpp>
-#include <opencv2/objdetect/objdetect.hpp>
-
-using namespace cv;
-using namespace std;
-using namespace cv::cuda;
-
-#include <opencv2/cudaobjdetect.hpp>
-#include <opencv2/gpu/gpu.hpp>
-#include <vector>
-#include "fstream"
-#include <opencv2/ml/ml.hpp>
-#include <opencv2/core/gpumat.hpp>
-#include <opencv2/video/tracking.hpp>
-#include <opencv2/core/cuda.hpp>
-
-//cv::cuda::CascadeClassifier_CUDA cascade_gpu;
-
-Ptr<cuda::CascadeClassifier> cascade_gpu = cuda::CascadeClassifier::create("haarcascade_frontalface_default.xml");
-
-//vector<Rect> detect_faces(Mat& image)
-void detect_faces(Mat& image)
-{
-	vector<Rect> res;
-	bool findLargestObject = true;
-	bool filterRects = true;
-	int detections_num;
-	Mat faces_downloaded;
-	Mat im(image.size(), CV_8UC1);
-	GpuMat facesBuf_gpu;
-	if(image.channels()==3)
-	{
-		cvtColor(image, im, CV_BGR2GRAY);
-	}
-	else
-	{
-		image.copyTo(im);
-	}
-	GpuMat gray_gpu(im);
-	
-//	cascade_gpu.visualizeInPlace = false;
-//	findLargestObject=cascade_gpu->FindLargestObject;
-//	detections_num = cascade_gpu->detectMultiScale(gray_gpu, facesBuf_gpu, 1.2, (filterRects || findLargestObject) ? 4 : 0,Size(image.cols/4,image.rows/4));
-//	detections_num = cascade_gpu->detectMultiScale(gray_gpu, facesBuf_gpu);
-//	if(detections_num==0){return res;}
-	
-//	facesBuf_gpu.colRange(0, detections_num).download(faces_downloaded);
-//	Rect *faceRects = faces_downloaded.ptr<Rect>();
-
-//	for(int i = 0; i<detections_num; i++)
-//	{
-//		res.push_back(faceRects[i]);
-//	}
-	
[…]

### 2015-12-16 documents/paper/paper.tex
Corrected typos in paper

@@
 At the end of Sections \ref{edge-detection-section} and \ref{motion-detection-section}, we describe our GPU code implementation of these algorithms with NVIDIA CUDA.
-At the end of Section \ref{facial-recognition-section}, we comment on the performance we achieve with a prebuilt, CUDA-based OpenCV GPU computation library, as opposed to that we achieve with a custom CUDA implementation as in Sections \ref{edge-detection-section} and \ref{motion-detection-section}.
-We present speedup results achieved with our CUDA implementation with respect to a reference serial, CPU implmentation in Section \ref{results}.
+At the end of Section \ref{facial-recognition-section}, we comment on the performance we achieve with a pre-built, CUDA-based OpenCV GPU computation library, as opposed to that we achieve with a custom CUDA implementation as in Sections \ref{edge-detection-section} and \ref{motion-detection-section}.
+We present speedup results achieved with our CUDA implementation with respect to a reference serial, CPU implementation in Section \ref{results}.
 Finally, we conclude in Section \ref{conclusion}.
@@
 The selective thresholding technique extends on the procedure of non-maximum suppression.
-For any pixel that passes crtiteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel.
+For any pixel that passes criteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel.
 If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted.
-This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satify the criterion themselves, in order to create more continuous edge lines.
+This technique attempts to catch pixels that might ``connect" ridge pixels identified by criteria (2) but fail to satisfy the criterion themselves, in order to create more continuous edge lines.
 \begin{figure}[]
@@
 	\centering
-	\caption{Speedup of Facial Recongition}
+	\caption{Speedup of Facial Recognition}
 	\label{facial-recognition-speedup}

### 2015-12-16 opencv-object-detect/facedetect.py
cleanup

@@ -7,10 +7,12 @@ import numpy
 cascade = 'haarcascade_frontalface_default.xml'
 faceCascade = cv2.CascadeClassifier(cascade)
 
-image1 = os.path.join(os.path.dirname(__file__), 'birthday.jpg')
+image1 = 'RichB.jpg'
 
 daftPunk = cv2.imread('DaftPunk.png', cv2.IMREAD_UNCHANGED)
 
+# train classifiers
+# apply a daft punk mask
 
 def faceDetectWebcam():
 
@@ -46,9 +48,6 @@ def faceDetectWebcam():
     cv2.destroyAllWindows()
 
 
-# train classifiers
-# train to recognize rich b's face, and apply a daft punk mask
-
 def faceDetectTiming(images):
     image = cv2.imread(images, cv2.IMREAD_UNCHANGED)
     start_time = dt.datetime.now()
@@ -68,11 +67,7 @@ def faceDetectTiming(images):
     print("detected in:" + " " + str(time) + "ms")
 
     for (x, y, w, h) in face:
-        # recog = cv2.resize(daftPunk, (w, h), interpolation=cv2.INTER_AREA)
-        # print(recog.shape, recog.shape[0], recog.shape[1], w, h)
-        # for c in range(0,3):
-        #   image[y:y+h, x:x+w, c] = recog[:,:,c] * (recog[:,:,3]/255.0) + image[y:y+h, x:x+w, c] * (1.0 - recog[:,:,3]/255.0)
-        # image[y:y+h, x:x+w] = recog
+ 
         cv2.rectangle(image, (x, y), (x+w, y+h), (0, 255, 0), 2)
 
     cv2.imshow("Faces found", image)
@@ -83,5 +78,5 @@ def faceDetectTiming(images):
         cv2.destroyAllWindows()
 
 
-faceDetectTiming(image1)
-#faceDetectWebcam()
+#faceDetectTiming(image1)
+faceDetectWebcam()

### 2015-12-16 documents/paper/paper.tex
Add Python listing for facial recognition

@@
 ]
+import cv2
+
 LOW_THRESHOLD = 100
 HIGH_THRESHOLD = 200
+
 edges = cv2.Canny(img, LOW_THRESHOLD, HIGH_THRESHOLD)
@@
 \end{figure}
+
+\subsection{High-Level Python Implementation}
+We again provide sample Python code to demonstrate how the facial recognition algorithm would be implemented on a high level, with the details of implementation abstracted out.
+The Python OpenCV library exposes an API for facial recognition via a cascade classifier, as shown in Listing \ref{python-facial-recognition-implementation}.
+\begin{lstlisting}[
+	caption=Python implementation of facial recognition with OpenCV,
+	label=python-facial-recognition-implementation,
+	language=Python
+]
+import cv2
+
+IMAGE = 'richb.jpg'
+CASCADE_DATA = 'haarcascade_frontalface_default.xml'
+
+face_cascade = cv2.CascadeClassifier(CASCADE_DATA)
+faces = face_cascade.detectMultiScale(
+	cv2.imread(IMAGE, cv2.IMREAD_GRAYSCALE),
+	scaleFactor=1.1,
+	minNeighbors=5,
+	minSize=(30, 30),
+	flags=cv2.CASCADE_SCALE_IMAGE,
+)
+# faces now contains tuples indicating the location and size of each detected face in the image
+\end{lstlisting}
 

### 2015-12-16 documents/paper/paper.tex
Corrected most grammatical errors in paper

@@
 Additionally, we explore the GPU-CPU speedup of existing OpenCV GPU computation libraries, namely, for facial recognition algorithms.
-Finally, we demonstrate the speedups as high as 10x we achieve with GPU parallelism, as compared to a reference serial CPU-based implementation.
+Finally, we demonstrate speedups as high as 10x we achieve with GPU parallelism, as compared to a reference serial CPU-based implementation.
 \end{abstract}
@@
 	\item Determine the two-dimensional gradient of the filtered image by applying partial derivatives in both the horizontal and vertical directions.
-	\item Classify edges by suppressing surrounding pixels in the gradient direction whose magnitude is lesser, and applying selective thresholding to a limited apron around each selected pixel.
+	\item Classify edges by accepting pixels of a minimum threshold gradient magnitude and suppressing its neighboring pixels in the gradient direction with a lesser gradient magnitude, and applying selective thresholding to a limited apron around each selected pixel.
 \end{enumerate}
@@
 The filtering necessary in steps (1) and (2) from our algorithm above involve the convolution of a two-dimensional image and a two-dimensional kernel filter of constant width and height.
-A na\"{\i}ve implementation of two-dimensional convolution would require a double sum across the kernel and image for each individual pixel:
+A na\"{\i}ve implementation of two-dimensional convolution would require a double sum across the convolution kernel and input image for each individual pixel in the output image:
 \begin{equation}
@@
 For example, consider the $5312 \times 2988$ pixel image (taken at a high ISO to deliberately introduce noise) of Figure \ref{noisy-image} and a plot of its grayscale intensity versus the image's spatial dimensions in Figure \ref{noisy-image-mesh}.
-It is clear from the mesh that there is much high-frequency noise, which can be removed with a low-pass filter.
+It is clear from the mesh that there exists much high-frequency noise, which can be removed with a low-pass filter.
 \begin{figure}[H]
@@
 	\includegraphics[width=0.45\textwidth]{noisy_image_mesh.jpg}
-	\caption{Grayscale intensity of unfiltered noisy image at each pixel}
+	\caption{Grayscale intensity of the unfiltered noisy image of Figure \ref{noisy-image} at each pixel, visualized as a three-dimensional mesh}
     \label{noisy-image-mesh}
@@
 	\includegraphics[width=0.45\textwidth]{fft_noisy_image.jpg}
-	\caption{FFT magnitude of unfiltered noisy image at each pixel (log scale)}
+	\caption{FFT magnitude of the unfiltered noisy image of Figure \ref{noisy-image} at each pixel (log scale)}
     \label{noisy-image-fft}
@@
 \end{equation}
-\par Applying a Gaussian filter with parameters $k = 5$ and $\sigma = 5$ to the above image significantly denoises the image without sacrificing edge precision, which can be seen from the spatial intensity plot in Figure \ref{filtered-noisy-image-mesh}.
+\par Applying a Gaussian filter with parameters $k = 5$ and $\sigma = 5$ to the noisy image of Figure \ref{noisy-image} significantly denoises the image without sacrificing edge precision, which can be seen from the spatial intensity plot in Figure \ref{filtered-noisy-image-mesh}.
 \begin{figure}[H]
@@
 	\includegraphics[width=0.45\textwidth]{filtered_noisy_image_mesh.jpg}
-	\caption{Grayscale intensity of Gaussian-filtered noisy image at each pixel}
+	\caption{Grayscale intensity of the Gaussian-filtered noisy image of Figure \ref{filtered-noisy-image} at each pixel, visualized as a three-dimensional mesh}
     \label{filtered-noisy-image-mesh}
@@
 	\includegraphics[width=0.45\textwidth]{fft_filtered_noisy_image.jpg}
-	\caption{FFT magnitude of Gaussian-filtered noisy image at each pixel (log scale)}
+	\caption{FFT magnitude of the Gaussian-filtered noisy image of Figure \ref{filtered-noisy-image} at each pixel (log scale)}
     \label{filtered-noisy-image-fft}
@@
 \end{equation}
-where $G_x$ and $G_y$ are the values of the partial derivatives along the $x$ and $y$ directions, respectively, at the pixel located at $(i, j)$ of the image $A$.
+where $G_x$ and $G_y$ are the values of the partial derivatives along the $x$ and $y$ directions, respectively, at the pixel located at $(i, j)$ of the image $\boldsymbol{A}$.
 $$G_x = \frac{\partial \boldsymbol{A}}{\partial x}\Bigr|_{\substack{(x, y) = (i, j)}}$$
@@
 	\includegraphics[width=0.45\textwidth]{sobel_pcb.jpg}
-	\caption{Magnitude of the gradient $||\boldsymbol{G}||$, as determined by applying a Sobel filter in both the $x$ and $y$ directions, at each pixel of the Gaussian-filtered original image with parameters $k = 5$ and $\sigma = 5$}
+	\caption{Magnitude of the gradient $||\boldsymbol{G}||$ as determined by applying a Sobel filter in both the $x$ and $y$ directions to the Gaussian-filtered original image of Figure \ref{pcb-original} with parameters $k = 5$ and $\sigma = 5$}
     \label{sobel-pcb}
@@
 	\includegraphics[width=0.45\textwidth]{sobel_pcb_mesh.jpg}
-	\caption{Gradient of Figure \ref{sobel-pcb} visualized as a three-dimensional mesh}
+	\caption{Gradient magnitude of Figure \ref{sobel-pcb} visualized as a three-dimensional mesh}
     \label{sobel-pcb-mesh}
@@
 \end{itemize}
-The selective thresholding technique extends on the procedure of non-maximum suppression.
-For any pixel that passes criteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha$ x $\alpha$ box around the pixel.
+\par The selective thresholding technique extends on the procedure of non-maximum suppression.
+For any pixel that passes criteria (2) (maximized, relative to the neighboring pixels, along the gradient direction), consider an $\alpha \times \alpha$ box around the pixel.
 If any pixel within the box exceeds the low threshold $t_l$, then the pixel is accepted.
@@
 In computing the output $y[m, n]$ at every index $(i, j)$, we launch a kernel with 16 threads per block, where the total number of blocks along one dimension is equal to the size of that dimension divided by 16.
-For our 5312 x 2988 test images, this corresponds to 332 blocks in the horizontal direction and 186 blocks in the vertical direction, each with 16 computational threads.
+For our 5312 $\times$ 2988 test images, this corresponds to 332 blocks in the horizontal direction and 186 blocks in the vertical direction, each with 16 computational threads.
 \begin{lstlisting}[
@@
 \end{lstlisting}
-\par The first stage of our edge detection algorithm uses the separable convolution GPU implementation above to apply a Gaussian low-pass filter in Equation \ref{gaussian-filter} to remove high-frequency noise from the image that might be falsely labeled as an edge.
+\par The first stage of our edge detection algorithm uses the separable convolution GPU kernel above to apply the Gaussian low-pass filter of Equation \ref{gaussian-filter} to remove high-frequency noise from the image that might be falsely labeled as an edge.
 Then, the same separable convolution kernel is used to apply the Sobel edge filter of Equations \ref{horizontal-sobel-filter} and \ref{vertical-sobel-filter} to the Gaussian low pass-filtered image in both the horizontal and vertical directions to approximate the horizontal and vertical partial derivatives, respectively.
@@
 	) {
-	const int r = blockIdx.y * blockDim.y + threadIdx.y;
-	const int c = blockIdx.x * blockDim.x + threadIdx.x;
-	const int i = r * (input_width + 2) + c;
+	int r = blockIdx.y * blockDim.y + threadIdx.y;
+	int c = blockIdx.x * blockDim.x + threadIdx.x;
+	int i = r * (input_width + 2) + c;
 
@@
 However, in practice, it is more useful to consider general regions of the image around which motion exists.
-For this reason, we build a density map of the matrix $\boldsymbol{D}$ using equal-sized rectangular sections of the input image that approximates the density of the difference per unit area (units pixels$^2$).
+For this reason, we build a density map of the matrix $\boldsymbol{D}$ using equal-sized rectangular sections of the input image that approximates the density of the difference per unit area.
 Thus, the original $M \times N$ difference matrix $\boldsymbol{D}$ is reduced to a matrix $\boldsymbol{D}'$ of size $m \times n$ such that
@@
 $$N = k_2 n$$
-where $k_1, k_2 \in \mathbb{Z}$.
+where $k_1, k_2 \in \mathbb{Z}^+$.
 Intuitively, choosing a higher $k_1$ and $k_2$ will result in a higher resolution over which the density is evaluated.
-Selection of $k_1$ and $k_2$ should depend on how large of a region over which motion should be estimated.
+Selection of $k_1$ and $k_2$ should depend on the desired size of the region over which motion should be estimated.
 The upper limits of $k_1$ and $k_2$ are $M$ and $N$, respectively, at which point $\boldsymbol{D}'$ = $\boldsymbol{D}$.
@@
 \par Regions of motion are the rectangular sections whose density exceeds a user-defined threshold $\Gamma$, denoting the number of difference pixels over the total number of pixels considered in the rectangular section.
-For each such region $\boldsymbol{R}$ in $\boldsymbol{A}$ of size $\frac{M}{m}$ x $\frac{N}{n}$,
+For each such region $\boldsymbol{R}$ in $\boldsymbol{A}$ of size $\frac{M}{m} \times \frac{N}{n}$,
 $$\textnormal{DifferenceDensity}(\boldsymbol{R}) = \frac{\textnormal{number of pixel differences in } \boldsymbol{R}}{\textnormal{number of pixels in } \boldsymbol{R}}$$
@@
 \end{lstlisting}
-The construction of the difference density matrix as in Algorithm \ref{difference-density-matrix-algorithm} is parallelized by considering the value of the thresholded difference matrix $\boldsymbol{D}'$ at each index $(r, c)$, mapping that index to the appropriate index in the difference density matrix $(i, j)$ where $i < m \wedge j < n$, then adding to the value of the array at $(i, j)$ at each encounter of a difference pixel in $\boldsymbol{D}'$.
+The construction of the difference density matrix as in Algorithm \ref{difference-density-matrix-algorithm} is parallelized by considering the value of the thresholded difference matrix $\boldsymbol{D}$ at each index $(r, c)$, mapping that index to the appropriate index in the difference density matrix $(i, j)$ where $i < m \wedge j < n$, then adding to the value of the array at $(i, j)$ at each encounter of a difference pixel in $\boldsymbol{D}$.
 Our CUDA implementation therefore relies on the fact that the density array is initialized to 0 before the kernel is launched; this is accomplished in practice with a \texttt{calloc} operation on the host, followed by copying that array to GPU memory.
@@
 \label{facial-recognition-section}
-Our facial recognition implementation is based off of the HAAR features-based cascade classifiers method proposed by Viola-Jones in 2001.
+Our facial recognition implementation is based off the HAAR feature-based cascade classifiers method proposed by Viola-Jones in 2001.
 It's split into two phases: training and detection.
@@
 \par Each feature is obtained by comparing the pixel intensity in the white boxes against the pixel intensity in the dark boxes for a wide sample of sub-windows across the entire image.
-A 24x24 pixel image window can produce over 160,000 features that need to be classified, in some sense, “oversampling” the face.
+A 24 $\times$ 24 pixel image window can produce over 160,000 features that need to be classified--in some sense, “oversampling” the face.
 \par The rectangle features can be computed by splitting the image into sub-windows called integral images.
@@
 \]
-\begin{figure}[h]
-	\centering
-	\includegraphics[width=0.45\textwidth]{integral_image.png}
-	\caption{Integral image on a face}
-    \label{integral-image-face}
-\end{figure}
 Since an integral image includes all previously-computed integral images contained in its window, in the image of Figure \ref{integral-image-face}, the region A can be computed as
@@
 Any such rectangular sum can be computed using four array references.
+\begin{figure}[h]
+	\centering
+	\includegraphics[width=0.45\textwidth]{integral_image.png}
+	\caption{Integral image on a face}
+    \label{integral-image-face}
+\end{figure}
 
 \subsection{Cascade Classifier Training}
-It's far too resource intensive to evaluate over 160,000 images for each 24x24 pixel window.
+It's far too resource intensive to evaluate over 160,000 images for each 24 $\times$ 24 pixel window.
 However, we can take advantage of the fact that only a very small fraction of extracted sub-windows in an uncropped image will correspond to accurate facial features.
@@
 \par We load the trained classifier, load our image, and then the object detect functions returns the position of any found faces as a Rectangle with identifiers (x, y, width, height).
-With the returned positions, we can draw a basic rectangle of the detected facial regions or even perhaps take it a step further and apply Daft Punk masks to any detected faces.
+With the returned positions, we can draw a basic rectangle of the detected facial regions or even perhaps take it a step further and apply Daft Punk masks to any detected faces (Figure \ref{daft-punk-masks}).
 \begin{figure}[H]
@@
 \par Unlike previously with the edge detection and motion detection, where there is no singularly accepted CUDA implementation of the aforementioned algorithms, with facial recognition, OpenCV's library has been heavily optimized over the years to provide that functionality.
-\par It would an interesting case study then to abstract the functionalities of their object detection libraries, and compare the speedups attained from using a heavily optimized 3rd party library against the speedups which we obtained from programs written by us from scratch.
+\par It would an interesting case study then to abstract the functionalities of their object detection libraries, and compare the speedups attained from using a heavily optimized third party library against the speedups which we obtained from programs written by us from scratch.
 \par OpenCV's CUDA implementation is similar to its serial CPU implementation, except that the cascade classifier is loaded in a CUDA optimized format, and the images are loaded and processed in CUDA-optimized matrices formats created by OpenCV.
